### 2021-11-08 code/analysis.Rmd
add prelim analysis

---
title: "Game Theory"
output:
  html_document: 
    toc: true
---

#TODO
- for less munging, put player role into round
- put players increment into record!
# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/pilot211108"

date_start=lubridate::ymd('2021-11-08')

#image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This was for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>% 
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>% 
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(5.33+bonus*4/3,2)) %>% 
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))

```

```{r, include=F, eval=F}

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
    filter(createdAt >= date_start)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role)

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  left_join(d.players) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2)) %>% 
  write_csv(here(data_location, 'raw_results.csv'))

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Metrics
Welp, people definitely took the "click through" approach ...

```{r}
d.round_results <- read_csv(here(data_location,"raw_results.csv"))

summary <- d.round_results %>% group_by(repNum, gameId) %>% 
           mutate(time= time %|% 180) %>% 
  summarize(max_time=max(time)) %>% 
  group_by(gameId) %>% 
  summarize(total_time=sum(max_time)/60,
            num_rounds=max(repNum))

summary

```

Not much talking, mostly one person

```{r}
chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text))

chat %>% select(gameId,playerId,repNum,text)

```

# What do people choose

win > coop > defect > 0
   "AA":{p1:coop,p2:coop},
   "AB":{p1:0,p2:win},
   "BA":{p1:win,p2:0},
   "BB":{p1:defect,p2:defect}

```{r}
outcome <- d.round_results %>% select(gameId:repNum, response,role) %>% 
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))

ggplot(outcome, aes(x=repNum, y=outcome, color=gameId))+geom_jitter(height=.1)

```

# What do they say about the game

```{r}
  read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId,human, workedWell, fair, chatUseful) %>% filter(!is.na(gameId))

```

### 2021-11-12 code/analysis.Rmd
edits for BoS

@@
 
-#TODO
+# TODO
 - for less munging, put player role into round
 - put players increment into record!
+
+- Raise the stakes (game length / pay / bonus)
+
+Next pilot 
+- second page of instructions about bonus v base pay
+- 40 trials (~15 minutes) aim for $3-4 bonus 
+- can you have a get to know you at the beginning for 30-60 seconds? 
+- switch to stag hunt? or other coord game (BoS)
+
+- switch to BoS -- what are payoffs? (A,B; 0,0)
+
 # Pre-process
@@
 # Metrics
+Paid $4 based on 20 minutes (including instructions, waiting, game, and end survey).
+
 Welp, people definitely took the "click through" approach ...
@@
 
+```
+
+```{r}
+pairs <- d.round_results %>% select(playerId,gameId) %>% unique()
+d.players <- read_csv(here(data_location, 'players.csv')) %>% select(playerId=`_id`,data.bonus) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)
+
+d.players
 ```
@@
 ```{r}
-  read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId,human, workedWell, fair, chatUseful) %>% filter(!is.na(gameId))
+  comments <- read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId, human, workedWell, fair, chatUseful) %>% filter(!is.na(gameId)) %>% arrange(gameId)
+
+comments %>% select(playerId, human, workedWell)
+
+comments %>% select(playerId,fair)
+
+comments %>% select(playerId, chatUseful)
 

### 2021-11-12 experiments/client/main.js

@@ -31,7 +31,7 @@ Empirica.consent(Consent);
 // At this point they have been assigned a treatment. You can return
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
-  const steps = [Overview];
+  const steps = [Payment, Overview];
   
 
   return steps;

### 2021-11-12 experiments/server/callbacks.js

@@ -62,6 +62,7 @@ Empirica.onStageStart((game, round, stage) => {
 // It receives the same options as onRoundEnd, and the stage that just ended.
 Empirica.onStageEnd((game, round, stage) => {
   if (stage.name=="selection"){
+    const scale=game.treatment.scale
     const players = game.players;
     const p1 = _.find(game.players, p => p.get('role') === "p1");
     const p2 = _.find(game.players, p => p.get('role') === "p2");
@@ -77,9 +78,11 @@ Empirica.onStageEnd((game, round, stage) => {
     players.forEach(player => {
       const currScore=player.get("bonus")
       const scoreIncrement=player.get("scoreIncrement")
-      player.set("bonus", scoreIncrement*.01 + currScore);
+      player.set("bonus", scoreIncrement*.01*scale + currScore);
       round.set('player_' + player._id + '_response', player.get('clicked'));
       round.set('player_' + player._id + '_time', player.get('timeClick'));
+      round.set('player_' + player._id + '_payoff', player.get('scoreIncrement'));
+      round.set('player_' + player._id + '_role', player.get('role'));
     });
   }
 }

### 2021-11-12 experiments/server/main.js

@@ -38,6 +38,22 @@ function createPDRewards(){
   return payoffs
 }
 
+function createBoSRewards(){
+  //off diagonal fixed at 0
+  // other rewards on 2-9 symmetric
+  var a=_.slice(_.shuffle(_.range(2,10)),0,2).sort()
+  var best=a[1]
+  var okay=a[0]
+  var miss=0
+  var payoffs={
+   "AA":{p1:best,p2:okay},
+   "AB":{p1:miss,p2:miss},
+   "BA":{p1:miss,p2:miss},
+   "BB":{p1:okay,p2:best}
+  }
+  return payoffs
+}
+
 // gameInit is where the structure of a game is defined.  Just before
 // every game starts, once all the players needed are ready, this
 // function is called with the treatment and the list of players.  You
@@ -80,7 +96,7 @@ Empirica.gameInit((game, treatment) => {
       // Loop through targets in block   
         const round = game.addRound();
         round.set('targets', chooseTargets(targets));
-        round.set('payoffs', createPDRewards());
+        round.set('payoffs', createBoSRewards());
         round.set('repNum', repNum)        
         round.addStage({
           name: "selection",

### 2021-11-13 code/analysis.Rmd
avoid committing files with Prolific ID

--- 2021-11-08 code/analysis.Rmd
@@
 ---
-title: "Game Theory"
+title: "BoS pilot"
 output:
@@
 
-#TODO
-- for less munging, put player role into round
-- put players increment into record!
+# TODO
+
+- why did time recording screw up for one group???
+
 # Pre-process
@@
 ##Data import constants
-data_location="data/pilot211108"
+data_location="data/pilot211112"
 
-date_start=lubridate::ymd('2021-11-08')
+date_start=lubridate::ymd('2021-11-12')
 
@@
 #          bonus=bonus %|% 0,
-#          cost=round(5.33+bonus*4/3,2)) %>% 
+#          cost=round(3.33+bonus*4/3,2)) %>%
 #   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))
@@
   unnest(payoffs) %>% 
-  left_join(d.players) %>% 
   mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
@@
 # Metrics
-Welp, people definitely took the "click through" approach ...
+Paid $2.50 for 15 minutes with up to $4 bonus stated.
+
+Something weird happened with time recording for Hy -- it shouldn't be possible to have times > 180 for each one, or > 120 total... Maybe it was just a time recording issue??
 
@@
 ```
+```{r}
 
-Not much talking, mostly one person
+ggplot(d.round_results, aes(x=repNum, y=time, color=playerId))+geom_point()
+
+```
+
+Some groups come close to maxxing bonuses
+
+```{r}
+pairs <- d.round_results %>% select(playerId,gameId) %>% unique()
+d.players <- read_csv(here(data_location, 'players.csv')) %>% select(playerId=`_id`,data.bonus) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)
+
+d.players
+```
+
+Number of utterances
 
@@
 
-chat %>% select(gameId,playerId,repNum,text)
+chat %>% group_by(gameId, playerId) %>% tally()
+
+```
+Looks like of the 5 groups:  3 talked (a little) / took longer / got more bonus and 2 didn't.
+
+When they talk
+
+```{r}
+chat %>% group_by(gameId, playerId, repNum) %>% tally() %>% filter(n>0) %>% 
+  ggplot(aes(x=repNum, y=playerId, color=gameId))+geom_point()
+
+```
+
+```{r}
+chat %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
+  select(gameId, playerId, repNum, text)
 
@@
 # What do people choose
+   "AA":Best for P1
+   "BB":Best for P2
 
-win > coop > defect > 0
-   "AA":{p1:coop,p2:coop},
-   "AB":{p1:0,p2:win},
-   "BA":{p1:win,p2:0},
-   "BB":{p1:defect,p2:defect}
+Looks like there's convergence by 20 for the groups that converge, but it may not be obvious they have -- 40 rounds looks right.
 
@@
 
-ggplot(outcome, aes(x=repNum, y=outcome, color=gameId))+geom_jitter(height=.1)
+ggplot(outcome, aes(x=repNum, y=outcome, color=gameId))+geom_jitter(height=.1)+facet_wrap(.~gameId)
+
+```
+
+Player 2 prefers option B, player 1 prefers action A. In one of the non-talky games, P2 usually choose their better one.
+
+```{r}
+outcome <- d.round_results %>% select(gameId:repNum, response,role)
+
+ggplot(outcome, aes(x=role, fill=response))+geom_bar(position="dodge")+facet_wrap(.~gameId)
 
@@
 
+... I guess maybe we do need a free chat at the beginning so people will think they're playing with another human...
+
 ```{r}
-  read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId,human, workedWell, fair, chatUseful) %>% filter(!is.na(gameId))
+  comments <- read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId, human, workedWell, fair, chatUseful) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% filter(!is.na(gameId)) %>% arrange(gameId)
+
+comments %>% select(playerId, human, workedWell)
+
+comments %>% select(playerId,fair)
+
+comments %>% select(playerId, chatUseful)
 

### 2021-11-30 code/analysis.Rmd
add initial chat /turing test

@@
 library(viridis)
+library(knitr)
+
 rstan_options(auto_write = TRUE)
@@
 chat %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
-  select(gameId, playerId, repNum, text)
+  select(gameId, playerId, repNum, text) %>% kable()
 

### 2021-11-30 experiments/client/main.js

@@ -25,7 +25,7 @@ import customBreadcrumb from "./game/Breadcrumb.jsx";
  Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
-Empirica.consent(Consent);
+//Empirica.consent(Consent);
 
 // Introduction pages to show before they play the game (optional).
 // At this point they have been assigned a treatment. You can return
@@ -34,8 +34,8 @@ Empirica.introSteps((game, treatment) => {
   const steps = [Payment, Overview];
   
 
-  return steps;
-  //return [];
+  //return steps;
+  return [];
 });
 
 // The Round component containing the game UI logic.

### 2021-11-30 experiments/server/main.js

@@ -90,6 +90,11 @@ Empirica.gameInit((game, treatment) => {
   // Make role list
     game.set('roleList', createRoles(_.map(game.players, '_id')));
 
+    const round=game.addRound()
+    round.addStage({ name: "turing",
+    displayName: "Chat",
+    durationInSeconds: treatment.turingDuration
+  })
     // Loop through repetition blocks
     _.times(reps, repNum => {
         //mixed_targets=_.shuffle(targets)

### 2021-12-08 experiments/client/main.js
fixes to turning

@@ -25,7 +25,7 @@ import customBreadcrumb from "./game/Breadcrumb.jsx";
  Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
-//Empirica.consent(Consent);
+Empirica.consent(Consent);
 
 // Introduction pages to show before they play the game (optional).
 // At this point they have been assigned a treatment. You can return
@@ -34,8 +34,8 @@ Empirica.introSteps((game, treatment) => {
   const steps = [Payment, Overview];
   
 
-  //return steps;
-  return [];
+  return steps;
+  //return [];
 });
 
 // The Round component containing the game UI logic.

### 2021-12-09 code/analysis.Rmd
new pilot now with pre-chat

@@
 # TODO
-
-- why did time recording screw up for one group???
 
@@
 ##Data import constants
-data_location="data/pilot211112"
+data_location="data/pilot211209"
 
-date_start=lubridate::ymd('2021-11-12')
+date_start=lubridate::ymd('2021-12-09')
 
@@
 #          bonus=bonus %|% 0,
-#          cost=round(3.33+bonus*4/3,2)) %>%
+#          cost=round(bonus*4/3,2)) %>%
 #   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))
@@
 
-Something weird happened with time recording for Hy -- it shouldn't be possible to have times > 180 for each one, or > 120 total... Maybe it was just a time recording issue??
+Sigh for those couple games.
 
@@
 
-ggplot(d.round_results, aes(x=repNum, y=time, color=playerId))+geom_point()
+#ggplot(d.round_results, aes(x=repNum, y=time, color=playerId))+geom_point()
 
@@
 
+pre-chat -- really we want word count not utterances, but eh. 
+From skimming people do talk some. 
 ```{r}
-chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text))
+#this is a sloppy way to split, but eh, 
+chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(is.na(targets))
 
 chat %>% group_by(gameId, playerId) %>% tally()
+```
+
+```{r}
+chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(!is.na(targets))
+
+chat %>% group_by(gameId, playerId) %>% tally() %>% full_join(d.players)
 
 ```
-Looks like of the 5 groups:  3 talked (a little) / took longer / got more bonus and 2 didn't.
 
@@
 ```
-
-```{r}
-chat %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
-  select(gameId, playerId, repNum, text) %>% kable()
-
-```
+(Full chat transcript at bottom)
 
@@
 
-Player 2 prefers option B, player 1 prefers action A. In one of the non-talky games, P2 usually choose their better one.
+Player 2 prefers option B, player 1 prefers action A.
 
@@
 # What do they say about the game
-
-... I guess maybe we do need a free chat at the beginning so people will think they're playing with another human...
 
@@
 ```
+# Chat
+
+```{r}
+read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
+  select(gameId, playerId, repNum, text) %>% kable()
+
+```

### 2022-01-05 experiments/server/main.js
tweaks for 2x2

@@ -54,6 +54,12 @@ function createBoSRewards(){
   return payoffs
 }
 
+function createRewards(game){
+  //wrapper
+  if (game=="PD") return createPDRewards()
+  if (game=="BoS") return createBoSRewards()
+}
+
 // gameInit is where the structure of a game is defined.  Just before
 // every game starts, once all the players needed are ready, this
 // function is called with the treatment and the list of players.  You
@@ -101,7 +107,7 @@ Empirica.gameInit((game, treatment) => {
       // Loop through targets in block   
         const round = game.addRound();
         round.set('targets', chooseTargets(targets));
-        round.set('payoffs', createBoSRewards());
+        round.set('payoffs', createRewards(treatment.gameType));
         round.set('repNum', repNum)        
         round.addStage({
           name: "selection",

### 2022-01-07 code/analysis.Rmd
study 1 2x2 data

@@
 ---
-title: "BoS pilot"
+title: "Study 1"
 output:
@@
 ---
+TODO notes:
 
-# TODO
+ - get empirica time recording to behave (grumble)
+ - read feedback
+ - remove incomplete game(s)
+
+# Study 1 notes
+
+2x2 study of BoS vs PD crossed with chat or no-chat. All did 40 trials. All did a 3 minute "turing-test" free chat at the start. Run Jan 7. Target was 20 pairs in each condition.
 
@@
 ##Data import constants
-data_location="data/pilot211209"
+data_location="data/study1_220107"
 
-date_start=lubridate::ymd('2021-12-09')
+date_start=lubridate::ymd('2022-01-07')
 
@@
 #          cost=round(bonus*4/3,2)) %>%
-#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>% write_csv(here(data_location,"for_prolific.csv"))
+#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
+#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
+
+```
+
+```{r}
+
+d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% 
+  rename(treatmentId=`_id`) %>% 
+  select(treatmentId, name) %>% 
+  separate(name, c("game_cond", "chat_cond"))
+
+d.games <- read_csv(here(data_location, 'games.csv')) %>% 
+  rename(gameId = `_id`) %>% 
+  left_join(d.treatments) %>% 
+    filter(createdAt >= date_start) %>% 
+  select(gameId,playerIds,game_cond,chat_cond)
+
+d.gameplayers <- d.games %>% separate(playerIds, c("p1","p2")) %>% 
+  pivot_longer(p1:p2, values_to="playerId") %>% 
+  select(-name)
+
+d.players <- read_csv(here(data_location, 'players.csv')) %>% 
+  select(playerId=`_id`,role=data.role, data.bonus) %>% 
+  left_join(d.gameplayers)
 
@@
 ```{r, include=F, eval=F}
-
-d.games <- read_csv(here(data_location, 'games.csv')) %>% 
-  rename(gameId = `_id`) %>% 
-    filter(createdAt >= date_start)
-
-d.players <- read_csv(here(data_location, 'players.csv')) %>% 
-  select(playerId=`_id`,role=data.role)
 
@@
   unnest(data.chat) %>% 
-  select( -ends_with('response'), -ends_with('time')) %>%
+  select( -ends_with('response'), -ends_with('time'), -ends_with('payoff'), -ends_with('role')) %>%
   rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
+  left_join(d.games) %>% 
   write_csv(here(data_location, 'raw_chat.csv'))
@@
          BB=ifelse(role=="p1", BB.p1, BB.p2)) %>% 
+  left_join(d.games) %>% 
   write_csv(here(data_location, 'raw_results.csv'))
@@
 
-# Metrics
-Paid $2.50 for 15 minutes with up to $4 bonus stated.
+# Game length
+Estimated 15 mins total including intro/exit/waiting turing etc
 
-Sigh for those couple games.
+How long did people spend playing? (Not counting intro chat round)
 
-```{r}
+```{r total-time}
+# game KxJzRQeo5HikZxreo, player BkAWA2P83AnBmRcif has consistently unreasonable times (3300+, should be <180, so something's messing up)
+
+d.round_results %>% filter(gameId=="KxJzRQeo5HikZxreo") %>% select(repNum, time, playerId) %>% arrange(time)
 d.round_results <- read_csv(here(data_location,"raw_results.csv"))
 
-summary <- d.round_results %>% group_by(repNum, gameId) %>% 
-           mutate(time= time %|% 180) %>% 
+summary <- d.round_results %>% group_by(repNum, gameId, game_cond, chat_cond) %>% 
+  filter(time<181) %>% 
   summarize(max_time=max(time)) %>% 
-  group_by(gameId) %>% 
+  group_by(gameId, game_cond, chat_cond) %>% 
   summarize(total_time=sum(max_time)/60,
-            num_rounds=max(repNum))
+            num_rounds=max(repNum)) %>% 
+  filter(num_rounds==39)
 
-summary
-
-```
-```{r}
-
-#ggplot(d.round_results, aes(x=repNum, y=time, color=playerId))+geom_point()
-
+ggplot(summary, aes(x=str_c(game_cond," ", chat_cond), y=total_time, color=str_c(game_cond," ", chat_cond)))+geom_jitter(width=.1)+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+
+  labs(title="Game time in min", y="mins", x="condition")+theme(legend.position = "none")
 ```
 
-Some groups come close to maxxing bonuses
+# Bonuses
+
+Note that bonuses are not directly comparable between BoS & PD! 
+But it looks like chat helps in BoS and not in PD.
 
@@
 pairs <- d.round_results %>% select(playerId,gameId) %>% unique()
-d.players <- read_csv(here(data_location, 'players.csv')) %>% select(playerId=`_id`,data.bonus) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)
+bonuses <- d.players %>%inner_join(pairs) %>%  select(playerId,data.bonus, chat_cond, game_cond) %>% filter(!is.na(data.bonus)) %>% left_join(pairs) %>% arrange(gameId)
 
-d.players
+ggplot(bonuses, aes(x=str_c(game_cond," ", chat_cond), y=data.bonus, color=str_c(game_cond," ", chat_cond)))+geom_jitter(width=.1)+
+  labs(title="Bonus earned", y="bonus", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+scale_y_continuous(limits=c(0,6))
 ```
 
-Number of utterances
-
-pre-chat -- really we want word count not utterances, but eh. 
-From skimming people do talk some. 
+# Pre-chat 
 ```{r}
 #this is a sloppy way to split, but eh, 
-chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(is.na(targets))
+chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(is.na(targets)) %>% 
+  mutate(words=str_count(text, "\\w+")) %>%  group_by(gameId, playerId, chat_cond, game_cond) %>% summarize(words=sum(words))
 
-chat %>% group_by(gameId, playerId) %>% tally()
+ggplot(chat, aes(x=str_c(game_cond," ", chat_cond), y=words))+geom_jitter(width=.1)+
+  labs(title="Words produced in pre-chat", y="words", x="condition")
 ```
 
+# During chat
+
+This chat is raw. There are outliers that have been chopped off. 
+People talk more during BoS than PD.
+
 ```{r}
-chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% filter(!is.na(targets))
+game_chat <- read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(targets)) %>% filter(chat_cond=="chat") %>% 
+  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, game_cond, repNum) %>% summarize(words=sum(words))
 
-chat %>% group_by(gameId, playerId) %>% tally() %>% full_join(d.players)
+chat_by_game <- game_chat %>% group_by(gameId, game_cond, repNum) %>% summarize(words=sum(words))
+
+is_chat <- chat_by_game %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(game_cond, repNum) %>% summarize(pct_chat=mean(is.chat))
+
+ggplot(game_chat, aes(x=repNum, y=words, color=game_cond))+geom_jitter(width=.1, size=1, alpha=.2)+facet_wrap(~game_cond)+scale_y_continuous(limits=c(0,30))+
+  geom_smooth()+
+  labs(title="Words produced per round per player", y="words")
+
+ggplot(chat_by_game, aes(x=repNum, y=words, color=game_cond))+geom_jitter(width=.1, size=1, alpha=.2)+facet_wrap(~game_cond)+scale_y_continuous(limits=c(0,30))+
+  geom_smooth()+
+  labs(title="Words produced per round per game", y="words")
+
+ggplot(is_chat, aes(x=repNum, y=pct_chat, color=game_cond))+geom_point()+
+  labs(title="Pct games who said anything", y="pct")
+```
+Currently open questions -- is that game-related chat, or is it just that it's a friendlier game? 
+Does the chatting help?
+
+To look at if chatting helps -- we can compare forced no-chat to chat (see bonuses above). We can also compare used chat (0/1 or amount) within those that could chat.
+
+Options for dependent variable are bonus (although there's some noise there) or option chosen.
+
+```{r}
+
+group_chat <- chat_by_game %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(game_cond, gameId) %>% summarize(pct_chat=mean(is.chat))
+
+all <- d.players %>% left_join(group_chat) %>% filter(data.bonus>0)
+
+ggplot(all, aes(x=pct_chat, y=data.bonus, color=game_cond))+geom_point()
 
 ```
-
-When they talk
-
-```{r}
-chat %>% group_by(gameId, playerId, repNum) %>% tally() %>% filter(n>0) %>% 
-  ggplot(aes(x=repNum, y=playerId, color=gameId))+geom_point()
-
-```
-(Full chat transcript at bottom)
+Looks like chatting helps for BoS, doesn't get used much for PD.
 
 # What do people choose
-   "AA":Best for P1
-   "BB":Best for P2
 
-Looks like there's convergence by 20 for the groups that converge, but it may not be obvious they have -- 40 rounds looks right.
+In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both.
+
+In PD:
 
@@
 
-Player 2 prefers option B, player 1 prefers action A.
-
 ```{r}
@@
 ```
-
-# What do they say about the game
-
-```{r}
-  comments <- read_csv(here(data_location,'exit.csv')) %>% select(playerId, gameId, human, workedWell, fair, chatUseful) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% filter(!is.na(gameId)) %>% arrange(gameId)
-
-comments %>% select(playerId, human, workedWell)
-
-comments %>% select(playerId,fair)
-
-comments %>% select(playerId, chatUseful)
-
-```
-# Chat
-
-```{r}
-read_csv(here(data_location, "raw_chat.csv")) %>% filter(!is.na(text)) %>% mutate(gameId=str_sub(gameId, 0,2),playerId=str_sub(playerId, 0,2)) %>% 
-  select(gameId, playerId, repNum, text) %>% kable()
-
-```

### 2022-01-08 code/analysis.Rmd
2x2 analysis

@@
 
-d.round_results %>% filter(gameId=="KxJzRQeo5HikZxreo") %>% select(repNum, time, playerId) %>% arrange(time)
 d.round_results <- read_csv(here(data_location,"raw_results.csv"))
@@
   labs(title="Game time in min", y="mins", x="condition")+theme(legend.position = "none")
+
+summary %>% group_by(game_cond, chat_cond) %>% tally()
 ```
+A respectable rate of game completion.
 
@@
 Note that bonuses are not directly comparable between BoS & PD! 
-But it looks like chat helps in BoS and not in PD.
+But it looks like chat helps in BoS and not (much?) in PD.
 
@@
 ```
+People talked some. I haven't looked at the transcripts at all.
 
@@
 
+# Does chat help?
+
 To look at if chatting helps -- we can compare forced no-chat to chat (see bonuses above). We can also compare used chat (0/1 or amount) within those that could chat.
@@
 
-# What do people choose
+## Choices
 
@@
 
-In PD:
+In PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. (BB is Nash equilibrium, AA is Pareto dominant.)
+
+Not really sure how to visualize this efficiently?
+
+Being in the chat condition v not
 
 ```{r}
-outcome <- d.round_results %>% select(gameId:repNum, response,role) %>% 
+outcome <- d.round_results %>% select(gameId:repNum, response,role, game_cond, chat_cond) %>% 
   pivot_wider(names_from=role, values_from=response) %>% 
@@
 
-ggplot(outcome, aes(x=repNum, y=outcome, color=gameId))+geom_jitter(height=.1)+facet_wrap(.~gameId)
+ggplot(outcome, aes(x=repNum, y=outcome))+geom_jitter(width=.1, alpha=.25)+facet_grid(game_cond~chat_cond)+theme(legend.position="none")
+
+outcome_summary <- outcome %>% group_by(game_cond, chat_cond, repNum,outcome) %>% tally() %>% group_by(game_cond, chat_cond, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)
+
+ggplot(outcome_summary, aes(x=repNum, y=pct, color=chat_cond))+geom_line()+facet_grid(game_cond~outcome)
 
@@
 
+Actually using the chat that round
+
 ```{r}
-outcome <- d.round_results %>% select(gameId:repNum, response,role)
+outcome <- d.round_results %>% select(gameId:repNum, response,role, game_cond, chat_cond) %>% 
+  pivot_wider(names_from=role, values_from=response) %>% 
+  mutate(outcome=str_c(p1,p2)) %>% filter(chat_cond=="chat") %>% left_join(chat_by_game) %>% mutate(is.chat=ifelse(words>0, 1,0))
 
-ggplot(outcome, aes(x=role, fill=response))+geom_bar(position="dodge")+facet_wrap(.~gameId)
+outcome_summary <- outcome %>% group_by(game_cond, is.chat, repNum,outcome) %>% tally() %>% group_by(game_cond, is.chat, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)
+
+ggplot(outcome_summary, aes(x=repNum, y=pct, color=as.factor(is.chat)))+geom_line()+facet_grid(game_cond~outcome)
 

### 2022-02-05 code/analysis.Rmd
start text analysis

@@
  - read feedback
- - remove incomplete game(s)
 
@@
 2x2 study of BoS vs PD crossed with chat or no-chat. All did 40 trials. All did a 3 minute "turing-test" free chat at the start. Run Jan 7. Target was 20 pairs in each condition.
+
+## Pre-reg
+According to https://osf.io/8fnze, the analyses I'm going to do:
+
+This is exploratory research, here are some things we plan to explore. 
+role of language on score. For each type of game, do games with language have higher overall scores ? Is this related to if language was actually used? 
+Quantity and type of language over rounds. How much language is produced each round in PD v BoS? Does it decline in later rounds? 
+What type of language is used? Is it just identifying what to select or is there meta-strategy/negotiation?
+Are pairs converging to predictable strategies?
 
@@
          BB=ifelse(role=="p1", BB.p1, BB.p2)) %>% 
-  left_join(d.games) %>% 
-  write_csv(here(data_location, 'raw_results.csv'))
+  left_join(d.games)
 
@@
   left_join(d.games, by = c('gameId')) %>%
-    rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
+    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
+
+d.complete <- d.round_results.raw %>% group_by(gameId, playerId) %>% 
+  summarize(max_rep=max(repNum)) %>% filter(max_rep==39) %>% select(gameId,playerId) %>% 
+  write_csv(here(data_location,"complete.csv"))
+
+#remove that one game that didn't complete
+d.chat <- d.chat.raw %>% inner_join(d.complete) %>% 
+  write_csv(here(data_location, 'raw_chat.csv'))
+d.round_results <- d.round_results.raw %>% inner_join(d.complete) %>% 
+  write_csv(here(data_location, 'raw_results.csv'))
+d.exit <- d.exit.survey %>% inner_join(d.complete) %>% 
   write_csv(here(data_location,'exit.csv'))
@@
   summarize(total_time=sum(max_time)/60,
-            num_rounds=max(repNum)) %>% 
-  filter(num_rounds==39)
+            num_rounds=max(repNum))
 
@@
 ```
+
 People talked some. I haven't looked at the transcripts at all.
@@
 ```
+
 Currently open questions -- is that game-related chat, or is it just that it's a friendlier game? 
@@
 
+## What are people saying?
+
+```{r}
+library(qdap)
+
+game_chat %>% ungroup() %>%  summarize(s=sum(words)) #2093 total words
+
+chat <-  read_csv(here(data_location, "raw_chat.csv")) %>%
+  filter(!is.na(targets)) %>% 
+  filter(chat_cond=="chat") %>% 
+  filter(!is.na(text)) %>% 
+  select(gameId,game_cond,repNum, text) %>% 
+  mutate(text=strip(text)) %>% 
+  mutate(length=str_count(text, "\\w+")) %>% 
+  filter(length>0)
+
+ggplot(chat, aes(length))+geom_bar()
+
+```
+
+Most utterances are single word utterances, we'll consider these and multi word utterances separately.
+
+```{r}
+#single world
+words <- chat %>% filter(length==1) %>% pull(text) %>% word_list() %>% pull(fwl)
+
+head(words$all, 20)
+#chat %>% filter(length==1) %>% nrow()
+colors <- c("green", "blue", "pink","purple", "yellow", "red", "orange", "brown", "grey","white", "salmon", "gray")
+
+chat %>% filter(length==1) %>% filter(text %in% colors) %>% nrow()
+```
+
+There are 743 singletons.
+Of these 100 are ok/okay, at least 433 are color words, more once you count rarities and misspellings.
+There's also a bunch of singleton letters -- these seem to be single letter color abbreviations?
+
+From skimming transcripts, looks like multi's occur earlier in general. But we're confounded by pleasantries at random times (talking about where they live, etc).
+
+```{r}
+
+chat %>% mutate(singleton=ifelse(length==1, "single", "multi")) %>% ggplot(aes(repNum, fill=singleton))+geom_bar()+facet_grid(singleton~game_cond)
+
+```
+
+```{r}
+chat %>% filter(length>1) %>% select(game_cond, repNum, text) %>% head(40) %>% knitr::kable()
+
+```
 # Does chat help?
@@
 ```
+
 Looks like chatting helps for BoS, doesn't get used much for PD.
@@
 ```
+
+# Is the other player a human?
+
+```{r bot}
+
+bot <- read_csv(here(data_location, "exit.csv")) %>% 
+  filter(!is.na(game_cond) & !is.na(chat_cond)) %>% 
+  group_by(game_cond, chat_cond, human) %>% 
+  tally() %>% 
+  pivot_wider(names_from=human, values_from=n)
+
+bot
+```
+
+Only 10-20% of players think the other person is a bot!

### 2022-02-17 code/analysis.Rmd
investigate & tweak timer, add PD/BoS mix

@@
 Only 10-20% of players think the other person is a bot!
+
+# Poking at time recording
+
+Conclusion -- doing things player side is just a bad idea! and the times are worse than thought. 
+```{r total-time}
+# game KxJzRQeo5HikZxreo, player BkAWA2P83AnBmRcif has consistently unreasonable times (3300+, should be <180, so something's messing up)
+
+d.round_results <- read_csv(here(data_location,"raw_results.csv"))
+
+summary <- d.round_results %>% group_by(repNum, gameId, game_cond, chat_cond) %>% 
+  filter(time<181) %>% 
+  summarize(max_time=max(time)) %>% 
+  group_by(gameId, game_cond, chat_cond) %>% 
+  summarize(total_time=sum(max_time)/60,
+            num_rounds=max(repNum))
+
+ggplot(summary, aes(x=str_c(game_cond," ", chat_cond), y=total_time, color=str_c(game_cond," ", chat_cond)))+geom_jitter(width=.1)+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+
+  labs(title="Game time in min", y="mins", x="condition")+theme(legend.position = "none")
+
+summary %>% group_by(game_cond, chat_cond) %>% tally()
+```
+
+Empirica recommends using stages.csv startTimeAt and players-stages.csv submittedAt
+
+```{r}
+d.stages <- read_csv(here(data_location,"stages.csv")) %>% rename(stageId=`_id`)
+
+d.player_stages <- read_csv(here(data_location,"player-stages.csv")) %>%
+  left_join(d.stages) %>% 
+  filter(!is.na(submittedAt)) %>% 
+  mutate(timedif=submittedAt-startTimeAt)
+
+d.player_stages %>% filter(playerId=="BkAWA2P83AnBmRcif") %>% View()
+
+blah <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
+  filter(createdAt >= date_start) %>% 
+  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
+  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
+    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
+    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
+  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
+  select(-chat) %>% 
+  gather(key, value, starts_with('player')) %>% 
+  separate(key, into = c('blah', 'playerId', 'info')) %>% 
+  spread(info, value) %>% 
+  rename(roundId=`_id`) %>% 
+  select(roundId, playerId, time)
+foobar <- d.player_stages %>% select(playerId, roundId, timedif) %>% left_join(blah) %>% mutate(time=as.numeric(time)/1000) %>% mutate(check=timedif-time)
+
+View(foobar)
+
+```

### 2022-02-17 experiments/server/callbacks.js

@@ -39,9 +39,10 @@ Empirica.onRoundStart((game, round) => {
     player.set("targets", _.shuffle(round.get("targets")));
 
     player.set('clicked', false);
-    player.set('timeClick', false);
     player.set("scoreIncrement",0)
   });
+  console.log(round)
+  console.log(players)
 });
 
 // onRoundStart is triggered before each stage starts.
@@ -80,7 +81,7 @@ Empirica.onStageEnd((game, round, stage) => {
       const scoreIncrement=player.get("scoreIncrement")
       player.set("bonus", scoreIncrement*.01*scale + currScore);
       round.set('player_' + player._id + '_response', player.get('clicked'));
-      round.set('player_' + player._id + '_time', player.get('timeClick'));
+      round.set('player_' + player._id + '_time', player.stage.submittedAt-stage.startTimeAt);
       round.set('player_' + player._id + '_payoff', player.get('scoreIncrement'));
       round.set('player_' + player._id + '_role', player.get('role'));
     });

### 2022-02-17 experiments/server/main.js

@@ -58,6 +58,10 @@ function createRewards(game){
   //wrapper
   if (game=="PD") return createPDRewards()
   if (game=="BoS") return createBoSRewards()
+  if (game=="mixPDBoS") {
+    if(Math.random()>.5) return createPDRewards()
+    else return createBoSRewards()
+  }
 }
 
 // gameInit is where the structure of a game is defined.  Just before

### 2022-04-14 code/mixed.Rmd
add study2 and analysis thereof

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

Future notes:

- would be better to have game type coded as "gametype" to avoid namespace collision
- TODO: trends over game time!
- TODO: transcript reading
- TODO: game time & feedback

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study2_220413"

date_start=lubridate::ymd('2022-04-13')

study1_loc="data/study1_220107"
#image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This is for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))

```

```{r, include=F, eval=F}

d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% 
  rename(treatmentId=`_id`) %>% 
  select(treatmentId, name) %>% 
  mutate(name=ifelse(name=="mix-chat", "mix-nochat", name)) %>% #VB 220413: turns out naming conventions matter...
  separate(name, c("game_cond", "chat_cond"))

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
  left_join(d.treatments) %>% 
    filter(createdAt >= date_start) %>% 
  select(gameId,playerIds,game_cond,chat_cond)

d.gameplayers <- d.games %>% separate(playerIds, c("p1","p2")) %>% 
  pivot_longer(p1:p2, values_to="playerId") %>% 
  select(-name)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role, data.bonus) %>% 
  left_join(d.gameplayers)

```

```{r, include=F, eval=F}

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>% 
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>% 
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time'), -ends_with('payoff'), -ends_with('role')) %>% 
  rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  left_join(d.games) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
    rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2),
         type=ifelse(AB==0 & BA==0, "BoS", "PD")) %>% 
  left_join(d.games)

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.complete <- d.round_results.raw %>% group_by(gameId, playerId) %>% 
  summarize(max_rep=max(repNum)) %>% filter(max_rep==39) %>% select(gameId,playerId) %>% 
  write_csv(here(data_location,"complete.csv"))

d.chat <- d.chat.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))
d.round_results <- d.round_results.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_results.csv'))
d.exit <- d.exit.survey %>% inner_join(d.complete) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Read data

```{r comparison data}
study1.chat <- read_csv(here(study1_loc,"raw_chat.csv")) %>% 
    mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)
study1.rounds <- read_csv(here(study1_loc,"raw_results.csv")) %>% 
  mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

study2.chat <- read_csv(here(data_location,"raw_chat.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)
study2.rounds <- read_csv(here(data_location,"raw_results.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

rounds <- study1.rounds %>% union(study2.rounds)
chat <- study1.chat %>% union(study2.chat)
```
# Summary of expts
All expts reported here are from pairs of participants who played 40 rounds of a game-theory type game. At the start, each pair had 3 minutes of free chat, and then played the game. In exp 1, pairs were assigned to either all-PD or all-BoS games crossed with either access to the chat during the game or not.

In exp 2, all pairs played a random mix of PD and BoS rounds with or without access to the chat.

The exact payoffs of each round were determined randomly.

Throughout, I split up the results of the mixed games by which type to compare to the pure games. Thus, while 6 game conditions were run, 8 conditions are discussed.

```{r}

chat %>% select(gameId, game_cond, chat_cond) %>% unique() %>% group_by(game_cond, chat_cond) %>% tally()

```

# Overall outcomes

To compare overall performance across these conditions, we can look at the average payoff for each player. Here we split the results of the mixed games by which game (PD or BoS).

```{r}
bonuses <- rounds %>%group_by(playerId,gameId, chat_cond,gametype,mix) %>% summarize(payoff=mean(payoff))

ggplot(bonuses, aes(x=str_c(mix," ",chat_cond), y=payoff, color=chat_cond))+geom_jitter(width=.1)+
  facet_grid(.~gametype)+
  labs(y="mean_reward", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+scale_y_continuous(limits=c(0,6))
```

PD performance appears constant across games and chat conditions. BoS performance benefits from having a chat in pure and mixed games.

# Pre-chat

We hope that players use the pre-chat to communicate with each other and convince themselves that the other person is a human. All games are the same at this point, so we don't expect differential amounts of talk.

```{r}
prechat <- chat %>% filter(!is.na(text)) %>% 
  filter(is.na(targets)) %>% mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, game_cond, chat_cond) %>% summarize(words=sum(words))

ggplot(prechat, aes(x=str_c(game_cond," ",chat_cond), y=words))+geom_jitter(width=.1, color="lightblue")+ stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+labs(y="pre-chat words / player", x="game type")
```

# During chat

This is raw chat, so it's unclear if these are referential expressions or chit chat or what.

People in pure PD games are less likely to talk than those on BoS or mixed games. Amount of talking seems to decline over rounds.

Because many of the talkings are short, we also binarize and look at whether any chat was produced in a given round.

Mixed games seem to fall in between the talking amounts of the pure games.

```{r}
game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, gametype,  mix, repNum) %>% summarize(words=sum(words))

chat_by_pair <- game_chat %>% group_by(gameId, gametype, mix, repNum) %>% summarize(words=sum(words))

is_chat <- chat_by_pair %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gametype, mix, repNum) %>% summarize(pct_chat=mean(is.chat))

# ggplot(game_chat, aes(x=repNum, y=words, color=mix))+geom_jitter(width=.1, size=1, alpha=.2)+facet_wrap(~gametype)+scale_y_continuous(limits=c(0,30))+
#   geom_smooth()+
#   labs(title="Words produced per round per player", y="words")

ggplot(chat_by_pair, aes(x=repNum, y=words, color=gametype))+geom_jitter(width=.1, size=1, alpha=.2)+facet_grid(mix~gametype)+scale_y_continuous(limits=c(0,30))+
  geom_smooth()+
  labs(title="Words produced per round per game", y="words")

ggplot(is_chat, aes(x=repNum, y=pct_chat, color=gametype))+geom_point()+
  facet_grid(gametype~mix)+
  labs(y="pct of games who said anything each round")
```

# Does chat help?

We look at how many rounds had non-zero chat (of those with the option to chat) compared to mean payoff. There's some serious sparsity, but it looks like talking helps for BoS.

```{r}
success <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, mix, payoff) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId, gametype, mix) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat))

ggplot(success, aes(x=mean_chat, y=mean_payout, color=gametype))+geom_point()+facet_wrap(~mix)

```

## Choices

Rather than collapsing over outcome (which has some randomness), we can look at per-round, which quadrant in the game people chose.

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both.

In PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. (BB is Nash equilibrium, AA is Pareto dominant.)

This is a lot of cells, so we break up the graphs into whether chat was unavailable, used, or not-used. Blue lines are at 25% indicating chance from random or uncoordinated clicking.

```{r}
outcome <- rounds %>% select(-playerId, -time,-payoff, -box_A, -box_B, -AA, -AB, -BA, -BB) %>%  
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))  %>% left_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(gametype, chat_cond, mix, is.chat, repNum,outcome) %>% tally() %>% group_by(gametype, chat_cond, mix, is.chat, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)

#ggplot(outcome_summary, aes(x=repNum, y=pct, color=as.factor(is.chat)), )+geom_point(size=1)+facet_grid(str_c(gametype,mix)~outcome)

ggplot(outcome_summary %>% filter(is.na(is.chat)), aes(x=repNum, y=pct))+
    geom_hline(yintercept=.25, color="blue", size=1)+
  geom_point(size=1)+
  facet_grid(str_c(gametype,mix)~outcome)+labs(title="Rounds without the option for chat")

ggplot(outcome_summary %>% filter((is.chat)==1), aes(x=repNum, y=pct))+
    geom_hline(yintercept=.25, color="blue", size=1)+
  geom_point(size=1)+
  facet_grid(str_c(gametype,mix)~outcome)+labs(title="Rounds that did use chat")

ggplot(outcome_summary %>% filter((is.chat)==0), aes(x=repNum, y=pct))+
    geom_hline(yintercept=.25, color="blue", size=1)+
  geom_point(size=1)+
  facet_grid(str_c(gametype,mix)~outcome)+labs(title="Rounds that could use chat but didn't")
```

### 2022-05-05 code/mixed.Rmd
work on analysis

@@
 - TODO: trends over game time!
-- TODO: transcript reading
 - TODO: game time & feedback
@@
 study1_loc="data/study1_220107"
-#image_location="write-ups/images"
+image_location="write-ups/images"
 
@@
 chat <- study1.chat %>% union(study2.chat)
+
+study1.exit <- read_csv(here(study1_loc,"exit.csv"))
+
+study2.exit <- read_csv(here(data_location, "exit.csv"))
+
+s1 <- read_csv(here(study1_loc,"raw_results.csv")) %>% 
+  select(chat_cond, game_cond, playerId) 
+players <- read_csv(here(data_location,"raw_results.csv")) %>% 
+  select(chat_cond, game_cond, playerId) %>% union(s1) %>% 
+  unique()
+exit <- study1.exit %>% union(study2.exit) %>% inner_join(players)
 ```
@@
 
-ggplot(bonuses, aes(x=str_c(mix," ",chat_cond), y=payoff, color=chat_cond))+geom_jitter(width=.1)+
+ggplot(bonuses, aes(x=str_c(mix,"\n",chat_cond), y=payoff, color=str_c(mix," ",chat_cond)))+geom_jitter(width=.1, alpha=.5)+
   facet_grid(.~gametype)+
-  labs(y="mean_reward", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+scale_y_continuous(limits=c(0,6))
+  scale_color_viridis(discrete=T)+
+  labs(y="mean_reward", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.7,color="black")+theme(legend.position = "none")+coord_cartesian(ylim=c(0,6))+labs(y="Mean reward / trial")+
+    theme(axis.text=element_text(size=12), 
+        axis.title=element_text(size=14),
+        strip.text=element_text(size=14) )
+
+ggsave(here(image_location, 'payoff.png'), width=6, height=3.5, device="png", type="cairo-png")
 ```
@@
 
-ggplot(prechat, aes(x=str_c(game_cond," ",chat_cond), y=words))+geom_jitter(width=.1, color="lightblue")+ stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+labs(y="pre-chat words / player", x="game type")
+ggplot(prechat, aes(x=str_c(game_cond,"\n",chat_cond), y=words))+geom_jitter(width=.1, color="lightblue")+ stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+labs(y="Pre-chat words / player", x="")+
+      theme(axis.text=element_text(size=12), 
+        axis.title=element_text(size=14),
+        strip.text=element_text(size=14) )
+
+ggsave(here(image_location, 'prechat.png'), width=6, height=3.5, device="png", type="cairo-png")
 ```
@@
 
-ggplot(chat_by_pair, aes(x=repNum, y=words, color=gametype))+geom_jitter(width=.1, size=1, alpha=.2)+facet_grid(mix~gametype)+scale_y_continuous(limits=c(0,30))+
+ggplot(chat_by_pair, aes(x=repNum, y=words, color=gametype))+geom_jitter(width=.1, size=1, alpha=.2)+facet_grid(mix~gametype)+coord_cartesian(ylim=c(0,15))+
   geom_smooth()+
-  labs(title="Words produced per round per game", y="words")
+      theme(axis.text=element_text(size=12), 
+        axis.title=element_text(size=14),
+        strip.text=element_text(size=14),
+        legend.position = "none")+labs(y="Words / pair / trial", x="Trial number")
 
+ggsave(here(image_location, 'words.png'), width=5, height=3.5, device="png", type="cairo-png")
+
+```
+
+```{r}
 ggplot(is_chat, aes(x=repNum, y=pct_chat, color=gametype))+geom_point()+
@@
   labs(y="pct of games who said anything each round")
+```
+
+```{r}
+# per game
+chatty <- game_chat %>% group_by(gameId, playerId, gametype,  mix, repNum) %>% mutate(did.chat=ifelse(words>0, 1,0))
+
+game_chatty <- chatty %>% group_by(gameId, gametype,mix, repNum)
 ```
@@
 
-ggplot(success, aes(x=mean_chat, y=mean_payout, color=gametype))+geom_point()+facet_wrap(~mix)
+ggplot(success, aes(x=mean_chat, y=mean_payout, color=gametype))+geom_point()+facet_wrap(~mix)+geom_smooth(method="lm")
 
@@
 ```
+
+# But what about language?
+
+```{r}
+
+game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% filter(!is.na(text)) %>% select(gameId,playerId, gametype,mix,repNum,text)
+
+per_game <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% filter(!is.na(text)) %>% select(gameId,role, gametype,mix,repNum,text) %>%  mutate(game_cond=ifelse(mix=="pure", gametype, "mix"),
+         words=str_count(text, "\\w+") %|% int(0)) %>% 
+  filter(words>0) %>% 
+  group_by(gameId,role,game_cond) %>% summarize(total_words=sum(words),total_utts=n()) %>% 
+  pivot_wider(names_from=role, values_from=c(total_words, total_utts), values_fill=0) %>% 
+  mutate(words_1=max(total_words_p1, total_words_p2),
+         words_2=min(total_words_p1, total_words_p2),
+         utts_1=max(total_utts_p1, total_utts_p2),
+         utts_2=min(total_utts_p1, total_utts_p2))
+
+words <- per_game %>% ungroup() %>% group_by(game_cond) %>% 
+  select(gameId,game_cond, total_words_p1, total_words_p2) %>% 
+  mutate(total=total_words_p1+total_words_p2,
+         n=row_number()) %>% 
+  pivot_longer(total_words_p1:total_words_p2)
+
+  ggplot(words,aes(x=reorder(n,total),y=value, fill=name))+geom_col()+facet_grid(game_cond~.)
+```
+
+So, types of language that could occur:
+- chitchat 
+- negotiation about strategy/ who gets to pick
+- someone naming / proposing a color to choose 
+- agreement / "ok"
+- other: discussion of mistakes, etc
+
+Could hand code a bunch of this -- spelling correction might be good for detecting color words!
+```{r}
+color_words <- c("red","blue", "green","pink","yellow","gray","grey","salmon","cyan","purple",
+                 "orange","white","brown","rose")
+
+agreement <- c("ok","yup","yes","yep","yea","okay")
+color_coded <- game_chat %>% select(text) %>%
+  mutate(has.color=str_detect(text, regex(paste(color_words, collapse="|"), ignore_case=T)),
+         is.color=ifelse(str_replace_all(text, "[^[:alnum:]]", "") %>% str_to_lower() %in% color_words, T,F),
+         is.ok = ifelse(str_replace_all(text, "[^[:alnum:]]", "") %>% str_to_lower() %in% agreement, T,F))
+
+color_coded %>% group_by(has.color,is.ok, is.color) %>% tally()
+```
+
+```{r}
+agreement <- c("ok","yup","yes","yep","yea","okay")
+
+color_pure <- read_csv(here("data/colorpure.csv"), guess_max = 2000) %>% mutate(gametype=game_cond)
+color_mixed <- read_csv(here("data/colormixed.csv"), guess_max=2000)
+
+color <- color_pure %>% union(color_mixed) %>% filter(!is.na(targets)) %>% filter(!is.na(text)) %>% 
+  mutate(is.color=ifelse(is.na(color),0,1), 
+         is.num= ifelse(str_detect(text, regex("[0-9]", ignore_case=T)),1,0),
+         is.ok = ifelse(str_replace_all(text, "[^[:alnum:]]", "") %>% str_to_lower() %in% agreement, 1,0)) %>% group_by(game_cond) %>% summarize(c=mean(is.color),num=mean(is.num),ok=mean(is.ok), total=n())
+```
+# Are they human?
+
+```{r bot}
+
+exit %>% 
+  group_by(game_cond, chat_cond, human) %>% 
+  tally() %>% 
+  pivot_wider(names_from=human, values_from=n) %>% 
+  mutate(pct=yes/(no+yes))
+
+```

### 2022-06-14 code/text_analysis.Rmd
first pass at sentence embedding, analysis

---
title: "Sbert"
output:
  html_document: 
    toc: true
---

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)
library(tidytext)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study2_220413"

date_start=lubridate::ymd('2022-04-13')

study1_loc="data/study1_220107"
image_location="write-ups/images"

#model_location="code/models"
```

```{r comparison data}
study1.chat <- read_csv(here(study1_loc,"raw_chat.csv")) %>% 
    mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)

study1.rounds <- read_csv(here(study1_loc,"raw_results.csv")) %>% 
  mutate(gametype=game_cond,
         mix="pure") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

study2.chat <- read_csv(here(data_location,"raw_chat.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)

study2.rounds <- read_csv(here(data_location,"raw_results.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

rounds <- study1.rounds %>% union(study2.rounds)
chat <- study1.chat %>% union(study2.chat)

```

```{r prep}

raw_text <- chat %>% left_join(rounds) %>% 
  filter(!is.na(repNum)) %>% 
  filter(!is.na(text)) %>% 
  select(gameId,repNum,text, playerId, gametype,mix,game_cond,box_A,box_B,payoff,response,AA,AB,BA,BB) %>% 
  unnest_tokens(sentence, input=text, token="sentences") %>% 
  filter(!is.na(sentence))

#sample <- raw_text %>% filter(gameId=="3WzEi9zkHF77vdTw5") %>% write_csv("sample.csv")
pre_sbert <- raw_text %>% write_csv("pre_sbert.csv")
```

# Interlude
This is where they need to get embedding with jupyter

# Post-wrangle

```{r, include=F, eval=F}
library(reticulate)
np <- import("numpy")
mat = np$load('post_sbert.npy')
saveRDS(mat,'post_sbert.RData')
```

```{r, include=F,}
ssb <- pre_sbert %>% bind_cols(readRDS('post_sbert.RData')  %>% as_tibble())

grouped <- ssb %>%
  group_by(gameId,repNum, playerId, gametype,mix,game_cond,box_A,box_B,payoff,response,AA,AB,BA,BB) %>% 
  summarize(across(starts_with('V'), mean)) %>% 
  ungroup()

F_mat <- grouped %>% select(starts_with("V")) %>% as.matrix() #Features
M_mat <- grouped %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
```

We average across vectors for what each person said each rep.

```{r helpers}
# note: cor expects features to be in columns so we transpose
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_within_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          .$repNum)) %>%
    mutate(rep1 = as.numeric(dim1), 
           rep2 = as.numeric(dim2)) 
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}
```

# Analysis

## Convergence within groups

look within a game, across all reps and playerIDs

```{r }

game_convergence <- M_mat %>% 
  group_by(gameId,game_cond) %>% 
  mutate(combinedId=str_c(playerId,repNum,sep="_")) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, into=c("playerId_1","repNum_1"), convert=T, sep="_") %>% 
  separate(dim2, into=c("playerId_2","repNum_2"), convert=T, sep="_") %>% 
  mutate(repDist=abs(repNum_1-repNum_2),
         earlierRep=ifelse(repNum_1>repNum_2,repNum_2,repNum_1),
         samePlayer=ifelse(playerId_1==playerId_2,1,0)) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  ungroup()

m_convergence <- lmer(sim ~ repDist*game_cond+samePlayer+earlierRep*game_cond + (1|gameId), data=game_convergence)

summary(m_convergence)
```

Unclear what interactions we should be including.

repDistance: reps that are further apart from each other are less similar (note this is confounded w/ earlier rep)

being said by the same person increases similarity

## Divergence between groups

```{r }

game_divergence <- M_mat %>% 
  group_by(game_cond) %>% 
  mutate(combinedId=str_c(gameId,repNum,sep="_")) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, into=c("gameId_1","repNum_1"), convert=T, sep="_") %>% 
  separate(dim2, into=c("gameId_2","repNum_2"), convert=T, sep="_") %>% 
  filter(gameId_1!=gameId_2) %>% 
  mutate(repDist=abs(repNum_1-repNum_2),
         earlierRep=ifelse(repNum_1>repNum_2,repNum_2,repNum_1)) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  ungroup()

m_divergence <- lm(sim ~ repDist*game_cond+earlierRep*game_cond, data=game_convergence)

summary(m_divergence)
```

There's more difference the further apart the reps are, and maybe slight differences between the game types, but we're not seeing the clear divergence patterns that arise in say tangrams. (Which makes sense if groups are mostly converging on a limited number of patterns we wouldn't expect difference between them.)

### 2022-06-14 code/text_analysis.Rmd
try add graphs

@@
   mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
+  filter(!is.na(samePlayer)) %>% 
   ungroup()
 
+```
+
+0 is not same speaker, 1 is same speaker.
+
+```{r}
+game_convergence %>% filter(repDist<5) %>% ggplot(aes(x=earlierRep, y=sim, color=repDist))+geom_point(alpha=.1)+geom_smooth()+facet_grid(game_cond~samePlayer)+labs(title="Rep distance <5")
+
+game_convergence %>% ggplot(aes(x=repDist, y=sim, color=earlierRep))+geom_point(alpha=.01)+geom_smooth()+facet_grid(game_cond~samePlayer)
+
+```
+
+We're not really seeing much, similarity is higher in rounds that are closer together, but if we only look at relatively close reps (within 5 of each other), there's not much by way of pattern.
+
+```{r, eval=F}
 m_convergence <- lmer(sim ~ repDist*game_cond+samePlayer+earlierRep*game_cond + (1|gameId), data=game_convergence)
@@
   ungroup()
+```
+
+```{r}
+game_divergence %>% filter(repDist==0) %>% ggplot(aes(x=earlierRep, y=sim))+geom_point(alpha=.1)+geom_smooth()+facet_grid(game_cond~.)+labs(title="Rep distance == 0")
+
+game_divergence %>% ggplot(aes(x=repDist, y=sim, color=earlierRep))+geom_point(alpha=.01)+geom_smooth()+facet_grid(game_cond~.)
+
+```
+
+Not seeing anything useful here.
+
+```{r, eval=F}
 

### 2022-06-15 experiments/server/main.js
add spiking version, and split PD variants

@@ -12,6 +12,11 @@ function createRoles(players) {
   return roles;
 }
 
+function compareNumbers(a, b) //deal with the fact that js defaults to lexicographic sorting....
+{
+    return a - b;
+}
+
 function chooseTargets(targets){
   //takes a list of flower objects, a list of utilties and a number to select
   //returns the context
@@ -26,6 +31,44 @@ function createPDRewards(){
   //sucker payoff fixed at 0
   // other rewards on 1-9, but have to respect ordering win > coop > defect 
   var a=_.slice(_.shuffle(_.range(1,10)),0,3).sort()
+  var coop=a[1]
+  var defect=a[0]
+  var win=a[2]
+  var payoffs={
+   "AA":{p1:coop,p2:coop},
+   "AB":{p1:0,p2:win},
+   "BA":{p1:win,p2:0},
+   "BB":{p1:defect,p2:defect}
+  }
+  return payoffs
+}
+function createEasyPDRewards(){
+  //sucker payoff fixed at 0
+  // other rewards on 1-12, but have to respect ordering win > coop > defect 
+  // we redraw until we have a set where 2*coop \ge win 
+  var a=_.slice(_.shuffle(_.range(1,13)),0,3).sort(compareNumbers)
+  while (2*a[1]<a[2]){
+    a=_.slice(_.shuffle(_.range(1,13)),0,3).sort(compareNumbers)
+  }
+  var coop=a[1]
+  var defect=a[0]
+  var win=a[2]
+  var payoffs={
+   "AA":{p1:coop,p2:coop},
+   "AB":{p1:0,p2:win},
+   "BA":{p1:win,p2:0},
+   "BB":{p1:defect,p2:defect}
+  }
+  return payoffs
+}
+function createHardPDRewards(){
+  //sucker payoff fixed at 0
+  // other rewards on 1-12, but have to respect ordering win > coop > defect 
+  // we redraw until we have a set where 2*coop < win 
+  var a=_.slice(_.shuffle(_.range(1,13)),0,3).sort(compareNumbers)
+  while (2*a[1]>=a[2]){
+    a=_.slice(_.shuffle(_.range(1,13)),0,3).sort(compareNumbers)
+  }
   var coop=a[1]
   var defect=a[0]
   var win=a[2]
@@ -54,19 +97,49 @@ function createBoSRewards(){
   return payoffs
 }
 
+function createSpikeBoSRewards(){
+  //off diagonal fixed at 0
[…]

### 2022-06-27 code/exp2_mixed.Rmd
data and prelim analysis of new study

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

# Pre-process

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

##Data import constants
data_location="data/study3_220627"

date_start=lubridate::ymd('2022-06-24')

image_location="write-ups/images"

#model_location="code/models"
```

```{r bonus}
##This is for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)+) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# 
# cost <- read_csv(here(data_location, "player_payments.csv"))
# 
# sum(cost$cost)
```

```{r, include=F, eval=F}

d.treatments <- read_csv(here(data_location, 'treatments.csv')) %>% 
  rename(treatmentId=`_id`) %>% 
  select(treatmentId, name) %>% 
  mutate(name=ifelse(name=="mix-chat", "mix-nochat", name)) %>% #VB 220413: turns out naming conventions matter...
  separate(name, c("game_cond", "chat_cond"))

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
  left_join(d.treatments) %>% 
    filter(createdAt >= date_start) %>% 
  select(gameId,playerIds,game_cond,chat_cond)

d.gameplayers <- d.games %>% separate(playerIds, c("p1","p2")) %>% 
  pivot_longer(p1:p2, values_to="playerId") %>% 
  select(-name)

d.players <- read_csv(here(data_location, 'players.csv')) %>% 
  select(playerId=`_id`,role=data.role, data.bonus) %>% 
  left_join(d.gameplayers)

```

```{r, include=F, eval=F}

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>% 
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>% 
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>% 
  select( -ends_with('response'), -ends_with('time'), -ends_with('payoff'), -ends_with('role')) %>% 
  rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  left_join(d.games) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
    rename(data.gametype=data.type) %>% #VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("player", "player_", .x, fixed=T)) %>% 
    rename_with ( ~ gsub("response", "_response", .x, fixed=T)) %>% 
  rename_with( ~ gsub("time", "_time", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah) %>% 
  select(-stageIds,-index,-createdAt,-submitted) %>% 
  filter(!is.na(response)) %>% 
  mutate(time=as.numeric(time)/1000) %>% 
  mutate(targets=map(targets, .f=ParseJSONColumn)) %>% 
  unnest(targets) %>% 
  mutate(image = gsub('/experiment/', '', image, fixed=TRUE),
        image = gsub('.jpeg', '', image, fixed=TRUE)) %>% 
  pivot_wider(names_from=label, names_prefix="box_", values_from=image) %>% 
    mutate(payoffs=map(payoffs, .f=ParseJSONColumn)) %>%
  unnest(payoffs) %>% 
  mutate(AA=ifelse(role=="p1", AA.p1, AA.p2),
         AB=ifelse(role=="p1", AB.p1, AB.p2),
         BA=ifelse(role=="p1", BA.p1, BA.p2),
         BB=ifelse(role=="p1", BB.p1, BB.p2),
         type=ifelse(AB==0 & BA==0, "BoS", "PD")) %>% 
  left_join(d.games)

d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.complete <- d.round_results.raw %>% group_by(gameId, playerId) %>% 
  summarize(max_rep=max(repNum)) %>% filter(max_rep==39) %>% select(gameId,playerId) %>% 
  write_csv(here(data_location,"complete.csv"))

d.chat <- d.chat.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))
d.round_results <- d.round_results.raw %>% inner_join(d.complete) %>% 
  write_csv(here(data_location, 'raw_results.csv'))
d.exit <- d.exit.survey %>% inner_join(d.complete) %>% 
  write_csv(here(data_location,'exit.csv'))

```

# Read data

```{r}
chat <- read_csv(here(data_location,"raw_chat.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,text, type, repNum, playerId, chat_cond, gametype,mix, game_cond, targets)
rounds <- read_csv(here(data_location,"raw_results.csv")) %>% 
  mutate(mix="mix") %>% 
  select(gameId,repNum, playerId,payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype,mix, role)

```

# Summary of experiment
In the expt reported here pairs of participants  played 40 rounds of a game-theory type game. At the start, each pair had 3 minutes of free chat, and then played the game. We recruited for 20 games in chat and 20 games in no-chat conditions.

 - 4 "spiked" BoS trials where one of the rewards is high (25-30) and the other is normal (3-7) These are hard coded to appear on certain trials (7,16,23,32 / 40 trials total).
  - ~16 normal BoS trials where both rewards are from [1-9]
  - ~10 normal PD trials where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] respecting this.
  - ~10 “sacrifice” PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff). All rewards are chosen from [1-12] respecting this. 
Aside from the 4 “spiked” trials that occur at set places, the others are determined by randomly drawing from a 16:10:10 ratio distribution, so games will vary in exact proportions and locations of these trials.

```{r}

chat %>% select(gameId, game_cond, chat_cond) %>% unique() %>% group_by(game_cond, chat_cond) %>% tally()

```

Fewer chat games than non-chat games -- not sure why the differential loss -- maybe due to order of running??

## Pre-reg'd analysis plan

This is exploratory research, here are comparisons we plan to make (i.e. with graphs).

- [x] Do games with the option to chat get better scores than games without (total, and subdivided by trial type)?

- [x] Which quadrant (combination of A,B responses) do players land in as a function of chat/nochat and trial condition? (Do games with chat choose the off-diagonal option in “sacrifice” PD trials?)

- [x] How much language do players use as a function of trial number and type?

- (More qualitative) What heuristics/strategies do players use? (gleaned from post-experiment question and chat logs) How successful are they and which ones require how much language to coordinate?

- Are there signs of conventionalizing on ways of describing trials / strategies beyond names for boxes?

## other stuff to explore

- need to do intent to treat analysis
- look at what players write as their strategy
- can we tell what heuristics are being used (by trial, as function of alternating, as function of score so far?)
- [it's a mess] correlation w/i game between outcomes on different trial types

# Overall outcomes

A coarse first pass is to look at how much bonus people earned, as bonus is propotional to points.

By trial type:

```{r}
bonuses <- rounds %>%group_by(playerId,gameId, chat_cond, gametype) %>% summarize(payoff=mean(payoff))

ggplot(bonuses, aes(x=str_c(chat_cond), y=payoff, color=str_c(chat_cond)))+geom_jitter(width=.1, alpha=.5)+
  facet_grid(.~gametype)+
  scale_color_viridis(discrete=T)+
  labs(y="mean_reward", x="condition")+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.7,color="black")+theme(legend.position = "none")+
  coord_cartesian(ylim=c(0,20))+
  labs(y="Mean reward / trial")+
    theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

Aggregate:

```{r}
rounds %>%group_by(playerId,gameId, chat_cond) %>% summarize(payoff=sum(payoff)) %>% ggplot(aes(x=chat_cond, y=payoff, color=str_c(chat_cond)))+geom_jitter(width=.1, alpha=.5)+
  scale_color_viridis(discrete=T)+
 stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.7,color="black")+theme(legend.position = "none")+
  labs(y="Total reward")+
    theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

Clear aggregate difference based on condition.

## Pre-chat

We hope that players use the pre-chat to communicate with each other and convince themselves that the other person is a human. All games are the same at this point, so we don't expect differential amounts of talk.

```{r}
prechat <- chat %>% filter(!is.na(text)) %>% 
  filter(is.na(targets)) %>% mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, game_cond, chat_cond) %>% summarize(words=sum(words))

ggplot(prechat, aes(x=str_c(game_cond,"\n",chat_cond), y=words))+geom_jitter(width=.1, color="lightblue")+ stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.3), size=.5,color="black")+theme(legend.position = "none")+labs(y="Pre-chat words / player", x="")+
      theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=14),
        strip.text=element_text(size=14) )

```

## During chat

Chatting seems to increase slightly across the games, but many groups even in the chat condition don't use the chat.

```{r}
game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, playerId, gametype,  mix, repNum) %>% summarize(words=sum(words))

chat_by_pair <- game_chat %>% group_by(gameId, gametype, mix, repNum) %>% summarize(words=sum(words))

is_chat <- chat_by_pair %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(repNum) %>% summarize(pct_chat=mean(is.chat))
ggplot(is_chat, aes(x=repNum, y=pct_chat))+geom_point()+geom_smooth()+
  labs(y="pct of games who said anything each round")
```

# Are they human?

```{r bot}

exit <- read_csv(here(data_location,'exit.csv'))

exit %>% 
  group_by(game_cond, chat_cond, human) %>% 
  tally() %>% 
  pivot_wider(names_from=human, values_from=n) %>% 
  mutate(pct=yes/(no+yes))

```

Yay, they mostly think they're human!!

# Chat and payoff

We look at how many rounds had non-zero chat (of those with the option to chat) compared to mean payoff.

```{r}
overall_success <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(overall_success, aes(x=mean_chat, y=mean_payout))+geom_point()+geom_smooth(method="lm")+labs(y="Mean Payout", x="Proportion round with nonzero chat")

ggplot(overall_success, aes(x=mean_words, y=mean_payout))+geom_point()+geom_smooth(method="lm")+labs(y="Mean payout", x="Average words / round")

```

Using the chat more of the time, and using more words are both associated with higher payoffs.

We can break this down by trial type.

```{r}
success <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId, gametype) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(success, aes(x=mean_chat, y=mean_payout, color=gametype))+geom_point()+geom_smooth(method="lm")+labs(y="Mean Payout", x="Proportion round with nonzero chat")

ggplot(success, aes(x=mean_words, y=mean_payout, color=gametype))+geom_point()+geom_smooth(method="lm")+labs(y="Mean payout", x="Average words / round")

```

Chat helps a lot with BoS and spikeBos, maybe helps a little with hardPD, does not help with easyPD.

We can go even more granular and look at whether words on a trial is associated with more bonus * on that trial *.

```{r}
success_round <- rounds %>% filter(chat_cond=="chat") %>% select(playerId, gameId, gametype, payoff, repNum) %>% inner_join(chat_by_pair) %>% mutate(is.chat=ifelse(words>0, 1,0)) %>% group_by(gameId, gametype, repNum) %>% summarize(mean_payout=mean(payoff), mean_chat=mean(is.chat), mean_words=mean(words))

ggplot(success_round, aes(x=as.factor(mean_chat), y=mean_payout, color=gametype))+geom_jitter(alpha=.1, width=.2, height=0)+stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=0), size=.5)+labs(x="Was there chat this round", y="Mean payout this round")

ggplot(success_round, aes(x=mean_words, y=mean_payout, color=gametype))+geom_jitter(alpha=.1)+geom_smooth(method="lm")+labs(x="Words said this round", y="mean payout this round")
```

There are some real outlier rounds where people say a lot, but mostly they say not that much even when they do say things. 
Talking is better than not talking for most trial types, but there's not much of a dose-response pattern.

# Chat and choices

Rather than collapsing over outcome (which has some randomness), we can look at per-round, which quadrant in the game people chose.

A number of people in the "chat" condition didn't use chat, so we sometimes 3 way split for used chat / didn't use chat / couldn't chat (in contrast to just spliting by intent to treat).

```{r}
outcome <- rounds %>% select(-playerId, -time,-payoff, -box_A, -box_B, -AA, -AB, -BA, -BB, -mix) %>%  
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))  %>% left_join(chat_by_pair) #%>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(gametype, chat_cond, repNum,outcome) %>% tally() %>% group_by(gametype, chat_cond, repNum) %>% summarize(pct=n/sum(n), outcome=outcome)

```

## normal BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both.

```{r}
# outcome %>% filter(gametype=="BoS") %>%
#   mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
#   ggplot(aes(x=repNum, y=outcome_parity))+geom_jitter(width=0, height=.05)+facet_grid(.~chat_cond)+geom_smooth(method="lm")+
#   labs(title="BoS:fraction choosing AA or BB (good) outcomes", y="Good outcome?")

outcome %>% filter(gametype=="BoS") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="BoS:fraction choosing good outcomes",y="Good outcome?")
```

Near chance if you don't talk, far above chance if you do.

## spike BoS

In spike BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are bad for both. (same distribution for spikeBoS, just more skewed).

```{r}
outcome %>% filter(gametype=="spikeBoS") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA", "BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="spike BoS:fraction choosing good outcomes", y="Good outcome?")
```

Actually talking is conducive to good outcomes.

## easyPD

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

```{r}
outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="easyPD:fraction choosing good outcomes", y="Good outcome?")

outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AB","BA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="easyPD:fraction choosing AB or BA outcomes")

outcome %>% filter(gametype=="easyPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="easyPD:fraction choosing BB outcomes")
```
Everyone is likely to choose the good outcome, but those who talk are more reliable at it (especially early on). No real consistency in what is chosen otherwise.

## hardPD

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

```{r}
outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="hardPD:fraction choosing AA outcomes")

outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("AB","BA"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.5))+
  labs(title="hardPD:fraction choosing AB or BA outcomes")

outcome %>% filter(gametype=="hardPD") %>%
  mutate(outcome_parity=ifelse(outcome %in% c("BB"),1, 0)) %>% 
  mutate(talked=ifelse(chat_cond=="chat"&words>0,1,0)) %>% 
  ggplot(aes(x=repNum, y=outcome_parity, color=as.factor(talked)))+geom_jitter(width=1, height=.05)+facet_grid(.~str_c(chat_cond,talked))+geom_smooth(method="lm")+geom_hline(aes(yintercept=.25))+
  labs(title="hardPD:fraction choosing BB outcomes")

```
When chat isn't used, AA increases over the course of the game. When chat is used, AB/BA increases. These are mostly trading off with each other, the obviously bad BB option isn't selected much ever.

# When do people talk?

```{r}

game_chat <- chat %>% filter(chat_cond=="chat") %>% filter(!is.na(targets)) %>% select(-type) %>% left_join(rounds) %>% 
  mutate(words=str_count(text, "\\w+") %|% int(0)) %>%  group_by(gameId, gametype, repNum) %>% summarize(words=sum(words))

ggplot(game_chat, aes(x=repNum,y=words, color=gametype))+geom_jitter()

ggplot(game_chat, aes(x=reorder(gameId, words), y=words, color=gametype))+stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.4), size=.5)+theme(axis.text.x=element_blank())

game_chat %>% mutate(is.chat=ifelse(words>0,1,0)) %>% group_by(gameId, gametype) %>% summarize(pct_chat=mean(is.chat)) %>% ggplot(aes(x=reorder(gameId, pct_chat), y=pct_chat, color=gametype))+geom_jitter()+theme(axis.text.x=element_blank())
```
Seems to be substantial group to group variability in if they talk never/ sometimes/ always. Isn't fully driven by trial type.

How do we increase the rates of talkers!!??

# Does doing well in one type mean doing well in another??

For each game, look at fraction of each trial type where there did which quadrant and correlate??

```{r}
outcome <- rounds %>% select(-playerId, -time,-payoff, -box_A, -box_B, -AA, -AB, -BA, -BB, -mix) %>%  
  pivot_wider(names_from=role, values_from=response) %>% 
  mutate(outcome=str_c(p1,p2))  %>% left_join(chat_by_pair) #%>% mutate(is.chat=ifelse(words>0, 1,0))

outcome_summary <- outcome %>% group_by(gametype, chat_cond, gameId, outcome) %>% tally() %>% group_by(gametype, chat_cond, gameId) %>% summarize(pct=n/sum(n), outcome=outcome) %>% pivot_wider(names_from=c(gametype, outcome), values_from=pct, values_fill=0) %>% 
  mutate(good_BoS=BoS_AA+BoS_BB,
         good_spikeBoS=spikeBoS_AA+spikeBoS_BB,
         uneven_easyPD=easyPD_AB+easyPD_BA,
         uneven_hardPD=hardPD_AB+hardPD_BA)

library(GGally)

outcome_summary %>% filter(chat_cond=="chat")%>% ungroup() %>%  select(good_BoS, good_spikeBoS, uneven_easyPD, uneven_hardPD) %>% ggpairs()

outcome_summary %>% filter(chat_cond=="nochat")%>% ungroup() %>%  select(good_BoS, good_spikeBoS, uneven_easyPD, uneven_hardPD) %>% ggpairs()

```

# What do they say strategy is?

```{r}

exit %>% select(chat_cond, strategy) %>% arrange(chat_cond) %>% kable()
```

### 2024-05-15 experiments/client/main.js
modify for new expt

@@ -3,8 +3,6 @@ import { render } from "react-dom";
 import "@blueprintjs/icons/lib/css/blueprint-icons.css";
 import "@blueprintjs/core/lib/css/blueprint.css";
 
-
-
 import Empirica from "meteor/empirica:core";
 
 import Consent from "./intro/Consent.jsx";
@@ -22,20 +20,19 @@ import ExitSurvey from "./exit/ExitSurvey";
 
 import customBreadcrumb from "./game/Breadcrumb.jsx";
 
- Empirica.header(() => null);
+Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
-Empirica.consent(Consent);
+//Empirica.consent(Consent);
 
 // Introduction pages to show before they play the game (optional).
 // At this point they have been assigned a treatment. You can return
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
   const steps = [Payment, Overview];
-  
 
-  return steps;
-  //return [];
+  //return steps;
+  return [];
 });
 
 // The Round component containing the game UI logic.
@@ -53,7 +50,7 @@ Empirica.round(Round);
 // exit screen will be shown.
 Empirica.exitSteps((game, player) => {
   if (player.exitStatus !== "finished") {
-    return [Sorry,ExitSurvey,Thanks];
+    return [Sorry, ExitSurvey, Thanks];
   } else {
     return [ExitSurvey, Thanks];
   }

### 2024-05-15 experiments/server/callbacks.js

@@ -1,5 +1,11 @@
 import Empirica from "meteor/empirica:core";
-import {names, avatarNames, nameColors, targets, payoffs} from './constants.js';
+import {
+  names,
+  avatarNames,
+  nameColors,
+  targets,
+  payoffs,
+} from "./constants.js";
 import _, { pick } from "lodash";
 
 // //// Avatar stuff //////
@@ -11,10 +17,11 @@ Empirica.onGameStart((game) => {
   const players = game.players;
   console.debug("game ", game._id, " started");
 
-  const roleList = game.get('roleList');
+  const roleList = game.get("roleList");
   //const targets = game.get('context');
 
   players.forEach((player, i) => {
+    player.set("bonus_log", []);
     player.set("role", roleList[player._id]);
     player.set("name", names[i]);
     player.set("avatar", `/avatars/jdenticon/${avatarNames[i]}`);
@@ -27,19 +34,19 @@ Empirica.onGameStart((game) => {
 // It receives the same options as onGameStart, and the round that is starting.
 Empirica.onRoundStart((game, round) => {
   const players = game.players;
-  round.set("chat", []); 
+  round.set("chat", []);
   //round.set("countCorrect",0);
   //round.set('speaker', "")
-  round.set('submitted', false);
-  players.forEach(player => {
+  round.set("submitted", false);
+  players.forEach((player) => {
     //player.set('role', player.get('roleList')[round.index])
     // if (player.get('role')=="speaker"){
     //   round.set('speaker', player._id)
     // }
     player.set("targets", _.shuffle(round.get("targets")));
 
-    player.set('clicked', false);
-    player.set("scoreIncrement",0)
+    player.set("clicked", false);
+    player.set("scoreIncrement", 0);
   });
   //console.log(round)
   //console.log(players)
@@ -62,37 +69,67 @@ Empirica.onStageStart((game, round, stage) => {
 // onStageEnd is triggered after each stage.
 // It receives the same options as onRoundEnd, and the stage that just ended.
 Empirica.onStageEnd((game, round, stage) => {
-  if (stage.name=="selection"){
-    const scale=game.treatment.scale
+  if (stage.name == "selection") {
+    const scale = game.treatment.scale;
     const players = game.players;
-    const p1 = _.find(game.players, p => p.get('role') === "p1");
-    const p2 = _.find(game.players, p => p.get('role') === "p2");
+    const p1 = _.find(game.players, (p) => p.get("role") === "p1");
[…]

### 2024-05-15 experiments/server/main.js

@@ -1,145 +1,148 @@
 import Empirica from "meteor/empirica:core";
 
 import "./callbacks.js";
-import { payoffs, targets} from "./constants";
+import { payoffs, targets } from "./constants";
 import _ from "lodash";
-
 
 function createRoles(players) {
   const l = _.shuffle(players);
-  const roles=_.zipObject(l,["p1","p2"]);
+  const roles = _.zipObject(l, ["p1", "p2"]);
   console.log(roles);
   return roles;
 }
 
-function compareNumbers(a, b) //deal with the fact that js defaults to lexicographic sorting....
-{
-    return a - b;
+function compareNumbers(a, b) {
+  //deal with the fact that js defaults to lexicographic sorting....
+  return a - b;
 }
 
-function chooseTargets(targets){
+function chooseTargets(targets) {
   //takes a list of flower objects, a list of utilties and a number to select
   //returns the context
   // if blinded, then each flower is labelled with which player won't its utils
-  const f=_.slice(_.shuffle(targets),0,2)
-  const labs=["A","B"]
-  const values=_.zipWith(f,labs, (a,b)=>_.assign({"image":a,"label":b}))
-  return values
+  const f = _.slice(_.shuffle(targets), 0, 2);
+  const labs = ["A", "B"];
+  const values = _.zipWith(f, labs, (a, b) => _.assign({ image: a, label: b }));
+  return values;
 }
 
-function createPDRewards(){
+function createPDRewards() {
   //sucker payoff fixed at 0
-  // other rewards on 1-9, but have to respect ordering win > coop > defect 
-  var a=_.slice(_.shuffle(_.range(1,10)),0,3).sort()
-  var coop=a[1]
-  var defect=a[0]
-  var win=a[2]
-  var payoffs={
-   "AA":{p1:coop,p2:coop},
-   "AB":{p1:0,p2:win},
-   "BA":{p1:win,p2:0},
-   "BB":{p1:defect,p2:defect}
-  }
-  return payoffs
+  // other rewards on 1-9, but have to respect ordering win > coop > defect
+  var a = _.slice(_.shuffle(_.range(1, 10)), 0, 3).sort();
+  var coop = a[1];
+  var defect = a[0];
+  var win = a[2];
+  var payoffs = {
+    AA: { p1: coop, p2: coop },
+    AB: { p1: 0, p2: win },
+    BA: { p1: win, p2: 0 },
[…]

### 2024-05-15 experiments/client/main.js
prep expt

@@ -8,6 +8,7 @@ import Empirica from "meteor/empirica:core";
 import Consent from "./intro/Consent.jsx";
 import Payment from "./intro/Payment.jsx";
 import Overview from "./intro/Overview.jsx";
+import Bonus from "./intro/Bonus.jsx";
 import SocialInteractionDetails from "./intro/SocialInteractionDetails.jsx";
 import MoreAboutBonus from "./intro/MoreAboutBonus.jsx";
 import UIOverview from "./intro/UIOverview.jsx";
@@ -23,16 +24,16 @@ import customBreadcrumb from "./game/Breadcrumb.jsx";
 Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
-//Empirica.consent(Consent);
+Empirica.consent(Consent);
 
 // Introduction pages to show before they play the game (optional).
 // At this point they have been assigned a treatment. You can return
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
-  const steps = [Payment, Overview];
+  const steps = [Payment, Overview, Bonus];
 
-  //return steps;
-  return [];
+  return steps;
+  //return [];
 });
 
 // The Round component containing the game UI logic.

### 2024-05-17 code/early_expts_summary.Rmd
playing with models on old data

---
title: "Summary of 2022 experiments"
output:
  html_document: 
    toc: true
---

# Conditions
All experiments did both language and no language versions as a between groups manipulation.

[Expt 1](https://osf.io/8fnze): all PD v all BoS (between subjects)
 - PD from sampling 3 values 1-9, 0 for lowest (sucker payoff)
 - BoS: off diagonal payoff of 1, others from 2-9

[Expt 2](https://osf.io/5au2r): mix of PD and BoS 
 - PD from sampling 3 values 1-9, 0 for lowest (sucker payoff)
 - BoS: off diagonal payoff of 1, others from 2-9

[Expt 3](https://osf.io/c274z): special mix of PD and BoS (this is post bug-fix)

 - 4 "spiked" BoS trials where one of the rewards is high (25-30) and the other is normal (3-7)
 - ~16 normal BoS trials where both rewards are from [1-9]
 - ~10 normal ("easy") PD trials where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] respecting this.
 - ~10 “sacrifice” ("hard") PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff). All rewards are chosen from [1-12] respecting this.

Off diagonal PD rewards for expts 1 and 2 were misdisplayed for player 2. This was resolved for expt 3. 
Note: most PD from expts 1 and 2 will pattern with "easy" PD in expt 3, but also the ones that aren't can't be trusted per error.

# Read and prep

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

## Data import constants
study_1_loc <- "data/study1_220107"
study_1_date_start <- lubridate::ymd("2022-01-07")

study_2_loc <- "data/study2_220413"
study_2_date_start <- lubridate::ymd("2022-04-13")

study_3_loc <- "data/study3_220627"
study_3_date_start <- lubridate::ymd("2022-06-24")
```

Assuming for now that my pre-processing was correct then.

```{r}
# expt 1
one_chat <- read_csv(here(study_1_loc, "raw_chat.csv")) %>%
  mutate(gametype = game_cond) |>
  mutate(cond = "pure_1") |>
  select(gameId, text, type, repNum, playerId, chat_cond, gametype, cond, targets)

one_rounds <- read_csv(here(study_1_loc, "raw_results.csv")) %>%
  mutate(gametype = game_cond) |>
  mutate(cond = "pure_1") |>
  select(gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, cond, role)

# expt 2
two_chat <- read_csv(here(study_2_loc, "raw_chat.csv")) %>%
  mutate(cond = "mix_2") %>%
  select(gameId, text, type, repNum, playerId, chat_cond, gametype, cond, targets)

two_rounds <- read_csv(here(study_2_loc, "raw_results.csv")) %>%
  mutate(cond = "mix_2") %>%
  select(gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, cond, role)

# expt 3
three_chat <- read_csv(here(study_3_loc, "raw_chat.csv")) %>%
  mutate(cond = case_when(
    gametype == "easyPD" ~ "easy_3",
    gametype == "hardPD" ~ "hard_3",
    gametype == "spikeBoS" ~ "spike_3",
    gametype == "BoS" ~ "mix_3"
  )) |>
  mutate(gametype = ifelse(str_detect(gametype, "PD"), "PD", "BoS")) |>
  select(gameId, text, type, repNum, playerId, chat_cond, gametype, cond, targets)

three_rounds <- read_csv(here(study_3_loc, "raw_results.csv")) %>%
  mutate(cond = case_when(
    gametype == "easyPD" ~ "easy_3",
    gametype == "hardPD" ~ "hard_3",
    gametype == "spikeBoS" ~ "spike_3",
    gametype == "BoS" ~ "mix_3"
  )) |>
  mutate(gametype = ifelse(str_detect(gametype, "PD"), "PD", "BoS")) |>
  select(gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, cond, role)

all_chat <- one_chat |>
  bind_rows(two_chat) |>
  bind_rows(three_chat) |>
  filter(!is.na(gametype))
all_rounds <- one_rounds |>
  bind_rows(two_rounds) |>
  bind_rows(three_rounds)
```

# Overall outcomes

Note that points are comparable within type, but not between types. Types are named by game type _ expt number.

## Points earned

```{r}
bonuses <- all_rounds %>%
  group_by(playerId, gameId, chat_cond, cond, gametype) %>%
  summarize(payoff = mean(payoff))

ggplot(bonuses |> filter(gametype == "PD"), aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .5) +
  facet_grid(gametype ~ cond) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )

ggplot(bonuses |> filter(gametype == "BoS"), aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .5) +
  facet_grid(gametype ~ cond) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```

In BoS, earn more points in chat condition than no chat condition.

Minimal difference in PD.

## During chat

```{r}
game_chat <- all_chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(all_rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, cond, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

when_chat <- game_chat %>%
  filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(repNum, gametype, cond) %>%
  summarize(pct_chat = mean(is.chat))

ggplot(when_chat, aes(x = repNum, y = pct_chat, color = cond)) +
  geom_point() +
  facet_grid(~gametype) +
  geom_smooth(se = F, method = "lm") +
  labs(y = "pct of games who said anything each round")
```

In games where chat is an option, it still isn't used that much on a per trial basis. Mixing PD and BoS seems to have helped increase the talking overall.

## Per trial does more talking help

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
ggplot(game_chat |> filter(gametype == "BoS"), aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(game_chat |> filter(gametype == "PD"), aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Looks like using language tends to help, on the trials it's used on, and slightly on the trials it's not used on for BoS. Helps in some cases for PD.

Is there a dose-response relationship, or is one word enough?

```{r}
ggplot(game_chat |> filter(gametype == "BoS", words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(game_chat |> filter(gametype == "PD", words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Especially where we have more data, looks like one word is enough. Indicative of coordination rather than negotiation, probably?

## Per game does talking help?

One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.

So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?

```{r}
group_talk <- game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words), pct_chat = mean(is_chat))

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(gametype == "BoS", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

So, in mixed games, talking on more trials helps regardess of whether you talked on this particular trial. (Although overall talk might also be confounding with conscientiousness..., to fix you'd have to experimentally manipulate when chat is or isn't available)

Talking more might help with hard PD, maybe in mixed?

```{r}
ggplot(overall_talk |> filter(gametype == "BoS", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Here total words is going to mix how much you talk each time with how often you talk, so maybe this is just totally redundant with previous. More talk helps on BoS, even as transfer from other trials. PD seems messier?

# Are they human?

```{r bot}
exit_1 <- read_csv(here(study_1_loc, "exit.csv")) |>
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(game = "one")

exit_2 <- read_csv(here(study_2_loc, "exit.csv")) |>
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(game = "two")

exit_3 <- read_csv(here(study_3_loc, "exit.csv")) |>
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(game = "three", game_cond = "spike_mix")

exit <- exit_1 |>
  bind_rows(exit_2) |>
  bind_rows(exit_3) |>
  select(game, game_cond, chat_cond, pct)

exit
```

Across games, mostly think they are playing with a human, higher in chat than no chat, but not hugely.

# Quadrant analysis per trial

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
outcome <- all_rounds %>%
  select(gameId, repNum, response, chat_cond, gametype, cond, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

## BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
```

Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = repNum, y = outcome_parity, color = cond)) +
  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes", y = "Good outcome?")
```

## PD

Expts 1 and 2 are suspect, but on most trials will pattern with easy PD.

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing AA outcome")

outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AB", "BA"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing AB or BA outcome")

outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing BB outcome")
```

Can get a reasonable option no matter what. If you talk, you can get the uneven, but welfare-maximizing in hard PD.

# Language

How much language?

Filter only for games that talked at least a little.

Second graph filters for *trials* that talked.

```{r}
game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  mutate(expt = str_sub(cond, -1)) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point() +
  facet_wrap(~expt) +
  geom_smooth()

game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  filter(words > 0) |>
  mutate(expt = str_sub(cond, -1)) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point() +
  facet_wrap(~expt) +
  geom_smooth()
```

## how often do *both* people talk

```{r}
pair_chat <- all_chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(all_rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, cond, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  group_by(gameId, gametype, cond, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |>
  group_by(gameId, gametype, cond, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2)) |>
  filter(chat_cond == "chat")

ggplot(pair_chat, aes(x = cond, y = pct_both)) +
  geom_jitter(width = .2) +
  facet_wrap(~gametype)
```
Even in games that talk, there aren't *that* many trials where both people talk?

Probably going to need to manually look at the language for negotiation v "choose red" "yup" etc.

# Models

```{r}
game_chat <- all_chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(all_rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, cond, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  filter(cond %in% c("mix_2", "mix_3", "easy_3"))

never_chat <- game_chat |>
  group_by(gameId, chat_cond) |>
  summarize(words = sum(words)) |>
  mutate(use_chat = ifelse(words > 0, 1, 0))

for_mod <- all_rounds |>
  filter(cond %in% c("mix_2", "mix_3", "easy_3")) |>
  select(gameId, repNum, playerId, payoff, chat_cond, gametype) |>
  left_join(game_chat) |>
  left_join(never_chat) |>
  mutate(
    has_chat = ifelse(chat_cond == "chat", 1, 0),
    is_BoS = ifelse(gametype == "BoS", 1, 0)
  ) # recode with no-chat PD as baseline
```

## Condition assignment on performance

```{r}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

perf_mod <- brm(payoff ~ has_chat * is_BoS + repNum + 
                  (is_BoS | gameId) + 
                  (is_BoS | playerId),
  data = for_mod,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

perf_mod_tot <- brm(payoff ~ (use_chat | has_chat) * is_BoS +
                      repNum +
                      (is_BoS | gameId) + 
                      (is_BoS | playerId),
  data = for_mod, 
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

freq_tot <- lm(payoff ~ (use_chat | has_chat) * is_BoS +
                      repNum , data=for_mod )
```

## Words over time

log(words) ~ game_type * trial_number

```{r}
for_log_words <- for_mod |> filter(words>0) |> mutate(log_words=log(words))

lang_priors  <- c(set_prior("normal(5,5)", class = "Intercept"),
  set_prior("normal(0, 5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor"))

log_words_mod <- brm(log_words ~ is_BoS*repNum +
                       (is_BoS*repNum|gameId),
                     data=for_log_words,
                     prior=lang_priors,
                     control=list(adapt_delta=.95)
                     )

```

## Words on outcome

outcome ~ game_type * log(words)

```{r}
perf_word_mod <- brm(payoff ~ log_words * is_BoS + repNum + 
                  (is_BoS | gameId) + 
                  (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

```

outcome ~ log(words_first_10)*game_type*words_now ?

```{r}
early_words <- for_mod |> filter(repNum<10) |> group_by(gameId) |> summarize(early_words=sum(words))

for_early_word_mod <- for_mod |> filter(repNum>9) |> left_join(early_words) |> filter(early_words>0) |> mutate(log_early_words=log(early_words))

early_word_mod <- brm(payoff ~ log_early_words * is_BoS + is_chat*is_BoS+ repNum + 
                  (is_BoS | gameId) + 
                  (is_BoS | playerId),
  data = for_early_word_mod,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

```

### 2024-05-17 code/early_expts_summary.Rmd
poking at models

@@
 
+```{r}
+
+ggplot(for_log_words, aes(x=log_words, y=payoff, color=repNum))+geom_jitter(alpha=.5)+geom_smooth(method="lm")+facet_wrap(is_BoS~cond)
+```
+
+TODO think more about!!
 outcome ~ log(words_first_10)*game_type*words_now ?
@@
 
-early_word_mod <- brm(payoff ~ log_early_words * is_BoS + is_chat*is_BoS+ repNum + 
+early_word_mod <- brm(payoff ~ log_early_words * is_BoS* is_chat+ repNum + 
                   (is_BoS | gameId) + 

### 2024-05-20 code/analysis_may2024.Rmd
data and prelim analysis

---
title: "Mix of PD and BoS"
output:
  html_document: 
    toc: true
---

```{r set-up, include=F}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"), fig.width=7, fig.height=3)
options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

## Data import constants
data_location <- "data/new_study1_240520"

date_start <- lubridate::ymd("2024-05-20")

image_location <- "write-ups/images"

model_location <- "code/models"
```

```{r bonus}
## This is for determining participant bonuses using the version of data with PID

# d.players <- read_csv(here(data_location, 'players.csv')) %>%
#   rename(playerId=`_id`) %>%
#   select(data.bonus, playerId,id) %>%
#   mutate(bonus=round(data.bonus,2),
#          bonus=bonus %|% 0,
#          cost=round(bonus*4/3,2)) %>%
#   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
#   filter(bonus>0) %>% write_csv(here(data_location,"for_prolific.csv"))
# #
# cost <- read_csv(here(data_location, "player_payments.csv"))
# #
# sum(cost$cost)
```

```{r, include=F, eval=F}
d.treatments <- read_csv(here(data_location, "treatments.csv")) %>%
  rename(treatmentId = `_id`) %>%
  select(treatmentId, name) |>
  mutate(
    chat_cond = ifelse(str_detect(name, "no_chat"), "no_chat", "chat"),
    game_cond = "may2024"
  )

d.games <- read_csv(here(data_location, "games.csv")) %>%
  rename(gameId = `_id`) %>%
  left_join(d.treatments) %>%
  filter(createdAt >= date_start) %>%
  select(gameId, playerIds, game_cond, chat_cond)

d.gameplayers <- d.games %>%
  separate(playerIds, c("p1", "p2")) %>%
  pivot_longer(p1:p2, values_to = "playerId") %>%
  select(-name)

d.players <- read_csv(here(data_location, "players.csv")) %>%
  select(playerId = `_id`, role = data.role, data.bonus) %>%
  left_join(d.gameplayers)
```

```{r, include=F, eval=F}
d.chat.raw <- read_csv(here(data_location, "rounds.csv"), guess_max = 10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), "{}", data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>%
  select(-ends_with("response"), -ends_with("time"), -ends_with("payoff"), -ends_with("role")) %>%
  rename(data.gametype = data.type) %>% # VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  left_join(d.games) %>%
  write_csv(here(data_location, "raw_chat.csv"))

d.round_results.raw <- read_csv(here(data_location, "rounds.csv"), guess_max = 10000) %>%
  filter(createdAt >= date_start) %>%
  rename(data.gametype = data.type) %>% # VB 220413:this fixes a collision with the "type" column
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  rename_with(~ gsub("room", "player", .x, fixed = T)) %>%
  rename_with(~ gsub("player", "player_", .x, fixed = T)) %>%
  rename_with(~ gsub("response", "_response", .x, fixed = T)) %>%
  rename_with(~ gsub("time", "_time", .x, fixed = T)) %>%
  select(-chat) %>%
  gather(key, value, starts_with("player")) %>%
  separate(key, into = c("blah", "playerId", "info")) %>%
  spread(info, value) %>%
  select(-blah) %>%
  select(-stageIds, -index, -createdAt, -submitted) %>%
  filter(!is.na(response)) %>%
  mutate(time = as.numeric(time) / 1000) %>%
  mutate(targets = map(targets, .f = ParseJSONColumn)) %>%
  unnest(targets) %>%
  mutate(
    image = gsub("/experiment/", "", image, fixed = TRUE),
    image = gsub(".jpeg", "", image, fixed = TRUE)
  ) %>%
  pivot_wider(names_from = label, names_prefix = "box_", values_from = image) %>%
  mutate(payoffs = map(payoffs, .f = ParseJSONColumn)) %>%
  unnest(payoffs) %>%
  mutate(
    AA = ifelse(role == "p1", AA.p1, AA.p2),
    AB = ifelse(role == "p1", AB.p1, AB.p2),
    BA = ifelse(role == "p1", BA.p1, BA.p2),
    BB = ifelse(role == "p1", BB.p1, BB.p2),
    type = ifelse(AB == 0 & BA == 0, "BoS", "PD")
  ) %>%
  left_join(d.games)

d.exit.survey <- read_csv(here(data_location, "player-inputs.csv")) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c("gameId")) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.complete <- d.round_results.raw %>%
  group_by(gameId, playerId) %>%
  summarize(max_rep = max(repNum)) %>%
  filter(max_rep == 39) %>%
  select(gameId, playerId) %>%
  write_csv(here(data_location, "complete.csv"))

d.chat <- d.chat.raw %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "raw_chat.csv"))
d.round_results <- d.round_results.raw %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "raw_results.csv"))
d.exit <- d.exit.survey %>%
  inner_join(d.complete) %>%
  write_csv(here(data_location, "exit.csv"))
```

# Are they human?

```{r bot}
exit <- read_csv(here(data_location, "exit.csv"))

exit %>%
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes))
```

Yay, they mostly think they're playing with a human!!

Looks like we have 32 games in each condition (from less attrition than I anticipated).

```{r}
chat <- read_csv(here(data_location, "raw_chat.csv")) %>%
  select(gameId, text, type, repNum, playerId, chat_cond, gametype, targets)

rounds <- read_csv(here(data_location, "raw_results.csv")) %>%
  select(gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, role)
```

# Before chat

```{r}
games <- rounds |>
  select(gameId, chat_cond) |>
  unique()

before_chat <- chat |>
  filter(is.na(targets)) |>
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, chat_cond) |>
  summarize(words = sum(words)) |>
  full_join(games)

ggplot(before_chat, aes(x = chat_cond, y = words)) +
  geom_jitter(width = .1, color = "blue", alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")
```

People do talk to each other a little in the pre-game chat time.

# Points earned

Note that BoS has a lower points range than PD because of what range of random numbers is selected.

```{r}
bonuses <- rounds %>%
  group_by(playerId, gameId, chat_cond, gametype) %>%
  summarize(payoff = mean(payoff))

ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(chat_cond))) +
  geom_jitter(width = .1, alpha = .1) +
  facet_grid(. ~ gametype) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```
Not seeing much of a difference in either category...

# During chat

```{r}
game_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

when_chat <- game_chat %>%
  filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(repNum, gametype) %>%
  summarize(pct_chat = mean(is.chat))

ggplot(when_chat, aes(x = repNum, y = pct_chat, color = gametype)) +
  geom_point() +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim=c(0,1))+
  labs(title = "pct of games who said anything each round")

game_chat |> filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId, gametype) %>%
  summarize(pct_chat = mean(is.chat)) |> 
  ggplot(aes(x=reorder(gameId, pct_chat, mean), y=pct_chat, color=gametype))+geom_point()+
  theme(axis.text.x=element_blank())+
    coord_cartesian(ylim=c(0,1))
```

looks like people are more likely to talk in BoS than PD games. But there's not a lot of talking at all. Driven by group-differences -- some pairs never or almost never chatted even if they could -- others used chat a substantial fraction of the time.

## Per trial does more talking help

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
ggplot(game_chat, aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

Looks like using language is correlated with better outcome with BoS and slightly better outcome for PD. But if this language helping or "people who avail themselves of the option to use language are more competent".

Is there a dose-response relationship, or is one word enough?

```{r}
ggplot(game_chat |> filter(words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(. ~ gametype) +
  geom_smooth(method = "lm")
```

Especially where we have more data, looks like one word is enough. Indicative of coordination rather than negotiation, probably?

## Per game does talking help?

One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.

So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?

```{r}
group_talk <- game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words), pct_chat = mean(is_chat))

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  geom_jitter(alpha = .1) +
  geom_smooth(method = "lm") +
  facet_grid(. ~ gametype) +
  theme(legend.position = "bottom")
```

Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? Maaaybe? But there's a real data-sparsity issue.

# Option selected

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
outcome <- rounds %>%
  select(gameId, repNum, response, chat_cond, gametype, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

## BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
```

Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = repNum, y = outcome_parity)) +
  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
  stat_summary(fun.data="mean_cl_boot", geom="point")+
  geom_smooth(method = "lm", se = F) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes", y = "Good outcome?")
```

## PD

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) 
```

Can get a reasonable option no matter what. 
```{r}
outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = repNum, y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
  facet_grid(.~str_c(chat_cond, "_", is_chat))+
  geom_smooth(method="lm")+
  geom_hline(aes(yintercept = .5)) 
```

# Language

How much language?

Filter only for games that talked at least a little.

Second graph filters for *trials* that talked.

```{r}
game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point(alpha = .1) +
  geom_smooth()

game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  filter(words > 0) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point(alpha = .1) +
  geom_smooth()
```

## how often do *both* people talk

```{r}
pair_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  group_by(gameId, gametype, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |>
  group_by(gameId, gametype, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2)) |>
  filter(chat_cond == "chat")

ggplot(pair_chat, aes(x = gametype, y = pct_both)) +
  geom_jitter(width = .2) +
  stat_summary(fun.data = "mean_cl_boot", color = "red")
```
Even in games that talk, there aren't *that* many trials where both people talk?

# Models

```{r}
game_chat <- chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

never_chat <- game_chat |>
  group_by(gameId, chat_cond) |>
  summarize(words = sum(words)) |>
  mutate(use_chat = ifelse(words > 0, 1, 0))

for_mod <- rounds |>
  select(gameId, repNum, playerId, payoff, chat_cond, gametype) |>
  left_join(game_chat) |>
  left_join(never_chat) |>
  mutate(
    has_chat = ifelse(chat_cond == "chat", 1, 0),
    is_BoS = ifelse(gametype == "BoS", 1, 0)
  ) # recode with no-chat PD as baseline
```

## Condition assignment on performance

```{r}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

perf_mod <- brm(
  payoff ~ has_chat * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_mod,
  prior = perf_priors,
  file = here(model_location, "perf_expt1"),
  control = list(adapt_delta = .95)
)

summary(perf_mod)
```

Note the high group level variability...

## Words over time

log(words) ~ game_type * trial_number

```{r}
for_log_words <- for_mod |>
  filter(words > 0) |>
  mutate(log_words = log(words))

lang_priors <- c(
  set_prior("normal(5,5)", class = "Intercept"),
  set_prior("normal(0, 5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

log_words_mod <- brm(
  log_words ~ is_BoS * repNum +
    (is_BoS * repNum | gameId),
  data = for_log_words,
  file = here(model_location, "words_expt1"),
  prior = lang_priors,
  control = list(adapt_delta = .95)
)
summary(log_words_mod)
```

More talk in BoS

## Words on outcome

outcome ~ game_type * log(words)

```{r}
perf_word_mod <- brm(
  payoff ~ log_words * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  file = here(model_location, "perf_words_expt1"),
  control = list(adapt_delta = .99)
)

summary(perf_word_mod)
```

Again, big group level variation.

# TODOs

* Look at the actual language

* Look at what people said their strategies were

### 2024-10-07 code/analysis_may2024.Rmd
clean up

@@
 ---
+
+# Summary
+
+* 32 games each condition (chat, no-chat)
+* humanness ratings are okay (75-77% -- in the acceptable range, probably lower than ideal)
+* seeing very little effect of chat, probably because people don't use the chat very much (and group to group variation) 
+* unlike prior pilots, not seeing an effect of chatted other times but not this one
+  * could be chance / slightly different participant populations
+  * could be different strategy space -- we did take away the points-counter which may make it harder to use that as an agreement point
+  * could be this task is too easy
+* about half of all utterances contain a color word 
+* strategy wise, a lot of talk about fairness, also some about having an idea of a strategy but partner not getting it
+
+so...what now:
+* could give up on this project 
+* one big issue for both design and interpretation is that people's incentives are hard to control or predict and aren't the numbers we put out (strong values of fairness, for instance)
+  * the idea was to try to leverage game theory predictions, but people don't act like game
+  theory agents / the payoffs that a person is choosing over aren't what we're displaying
+  * possibly other areas of game theory would be better? 
+* could force chat usage (someone must send message before selection can occur -- heavy handed, but could get us ceiling) 
+* could up the difficulty -- more game types mixed together
+* could switch back to offering running total of reward
 
@@
 ```{r bonus}
+
 ## This is for determining participant bonuses using the version of data with PID
@@
 # #
-# cost <- read_csv(here(data_location, "player_payments.csv"))
+#cost <- read_csv(here("player_payments.csv")) |> select(id, bonus) |> filter(bonus>0) |> write_csv(here("for_prolific.csv"))
 # #
@@
 
-# TODOs
+# What people say strategy is
 
-* Look at the actual language
+```{r}
+exit <- read_csv(here(data_location, "exit.csv")) |> select(chat_cond, strategy) |> filter(!is.na(strategy))
 
-* Look at what people said their strategies were
+message("chat")
+kable(exit |> filter(chat_cond=="chat") |> select(-chat_cond))
+
+message("no-chat")
+kable(exit |> filter(chat_cond=="no_chat") |> select(-chat_cond))
+
+```
+
+# In game talk
+
+note that there's some non-game-relevant chit-chat that hasn't been cleaned up!
+
+```{r}
+color_words <- c("red","blue", "purple", "green","yellow", "brown", "pink", "violet","cyan", "grey", "peach", "purplish", "beige", "orange")
+agreement_words <- c("okay", "ok", "yes", "sure")
+color_regex <- str_c(color_words, collapse="|")
+agreement_regex <- str_c(agreement_words, collapse="|")
+
+in_game_said <- chat %>%
+  filter(!is.na(targets)) %>%
+  select(-type) %>%
+  full_join(rounds) |> filter(!is.na(text)) |>  mutate(gameId=str_sub(gameId, 1,3), playerId=str_sub(playerId, 1,3)) |> select(gameId, playerId, repNum, gametype, text) |> 
+  mutate(has_color=str_detect(str_to_lower(text), color_regex),
+         has_agree=str_detect(str_to_lower(text), agreement_regex))
+
+in_game_said |> group_by(has_color) |> tally()
+in_game_said |> group_by(has_agree) |> tally()
+
+kable(in_game_said |> select(-has_color, -has_agree))
+```

### 2024-10-07 code/early_expts_summary.Rmd
clean up

@@
   labs(y = "pct of games who said anything each round")
+```
+```{r}
+
+ggplot(when_chat |> filter(cond %in%c("mix_2", "mix_3", "easy_3")) |> mutate(condition=ifelse(str_detect(cond, "3"),3,2)), aes(x = repNum, y = pct_chat, color = gametype)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  coord_cartesian(ylim=c(0,1))+
+  labs(title = "pct of games who said anything each round")+facet_wrap(~condition)
+
+game_chat |> filter(chat_cond == "chat") |>filter(cond %in%c("mix_2", "mix_3", "easy_3")) |> 
+  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
+  group_by(gameId, gametype, cond) %>%
+  summarize(pct_chat = mean(is.chat)) |> 
+  mutate(condition=ifelse(str_detect(cond, "3"),3,2)) |> 
+  ggplot(aes(x=reorder(gameId, pct_chat, mean), y=pct_chat, color=gametype))+geom_point()+
+  theme(axis.text.x=element_blank())+
+    coord_cartesian(ylim=c(0,1))+facet_wrap(~condition)
+
 ```

### 2024-10-07 experiments/client/main.js
a believed to be expt 3 + updated turing test but
Veronica should check this!

@@ -3,12 +3,13 @@ import { render } from "react-dom";
 import "@blueprintjs/icons/lib/css/blueprint-icons.css";
 import "@blueprintjs/core/lib/css/blueprint.css";
 
+
+
 import Empirica from "meteor/empirica:core";
 
 import Consent from "./intro/Consent.jsx";
 import Payment from "./intro/Payment.jsx";
 import Overview from "./intro/Overview.jsx";
-import Bonus from "./intro/Bonus.jsx";
 import SocialInteractionDetails from "./intro/SocialInteractionDetails.jsx";
 import MoreAboutBonus from "./intro/MoreAboutBonus.jsx";
 import UIOverview from "./intro/UIOverview.jsx";
@@ -21,7 +22,7 @@ import ExitSurvey from "./exit/ExitSurvey";
 
 import customBreadcrumb from "./game/Breadcrumb.jsx";
 
-Empirica.header(() => null);
+ Empirica.header(() => null);
 
 // Set the Consent Component you want to present players (optional).
 Empirica.consent(Consent);
@@ -30,7 +31,8 @@ Empirica.consent(Consent);
 // At this point they have been assigned a treatment. You can return
 // different instruction steps depending on the assigned treatment.
 Empirica.introSteps((game, treatment) => {
-  const steps = [Payment, Overview, Bonus];
+  const steps = [Payment, Overview];
+  
 
   return steps;
   //return [];
@@ -51,7 +53,7 @@ Empirica.round(Round);
 // exit screen will be shown.
 Empirica.exitSteps((game, player) => {
   if (player.exitStatus !== "finished") {
-    return [Sorry, ExitSurvey, Thanks];
+    return [Sorry,ExitSurvey,Thanks];
   } else {
     return [ExitSurvey, Thanks];
   }

### 2024-10-07 experiments/server/callbacks.js

@@ -83,36 +83,10 @@ Empirica.onStageEnd((game, round, stage) => {
       p1.set("scoreIncrement", payout.p1);
       p2.set("scoreIncrement", payout.p2);
       //Save outcomes as property of round for later export/analysis
-      const bonus_trials = [7, 15, 32, 38];
-      //const bonus_trials = [0, 1, 2, 3];
-      if (bonus_trials.includes(round.get("repNum"))) {
-        console.log(round.get("repNum"));
-        let bonus_p1 = {
-          rep: round.get("repNum"),
-          own: payout.p1,
-          partner: payout.p2,
-        };
-        let bonus_p2 = {
-          rep: round.get("repNum"),
-          own: payout.p2,
-          partner: payout.p1,
-        };
-        let p2_old_bonus = p2.get("bonus_log");
-        let p1_old_bonus = p1.get("bonus_log");
-        p2_old_bonus.push(bonus_p2);
-        p1_old_bonus.push(bonus_p1);
-        p2.set("bonus_log", p2_old_bonus);
-        p1.set("bonus_log", p1_old_bonus);
-        players.forEach((player) => {
-          const currScore = player.get("bonus");
-          const scoreIncrement = player.get("scoreIncrement");
-          player.set("bonus", scoreIncrement * 0.1 + currScore);
-        });
-      }
       players.forEach((player) => {
-        //const currScore=player.get("bonus")
+        const currScore = player.get("bonus");
         const scoreIncrement = player.get("scoreIncrement");
-        //player.set("bonus", scoreIncrement*.01*scale + currScore);
+        player.set("bonus", scoreIncrement * 0.01 * scale + currScore);
         round.set("player_" + player._id + "_response", player.get("clicked"));
         round.set(
           "player_" + player._id + "_time",

### 2024-10-07 experiments/server/main.js

@@ -1,148 +1,145 @@
 import Empirica from "meteor/empirica:core";
 
 import "./callbacks.js";
-import { payoffs, targets } from "./constants";
+import { payoffs, targets} from "./constants";
 import _ from "lodash";
+
 
 function createRoles(players) {
   const l = _.shuffle(players);
-  const roles = _.zipObject(l, ["p1", "p2"]);
+  const roles=_.zipObject(l,["p1","p2"]);
   console.log(roles);
   return roles;
 }
 
-function compareNumbers(a, b) {
-  //deal with the fact that js defaults to lexicographic sorting....
-  return a - b;
+function compareNumbers(a, b) //deal with the fact that js defaults to lexicographic sorting....
+{
+    return a - b;
 }
 
-function chooseTargets(targets) {
+function chooseTargets(targets){
   //takes a list of flower objects, a list of utilties and a number to select
   //returns the context
   // if blinded, then each flower is labelled with which player won't its utils
-  const f = _.slice(_.shuffle(targets), 0, 2);
-  const labs = ["A", "B"];
-  const values = _.zipWith(f, labs, (a, b) => _.assign({ image: a, label: b }));
-  return values;
+  const f=_.slice(_.shuffle(targets),0,2)
+  const labs=["A","B"]
+  const values=_.zipWith(f,labs, (a,b)=>_.assign({"image":a,"label":b}))
+  return values
 }
 
-function createPDRewards() {
+function createPDRewards(){
   //sucker payoff fixed at 0
-  // other rewards on 1-9, but have to respect ordering win > coop > defect
-  var a = _.slice(_.shuffle(_.range(1, 10)), 0, 3).sort();
-  var coop = a[1];
-  var defect = a[0];
-  var win = a[2];
-  var payoffs = {
-    AA: { p1: coop, p2: coop },
-    AB: { p1: 0, p2: win },
-    BA: { p1: win, p2: 0 },
-    BB: { p1: defect, p2: defect },
-  };
-  return payoffs;
-}
-function createEasyPDRewards() {
-  //sucker payoff fixed at 0
-  // other rewards on 1-12, but have to respect ordering win > coop > defect
-  // we redraw until we have a set where 2*coop \ge win
-  var a = _.slice(_.shuffle(_.range(1, 13)), 0, 3).sort(compareNumbers);
[…]

### 2024-10-23 code/analysis_may2024.Rmd
amount of talking trials

@@
     coord_cartesian(ylim=c(0,1))
+
+game_chat |> filter(chat_cond == "chat") |>
+  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
+  group_by(gameId) %>%
+  summarize(pct_chat = mean(is.chat)) |> 
+  arrange(pct_chat) |> 
+  mutate(x=1/n()*row_number()) |> ggplot(aes(x=x, y=pct_chat))+geom_point()+
+    coord_cartesian(ylim=c(0,1))+geom_hline(yintercept=.1)+geom_hline(yintercept=.5)+labs(title="Game 4")
+
 ```

### 2024-10-23 code/early_expts_summary.Rmd
amount of talking trials

@@
   labs(y = "pct of games who said anything each round")
+```
+
+```{r}
+game_chat |> filter(chat_cond == "chat") |>
+  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
+  mutate(game_no=str_sub(cond, -1)) |> 
+  group_by(gameId, game_no) %>%
+  summarize(pct_chat = mean(is.chat)) |> 
+  arrange(game_no, pct_chat) |> 
+  group_by(game_no) |> 
+  mutate(x=1/n()*row_number()) |> ggplot(aes(x=x, y=pct_chat))+geom_point()+facet_wrap(~game_no)+
+    coord_cartesian(ylim=c(0,1))+geom_hline(yintercept=.1)+geom_hline(yintercept=.5)
+
 ```

### 2024-10-29 code/analysis_oct2024.Rmd
add analysis expt 5 prelim

--- 2024-05-20 code/analysis_may2024.Rmd
@@
 ---
-title: "Mix of PD and BoS"
+title: "Expt 5: Mix of PD and BoS with spikes and two types of PD trials"
 output:
@@
 library(rlang)
-library(lme4)
-library(brms)
-library(rstan)
 library(viridis)
 library(knitr)
-
-rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
 theme_set(theme_bw())
@@
 ## Data import constants
-data_location <- "data/new_study1_240520"
+data_location <- "data/study5_241028"
 
-date_start <- lubridate::ymd("2024-05-20")
+date_start <- lubridate::ymd("2024-10-26")
 
@@
 
+Note: one participant wasn't getting a partner, so they opened up a new tab, fiddled with the id and played themselves. 
+So we exclude the game with player id YSB8RYRgF2tQjym2e, which is game AQXAv4FKrZxBrTQgE
+
 ```{r bonus}
+
 ## This is for determining participant bonuses using the version of data with PID
@@
 #          bonus=bonus %|% 0,
-#          cost=round(bonus*4/3,2)) %>%
+#          cost=round(bonus*4/3,2)) 
 #   write_csv(here(data_location, "player_payments.csv")) %>% select(id,bonus) %>%
@@
 # #
-# cost <- read_csv(here(data_location, "player_payments.csv"))
+# cost <- read_csv(here(data_location,"player_payments.csv"))
 # #
@@
 ```
+
+# Summary
+
+Each game has:
+
+ - 4 "spiked" BoS trials where one of the rewards is high (25-30) and the other is normal (3-7)
+
+ - ~16 normal BoS trials where both rewards are from [1-9]
+
+ - ~10 normal ("easy") PD trials where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] respecting this.
+
+ - ~10 “sacrifice” ("hard") PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff).
+All rewards are chosen from [1-12] respecting this.
 
@@
   mutate(
-    chat_cond = ifelse(str_detect(name, "no_chat"), "no_chat", "chat"),
-    game_cond = "may2024"
+    chat_cond = ifelse(str_detect(name, "no chat"), "no_chat", "chat"),
+    game_cond = "oct2024"
   )
@@
   select(gameId, playerId) %>%
+  filter(gameId!="AQXAv4FKrZxBrTQgE") |> # the game where someone was clever and played themselves
   write_csv(here(data_location, "complete.csv"))
@@
 
+53 games in each condition
+
+```{r}
+read_csv(here(data_location,"raw_results.csv")) |> select(chat_cond, gameId) |> unique() |> group_by(chat_cond) |> tally()
+
+```
+
 # Are they human?
@@
 ```
-
-Yay, they mostly think they're playing with a human!!
-
-Looks like we have 32 games in each condition (from less attrition than I anticipated).
+mostly think partner is human, oh good.
 
@@
   labs(y = "mean_reward", x = "condition") +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black", geom="errorbar") +
     stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
@@
 ```
-Not seeing much of a difference in either category...
+
+Looks like chat does better in spike, and a little bit on normal BoS?
 
@@
 
-game_chat |> filter(chat_cond == "chat") |>
+how_many_chat <- game_chat |> filter(chat_cond == "chat") |>
   mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
-  group_by(gameId, gametype) %>%
+  group_by(gameId) %>%
   summarize(pct_chat = mean(is.chat)) |> 
-  ggplot(aes(x=reorder(gameId, pct_chat, mean), y=pct_chat, color=gametype))+geom_point()+
-  theme(axis.text.x=element_blank())+
-    coord_cartesian(ylim=c(0,1))
+  arrange(pct_chat)
+
+#how_many_chat |> nrow()
+#how_many_chat |> filter(pct_chat>=.1) |> nrow()
+
+how_many_chat |> 
+  mutate(x=1/n()*row_number()) |> ggplot(aes(x=x, y=pct_chat))+geom_point()+
+    coord_cartesian(ylim=c(0,1))+geom_hline(yintercept=.1)+geom_hline(yintercept=.5)+labs(title="Game 5")
+
 ```
-
-looks like people are more likely to talk in BoS than PD games.
-But there's not a lot of talking at all.
-Driven by group-differences -- some pairs never or almost never chatted even if they could -- others used chat a substantial fraction of the time.
+Not a lot of talking.
 
@@
 ggplot(game_chat, aes(x = str_c(chat_cond, "_", is_chat), y = mean_payout)) +
+  stat_summary(fun.data = "mean_cl_boot", geom="errorbar") +
     stat_summary(fun.data = "mean_cl_boot") +
+
   facet_grid(. ~ gametype) +
@@
 ```
+
+Looks like mostly the actually using the chat is what's helping.
 
@@
 overall_talk <- game_chat |> left_join(group_talk)
+
 ggplot(overall_talk |> filter(chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
@@
 
-Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? Maaaybe? But there's a real data-sparsity issue.
+Looks like talking on *other* trials might help in BoS if you didn't talk on this specific trial? But might just be fitting to outliers? Will need models.
 
@@
 
-Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.
+Near chance if you can't talk, or if you don't talk, far above chance if you do coordinate.
 
@@
 
-## PD
+## spike BoS
+
+```{r}
+outcome %>%
+  filter(gametype == "spikeBoS") %>%
+  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
+  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) +
+  labs(title = "Spiked BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
+
+outcome %>%
+  filter(gametype == "spikeBoS") %>%
+  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
+  ggplot(aes(x = repNum, y = outcome_parity)) +
+  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
+  stat_summary(fun.data="mean_cl_boot", geom="point")+
+  geom_smooth(method = "lm", se = F) +
+  geom_hline(aes(yintercept = .5)) +
+  labs(title = "spike BoS:fraction choosing good outcomes", y = "Good outcome?")
+```
+
+## easy PD
 
@@
 Can get a reasonable option no matter what.
+
 ```{r}
@@
   filter(gametype == "easyPD") %>%
+  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
+         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
+         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
+  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
+  ggplot(aes(x = repNum, y = selected, color=type)) +
+  stat_summary(fun.data = "mean_cl_boot", geom="point", alpha=.5) +
+  facet_grid(.~str_c(chat_cond, "_", is_chat))+
+  geom_smooth(method="lm")+
+  geom_hline(aes(yintercept = .5)) 
+```
+
+## hard PD
+
+In hard PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AB/BA is welfare maximizing.
+
+```{r}
+outcome %>%
+  filter(gametype == "hardPD") %>%
+  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
+         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
+         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
+  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
+  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = selected, color=type)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) 
+```
+So if you do use the chat, you tend to get the best option. (Using chat means that easy and hard PD look different)
+
+Whereas if you don't chat, they look more similar at least.
+
+```{r}
+outcome %>%
+  filter(gametype == "hardPD") %>%
   mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
@@
   group_by(gameId, gametype, chat_cond) |>
-  summarize(pct_both = mean(how_many_chat == 2)) |>
-  filter(chat_cond == "chat")
+  summarize(pct_both = mean(how_many_chat == 2),
+            pct_either=mean(how_many_chat > 0)) |> 
+  filter(chat_cond == "chat") |> 
+  pivot_longer(pct_both:pct_either)
 
-ggplot(pair_chat, aes(x = gametype, y = pct_both)) +
-  geom_jitter(width = .2) +
-  stat_summary(fun.data = "mean_cl_boot", color = "red")
+ggplot(pair_chat, aes(x = gametype, y = value, color=name, group=name)) +
+  geom_point(position=position_jitterdodge(dodge.width=.7,jitter.width = .2)) +
+  stat_summary(fun.data = "mean_cl_boot", color = "black", position=position_dodge(.7))
 ```
+
 Even in games that talk, there aren't *that* many trials where both people talk?
 
+<!--
 # Models
 
-```{r}
+```{r,eval=F}
 game_chat <- chat %>%
@@
 
-```{r}
+```{r, eval=F}
 perf_priors <- c(
@@
 
-```{r}
+```{r, eval=F}
 for_log_words <- for_mod |>
@@
 
-```{r}
+```{r, eval=F}
 perf_word_mod <- brm(
@@
 
-# TODOs
+# What people say strategy is
 
-* Look at the actual language
+```{r, eval=F}
+exit <- read_csv(here(data_location, "exit.csv")) |> select(chat_cond, strategy) |> filter(!is.na(strategy))
 
-* Look at what people said their strategies were
+message("chat")
+kable(exit |> filter(chat_cond=="chat") |> select(-chat_cond))
+
+message("no-chat")
+kable(exit |> filter(chat_cond=="no_chat") |> select(-chat_cond))
+
+```
+
+# In game talk
+
+note that there's some non-game-relevant chit-chat that hasn't been cleaned up!
+
+```{r, eval=F}
+color_words <- c("red","blue", "purple", "green","yellow", "brown", "pink", "violet","cyan", "grey", "peach", "purplish", "beige", "orange")
+agreement_words <- c("okay", "ok", "yes", "sure")
+color_regex <- str_c(color_words, collapse="|")
+agreement_regex <- str_c(agreement_words, collapse="|")
+
+in_game_said <- chat %>%
+  filter(!is.na(targets)) %>%
+  select(-type) %>%
+  full_join(rounds) |> filter(!is.na(text)) |>  mutate(gameId=str_sub(gameId, 1,3), playerId=str_sub(playerId, 1,3)) |> select(gameId, playerId, repNum, gametype, text) |> 
+  mutate(has_color=str_detect(str_to_lower(text), color_regex),
+         has_agree=str_detect(str_to_lower(text), agreement_regex))
+
+in_game_said |> group_by(has_color) |> tally()
+in_game_said |> group_by(has_agree) |> tally()
+
+kable(in_game_said |> select(-has_color, -has_agree))
+```
+
+-->

### 2025-06-14 code/analysis_oct2024.Rmd
do models for ms

@@
 
+# Subgroup analysis
+
+As an exploratory thing, what if we look at the people who talked *a lot* or *a moderate amount*
+
+high = 40%+ of trials (10 games)
+
+med = 10% - 39% of trials (10 games)
+
+minimal = < 10% of trials (33 games)
+
+no = couldn't chat (53 games)
+
+```{r}
+
+how_many_chat <- game_chat |> filter(chat_cond == "chat") |>
+  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
+  group_by(gameId) %>%
+  summarize(pct_chat = mean(is.chat)) |> 
+  arrange(pct_chat)
+
+chat_subgroup <- how_many_chat |> mutate(chat_subgroup=case_when(
+  pct_chat>.4 ~ "high",
+  pct_chat>.09 ~ "med",
+  T ~ "min"
+))
+
+subgroup_rounds <- rounds |> left_join(chat_subgroup) |>
+  mutate(chat_subgroup=ifelse(is.na(chat_subgroup), "no", chat_subgroup))
+
+subgroup_game_chat <- game_chat |> left_join(chat_subgroup) |>
+  mutate(chat_subgroup=ifelse(is.na(chat_subgroup), "no", chat_subgroup))
+```
+
+```{r}
+bonuses <- subgroup_rounds %>%
+  group_by(playerId, gameId, chat_subgroup, gametype) %>%
+  summarize(payoff = mean(payoff))
+
+ggplot(bonuses, aes(x = chat_subgroup, y = payoff, color = str_c(chat_subgroup))) +
+  geom_jitter(width = .1, alpha = .1) +
+  facet_grid(. ~ gametype) +
+  scale_color_viridis(discrete = T) +
+  labs(y = "mean_reward", x = "chat usage") +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black", geom="errorbar") +
+    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
+  theme(legend.position = "none") +
+  coord_cartesian(ylim = c(0, 20)) +
+  labs(y = "Mean reward / trial") +
+  theme(
+    axis.text = element_text(size = 12),
+    axis.title = element_text(size = 14),
+    strip.text = element_text(size = 14)
+  )
+
+```
+
+```{r}
+ggplot(subgroup_game_chat, aes(x = str_c(chat_subgroup, "_", is_chat), y = mean_payout)) +
+    geom_jitter(aes(color=chat_subgroup), alpha=.2)+
+
+  stat_summary(fun.data = "mean_cl_boot", geom="errorbar") +
+    stat_summary(fun.data = "mean_cl_boot") +
+
+facet_wrap(~gametype) +
+  geom_smooth(method = "lm")+
+  theme(legend.position = "none")
+```
+
+## Option
+
+```{r}
+subgroup_outcome <- subgroup_rounds %>%
+  select(gameId, repNum, response, chat_subgroup, gametype, role) |>
+  pivot_wider(names_from = role, values_from = response) %>%
+  mutate(outcome = str_c(p1, p2)) %>%
+  left_join(game_chat)
+```
+
+### BoS
+```{r}
+subgroup_outcome %>%
+  filter(gametype == "BoS") %>%
+  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
+  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = outcome_parity, color=chat_subgroup)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) +
+  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
+```
+
+### Spike BoS
+
+```{r}
+subgroup_outcome %>%
+  filter(gametype == "spikeBoS") %>%
+  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
+  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = outcome_parity, color=chat_subgroup)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) +
+  labs(title = "Spiked BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
+
+```
+
+### easy PD
+
+In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.
+
+```{r}
+subgroup_outcome %>%
+  filter(gametype == "easyPD") %>%
+  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
+         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
+         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
+  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
+  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = selected, color=type)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) 
+```
+
+### hard PD
+
+In hard PD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AB/BA is welfare maximizing.
+
+```{r}
+subgroup_outcome %>%
+  filter(gametype == "hardPD") %>%
+  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
+         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
+         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
+  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
+  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = selected, color=type)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  geom_hline(aes(yintercept = .5)) 
+```
+So if you do use the chat, you tend to get the best option. (Using chat means that easy and hard PD look different)
+
 <!--

### 2025-06-14 manuscript/manuscript.Rmd
do models for ms

---
title: "Game theory and language: a cautionary tale"
authors:
  - name: Veronica Boyce
    department: Department of Psychology
    affiliation: Stanford University
    email: vboyce@stanford.edu
  - name: Avidit Acharya
    department: Department of Political Science
    affiliation: Stanford University
  - name: Michael C. Frank
    department: Department of Psychology
    affiliation: Stanford University
abstract: |
  An attempt was made TODO
bibliography: references.bib
biblio-style: unsrt
output: 
  rticles::arxiv_article:
    extra_dependencies: ["float"]

---

# Introduction

How do we use language to coordinate with each other?

The use of language and the formation of linguistic conventions in pure reference games is studied frequently, but the role of language is studied less in other situations with more complicated goals that may not be fully aligned between participants.

In prior work, we tried to look at the role of language in a 3-player negotiation game where incentives were either fully-aligned or competitive between players. We found some evidence that language use improved performance, and that the referential language used to pick out targets showed some signs of conventionalization over time. However, this was a game we had created, so there were not theoretical or analytic predictions about what the optimal strategies were either with or without language, and therefore, when or how communication (via language) might improve performance.

Here, we attempted to bridge that gap, by looking at the role of language in two classic game theory games. By using set-ups that are well studied in the game theory literature, we had a rich body of work understanding how rational agents would play, and when linguistic coordination could be useful.

We focus on two games from within the 2x2 game space: Prisoner's Dilemma (PD) and Bach or Stravinsky (BoS).

TODO payout table

From game theory analysis, communication should not help in PD as it would be "cheap talk" -- that is, regardless of what the opponent player says, the equilibrium does not change, because there is not an incentive to be truthful. Thus, the prediction from game theory is that among rational agents playing a one-shot game of PD, they will end up in the Nash equilibrium.

For BoS, there are two pure Nash equilibrium, and thus there is a coordination challenge to agree on which one to go to. In this case, a message from one player saying what they will select will be credible -- having said that, they are incentivized to act on it, and their partner can do better by listening to it. There may be competition over which of the two equilibrium options to agree on, but both players want to go for the same cell. Thus, the prediction is that language will be beneficial in BoS.

To foreshadow the results, we found that *using* language showed some effect of helping improve results when people used it (including the possibility of cross-round transfer). However, the rate of uptake of the chat interface was low (despite our best efforts), and so we did not find robust effects of condition (chat or no chat). We also cannot know that differences between *those who could use and did use the chat* and other groups were due to the chat use as opposed to selection confounds on who chooses to use the chat. A qualitative look at the language and post-experiment reports reveals various ways overall strategies for achieving fairness (in both conditions) and various strategies for identifying the target each round (when language was used).

Thus, we still suspect that there probably is something interesting about when language is strategically useful and what sort of procedural and referential language is used. That said, we are unable to provide evidence because over 5 experiments (+ pilots) we continually ran into issues getting sufficient uptake of the condition manipulation, from which we eventually concluded that it was not worth our running additional experiments.

All data and code is available.

# Methods
```{r set-up, include=F}
knitr::opts_chunk$set(
  out.width = "\\textwidth",
  fig.align = "center", fig.width = 8, fig.height = 4, fig.crop = F, fig.align = "center", fig.env = "figure",
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = F,
  message = F, sanitize = T
)
library(tidyverse)
library(jsonlite)
library(here)
library(rlang)
library(lme4)
library(brms)
library(rstan)
library(viridis)
library(knitr)

theme_set(theme_bw())

source(here("manuscript/prep.R"))

stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

```

## Overall methods

All experiments were implemented in Empirica (v1). We recruited participants via Prolific and directed them to our website where they were paired up to play a game with another participant in real time.

The experiment consisted first of consent and instructions. Then participants (in all conditions) (TODO confirm which experiments this was true for) were given 3 minutes to freely chat with their partner using the chat box. This was intended to convince participants their partner really was human and not a bot (after high "is a bot" ratings in pilot studies). A few potential prompts for conversation were given. The exact instructions were changed in later experiments after LLM chatbots became common to explicitly tell participants TODO what we actually said.

Then participants moved onto the main experiment, where on each trial, they saw a payout grid for differently colored treasure chests and were asked to click which chest they wanted to open. Each pair completed 40 trials, with the payoffs dependent on the experiment condition (see below). In some experiments, we showed a running payout score for each participant. (TODO which expts).

After the experiment, participants completed a survey about their experience, including their strategy and whether they thought they were playing with a real person, before seeing a debrief and being redirected back to prolific to receive payment.

```{r}
all_rounds |>
  select(gameId, expt, chat_cond) |>
  unique() |>
  group_by(expt, chat_cond) |>
  tally() |>
  mutate(n = 2 * n) |>
  pivot_wider(names_from = chat_cond, values_from = n) |>
  kable(caption = "Number of participants in each experiment and each condition.")
```

## Conditions
For all experiments, we set parameters for the ratios between trial types and the ranges of payout values for each trial type. We used a sampling and randomization approach (see code for details), so different games saw different orders of trials and different values of trials, from the same distributions.

All experiments did both chat and no-chat versions as a between groups manipulation. 
TODO diagram 
Expt 1 (https://osf.io/8fnze): Between - group manipulation of either all PD trials or all BoS trials. PD trials were created by sampling 3 values from 1-9 (without replacement) for the payoffs, with 0 always as the sucker payoff, with the preference order corresponding to PD structure. For BoS, the off-diagonal payoff was always 1, the other payoffs were sampled from 2-9 (without replacement).

TODO confirm payoffs!

Expt 2 (https://osf.io/5au2r): A within-group version of experiment 1. Each game had roughly half and half PD and BoS trials, with trials constructed in the same way as experiment 1.

During data analysis for experiment 2, we realized that we had been misdisplaying the off diagonal rewards for PD for one player in each game. This was fixed for subsequent experiments.

Expt 3 (https://osf.io/c274z): Each game had a mix of two different tyes of PD and two different types of BoS, so that overall roughly half of trials were BoS and half were PD. Of the 40 trials, exactly 4, at pre-set locations in the trial order were "spiked" BoS where one of the rewards was high (sampled from 25-30) and the other was normal (sampled from 3-7). Due to the higher stakes, we thought pairs might coordinate more on these trials. We also included roughly 16 "normal" BoS trials with payouts sampled as in Experiments 1 and 2.

When randomly sampling numbers to populate PD trials, generally, the welfare maximizing option (highest summed reward) comes from both cooperating. Sometimes, however the payout for defecting is greater than twice the both-cooperate value and so the welfare maximizing option is to agree to defect-cooperate. Thus, in this situation, one might be able to cooperate for an overall better outcome than could otherwise be achieved, so we aimed to see whether people would recognize this type.

Of the roughly 20 trials in experiment 3 that were PD type, half were "easy" PD trials where where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff). All rewards are chosen from [1-12] conditional on this distribution. The "sucker" payoff was 0. The other half of PD trials (roughly 1/4 of the overall game) were "hard" PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff). All rewards are chosen from [1-12] conditional on this constraint.

Expt 4 (): In experiment 4, we tried to get a set-up that was closer to the predictions from game theory. We switched from providing a bonus proportional to the overall points earned to an incentive compatible bonus that sampled TODO N trials at random and paid out those points. We also removed the points counter, given that participants previously reported using the visible money earned so far as a coordination locus. TODO

Expt 5 (): The condition distribution of experiment 5 was identical to experiment 3.  TODO

## Data processing
All data processing was done in R, the code is available. TODO

# Results

the goal was to look at how people might use langauge to coordinate and thus achieve better outcomes than they could without language. We were also interested in what sort of coordination language might be used, and whether procedural or meta-conventions might occur.

In terms of outcomes, we can look at points earned (which includes the random noise from what the points spreads were). We can also look at what outcome quadrant participants are selecting.

One issue that arose was that we had some difficulty convincing participants their partners were indeed real humans, and substantial difficulty creating a situation where people in the chat condition would choose to use the chat.

We also do analysis breaking out our chat condition into whether or not people chose to use the chat. However, this breaks the randomization and introduces as confounds any exogenous factor (other than access to chat) that determines whether people chat (i.e. contentiousness, motivation, etc).

## Manipulation results

The interpretability of results depends on two factors: a) whether participants generally think they are playing with another human and b) whether chat use actually differs substantially by condition.

```{r bot}
exit_1 <- read_csv(here(study_1_loc, "exit.csv")) |>
  group_by(game_cond, chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(expt = str_c("1", "_", game_cond)) |>
  ungroup()

exit_2 <- read_csv(here(study_2_loc, "exit.csv")) |>
  group_by(chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(expt = "2")

exit_3 <- read_csv(here(study_3_loc, "exit.csv")) |>
  group_by(chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(expt = "3")

exit_4 <- read_csv(here(study_4_loc, "exit.csv")) |>
  group_by(chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(expt = "4")

exit_5 <- read_csv(here(study_5_loc, "exit.csv")) |>
  group_by(chat_cond, human) %>%
  tally() %>%
  pivot_wider(names_from = human, values_from = n) %>%
  mutate(pct = yes / (no + yes)) |>
  mutate(expt = "5")

exit <- exit_1 |>
  bind_rows(exit_2) |>
  bind_rows(exit_3) |>
  bind_rows(exit_4) |>
  bind_rows(exit_5) |>
  select(expt, chat_cond, pct) |>
  mutate(chat_cond = ifelse(chat_cond == "chat", "chat", "nochat"))

exit |>
  mutate(pct = round(pct * 100, 1)) |>
  pivot_wider(names_from = chat_cond, values_from = pct) |>
  kable(caption = "Percent of participants in each condition who thought their partner was human as reported in an exit survey.")
```

For the most part, we succeeded in convincing players their partner was indeed human. Rates of believing partner was human were slightly higher in the chat conditions than no-chat conditions.

Our other manipulation check was how well participants took up the manipulation of being able to use a chat to communicate with their partner.

As shown in Figure TODO, groups vary substantially in how many trials they use the chat on, but many groups rarely use the chat. The horizontal lines correspond to 10% and 50% of the time, so games lower than the 10% line are using chat on fewer than 4 (out of 40) trials. There's a lot of not chatting in the access-to-chat condition.

```{r chat-pct, fig.width=10, fig.height=3, }

game_chat <- all_chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(all_rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, gametype, cond, chat_cond, repNum, expt) %>%
  summarize(
    words = sum(words),
    mean_payout = mean(payoff),
    is_chat = ifelse(words > 0, 1, 0)
  )

game_chat |>
  filter(chat_cond == "chat") |>
  mutate(is.chat = ifelse(words > 0, 1, 0)) %>%
  group_by(gameId, expt) %>%
  summarize(pct_chat = mean(is.chat)) |>
  arrange(expt, pct_chat) |>
  group_by(expt) |>
  mutate(x = 1 / n() * row_number()) |>
  ggplot(aes(x = x, y = pct_chat)) +
  geom_point() +
  facet_grid(.~expt) +
  coord_cartesian(ylim = c(-.05, 1.05),xlim=c(-.05, 1.05), expand=F) +
  geom_hline(yintercept = .1) +
  geom_hline(yintercept = .5) +
  labs(x="Fraction games", y="Fraction trials with chat messages")
```

## Main results

The prediction is that having access to the communication modality will be helpful for aligning coordination in BoS where both players want to choose the same thing.

Coordination should be less helpful in PD. According to theory, for single-shot, rational actors should choose the nash equilibrium. Participants are nicer than that and instead coordinate on the unstable mutually beneficial equilibrium.

After observing that occasionally there were PD trials were the welfare-maximizing option was for one person to get a large payout and the other to get nothing (called "hard" or "sacrifice" PD trials in expt 3 onwards), we explicitly sampled both this "hard" structure where the best overall outcome is to coordinate on one person getting the big reward and balance out later by switching off.

In contrast, we refer to the trials in which only one option is welfare maximizing as "easy PD" where we do not expect language to help.

The first way we assess whether access to chat makes a difference on game performance is to look at the mean number of points recieved per trial. Because BoS and PD have different point structures, we compare within game type.

```{r points_pd, fig.cap="TODO fig cap"}
bonuses <- all_rounds %>%
  group_by(playerId, gameId, chat_cond, cond, gametype, expt) %>%
  summarize(payoff = mean(payoff))

ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(cond, "_", gametype), group = cond)) +
  geom_point(position = position_jitterdodge(jitter.width = .1, dodge.width = .7), alpha = .3) +
  facet_grid(gametype~expt) +
  scale_color_brewer(palette = "Dark2") +
  labs(y = "mean_reward", x = "condition") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .7), size = .7, color = "black") +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .7), size = .7, color = "black", geom="line") +

  theme(legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial", ) +
  theme(
    legend.title=element_blank(),
    axis.title.x=element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )
```

TODO set per-condition color scheme!

In BoS, there is a generally a slightly higher score in the chat condition than the nochat condition, whereas there is not a difference between condtions in PD.

Looking at points introduces some noise as the points were varying even within a type of game. To understand more what is being selected, we can look at what quadrant games are ending up in.

TODO include topline model results

### Quadrant analysis per trial

```{r}
outcome <- all_rounds %>%
  select(gameId, repNum, response, chat_cond, gametype, cond, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

For BoS, the desirable outcome is to both choose the same box. As can be seen in Figure TODO, this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.

```{r, fig.cap="In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both. "}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond), y = outcome_parity, color=cond, group=cond)) +
  facet_grid(~ expt) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")+
  theme(legend.position="bottom")
```
We again note that the off-diagonal rewards were misdisplayed for one of the players in experiments 1 and 2 for PD. On most trials "normal_PD" will pattern with easy PD.

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "PD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0)) %>%
  mutate(AB_BA= ifelse(outcome %in% c("AB", "BA"), 1, 0)) %>%
  mutate(BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(c("AA", "AB_BA", "BB"), names_to="outcome_type", values_to="selected") |> 
  ggplot(aes(x = chat_cond, y = selected, color=cond)) +
  facet_grid(outcome_type ~ expt) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
  geom_hline(aes(yintercept = .5)) +
  theme(legend.position="bottom")
```
From Figure TODO, we see that the preferred outcome is AA which is welfare-maximizing in easy PD. This is not what game theory predicts for rational actors, instead, participants are being nice and cooperative with each other.

There is a slight tendency for more AA selections in easy than hard PD, especially in the chat conditions, but slight.

# Analyses based on use of chat

## Per game does talking help?

One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.

So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?

```{r}
group_talk <- game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words), pct_chat = mean(is_chat))

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(gametype == "BoS", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

So, in mixed games, talking on more trials helps regardess of whether you talked on this particular trial. (Although overall talk might also be confounding with conscientiousness..., to fix you'd have to experimentally manipulate when chat is or isn't available)

Talking more might help with hard PD, maybe in mixed?

```{r}
ggplot(overall_talk |> filter(gametype == "BoS", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Here total words is going to mix how much you talk each time with how often you talk, so maybe this is just totally redundant with previous. More talk helps on BoS, even as transfer from other trials. PD seems messier?

## Quadrant analysis per trial

 - chat_0 is had access to chat and didn't use on that trial
 - chat_1 is used chat on that trial
 - nochat_0 did not have access to chat

```{r}
outcome <- all_rounds %>%
  select(gameId, repNum, response, chat_cond, gametype, cond, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

## BoS

In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
```

Near chance if you can't talk, above chance if you could but don't,  far above chance if you do.

```{r}
outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = repNum, y = outcome_parity, color = cond)) +
  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes", y = "Good outcome?")
```

## PD

Expts 1 and 2 are suspect, but on most trials will pattern with easy PD.

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

```{r}
outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing AA outcome")

outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AB", "BA"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing AB or BA outcome")

outcome %>%
  filter(gametype == "PD") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
  facet_grid(. ~ cond) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "PD:fraction choosing BB outcome")
```

Can get a reasonable option no matter what. If you talk, you can get the uneven, but welfare-maximizing in hard PD.

Looks like using language tends to help, on the trials it's used on, and slightly on the trials it's not used on for BoS. Helps in some cases for PD.

Is there a dose-response relationship, or is one word enough?

```{r}
ggplot(game_chat |> filter(gametype == "BoS", words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(game_chat |> filter(gametype == "PD", words > 0), aes(x = log(words), y = mean_payout)) +
  geom_point() +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Especially where we have more data, looks like one word is enough. Indicative of coordination rather than negotiation, probably?

## Per game does talking help?

One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.

So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?

```{r}
group_talk <- game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words), pct_chat = mean(is_chat))

overall_talk <- game_chat |> left_join(group_talk)
ggplot(overall_talk |> filter(gametype == "BoS", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
  # stat_summary(fun.data="mean_cl_boot")+
  # geom_point()+
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

So, in mixed games, talking on more trials helps regardess of whether you talked on this particular trial. (Although overall talk might also be confounding with conscientiousness..., to fix you'd have to experimentally manipulate when chat is or isn't available)

Talking more might help with hard PD, maybe in mixed?

```{r}
ggplot(overall_talk |> filter(gametype == "BoS", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")

ggplot(overall_talk |> filter(gametype == "PD", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
  facet_grid(gametype ~ cond) +
  geom_smooth(method = "lm")
```

Here total words is going to mix how much you talk each time with how often you talk, so maybe this is just totally redundant with previous. More talk helps on BoS, even as transfer from other trials. PD seems messier?

# Language results: what do people say
and what does this say about coordination strategies

## Language

How much language?

Filter only for games that talked at least a little.

Second graph filters for *trials* that talked.

```{r}
game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  mutate(expt = str_sub(cond, -1)) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point() +
  facet_wrap(~expt) +
  geom_smooth()

game_chat |>
  group_by(gameId) |>
  summarize(total_words = sum(words)) |>
  filter(total_words != 0) |>
  left_join(game_chat) |>
  filter(words > 0) |>
  mutate(expt = str_sub(cond, -1)) |>
  ggplot(aes(x = repNum, y = words)) +
  geom_point() +
  facet_wrap(~expt) +
  geom_smooth()
```

## how often do *both* people talk

```{r}
pair_chat <- all_chat %>%
  filter(!is.na(targets)) %>%
  select(-type) %>%
  full_join(all_rounds) %>%
  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
  group_by(gameId, playerId, gametype, cond, chat_cond, repNum) %>%
  summarize(
    words = sum(words),
    is_chat = ifelse(words > 0, 1, 0)
  ) |>
  group_by(gameId, gametype, cond, chat_cond, repNum) |>
  summarize(how_many_chat = sum(is_chat)) |>
  group_by(gameId, gametype, cond, chat_cond) |>
  summarize(pct_both = mean(how_many_chat == 2)) |>
  filter(chat_cond == "chat")

ggplot(pair_chat, aes(x = cond, y = pct_both)) +
  geom_jitter(width = .2) +
  facet_wrap(~gametype)
```
Even in games that talk, there aren't *that* many trials where both people talk?

Probably going to need to manually look at the language for negotiation v "choose red" "yup" etc.

# Discussion 
Some caveats: Despite our desire to use game-theory aligned paradigms for the analytic theory and predictions, the experiments had several factors that cause potential theoretical misalignment.

1. For practical reasons, we ran repeated trials with the same partners. This iterated approach is also studied, but potentially changes what exactly the predictions are.

2. The incentives of participants are not the same as the incentives of the "agents". People have values above money such as their reputation and desire to be nice to other people. Additionally, the monetary value participants get from the task is not what is shown in the boxes. They were paid a base rate for participation + a bonus aligned to task performance. Strategies of random clicking could be cost efficient if it means the task takes sufficiently shorter that base rate / short time is a better wage than (base rate + higher bonus)/ longer time.

3.

# Appendix: Model Results

Here we present the statistical results we ran. Note that these are not all the statistics we ever ran and may not be the pre-registered ones. These are what I thought best summarized the data descriptively and addressed the hypotheses at the time this summary was written.

```{r}
form <- function(model) {
  dep <- as.character(model$formula[2])
  ind <- as.character(model$formula[3])

  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("~", "~$\\\\sim$ ") |>
    str_replace_all("\\*", "~$\\\\times$ ") |>
    str_replace_all("\\+", "~+ ") |>
    str_replace_all("_", "")
}

do_table <- function(mod, cap, decimal = 2) {
  model <- read_rds(here(msum_loc, mod)) |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Term = str_replace_all(Term, "_", ""),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)

  spec <- read_rds(here(mform_loc, mod))

  kable(model |> select(Term, Est. = Estimate, `95\\% CrI` = `Credible Interval`),
    escape = F, format = "latex", booktabs = T, linesep = "",
    caption = str_c(cap, ": ", form(spec), sep = ""),
    align = "lll", position="H"
  )
}

msum_loc="manuscript/model_files/summary"
mform_loc="manuscript/model_files/formulae"
```

```{r}
do_table("payout_mega.rds", cap="TODO")
do_table("payout_mega_combo.rds", cap="TODO")
do_table("payout_combo_3.rds", cap="TODO")
do_table("payout_combo_4.rds", cap="TODO")
do_table("payout_combo_5.rds", cap="TODO")

do_table("BoS_aligned_mega.rds", cap="TODO")
do_table("BoS_aligned_3.rds", cap="TODO")
do_table("BoS_aligned_4.rds", cap="TODO")
do_table("BoS_aligned_5.rds", cap="TODO")

do_table("PD_cooperate_mega.rds", cap="TODO")
do_table("PD_cooperate_3.rds", cap="TODO")
do_table("PD_cooperate_4.rds", cap="TODO")
do_table("PD_cooperate_5.rds", cap="TODO")

do_table("PD_unneven_mega.rds", cap="TODO")
do_table("PD_unneven_3.rds", cap="TODO")
do_table("PD_unneven_4.rds", cap="TODO")
do_table("PD_unneven_5.rds", cap="TODO")

do_table("PD_nash_mega.rds", cap="TODO")
do_table("PD_nash_3.rds", cap="TODO")
do_table("PD_nash_4.rds", cap="TODO")
do_table("PD_nash_5.rds", cap="TODO")

```

 [1] "BoS_aligned_3.rds"     "BoS_aligned_4.rds"     "BoS_aligned_5.rds"    
 [4] "BoS_aligned_mega.rds"  "payout_mega_3.rds"     "payout_mega_4.rds"    
 [7] "payout_mega_combo.rds" "payout_mega.rds"       "PD_cooperate_3.rds"   
[10] "PD_cooperate_4.rds"    "PD_cooperate_5.rds"    "PD_cooperate_mega.rds"
[13] "PD_nash_3.rds"         "PD_nash_4.rds"         "PD_nash_5.rds"        
[16] "PD_nash_mega.rds"      "PD_unneven_3.rds"      "PD_unneven_4.rds"     
[19] "PD_unneven_5.rds"      "PD_unneven_mega.rds"

## payout models

First up, payout models.

## BoS alignment models

## PD alignment models

In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.

In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.

TODO run Bayesian
TODO save Bayesian
TODO tables

# Random crap

## Condition assignment on performance

```{r, eval=F}
perf_priors <- c(
  set_prior("normal(5,2)", class = "Intercept"),
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 2)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

perf_mod <- brm(
  payoff ~ has_chat * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_mod,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

perf_mod_tot <- brm(
  payoff ~ (use_chat | has_chat) * is_BoS +
    repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_mod,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)

freq_tot <- lm(payoff ~ (use_chat | has_chat) * is_BoS +
  repNum, data = for_mod)
```

## Words over time

log(words) ~ game_type * trial_number

```{r, eval=F}
for_log_words <- for_mod |>
  filter(words > 0) |>
  mutate(log_words = log(words))

lang_priors <- c(
  set_prior("normal(5,5)", class = "Intercept"),
  set_prior("normal(0, 5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

log_words_mod <- brm(
  log_words ~ is_BoS * repNum +
    (is_BoS * repNum | gameId),
  data = for_log_words,
  prior = lang_priors,
  control = list(adapt_delta = .95)
)
```

## Words on outcome

outcome ~ game_type * log(words)

```{r, eval=F}
perf_word_mod <- brm(
  payoff ~ log_words * is_BoS + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_log_words,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)
```

```{r, eval=F}
ggplot(for_log_words, aes(x = log_words, y = payoff, color = repNum)) +
  geom_jitter(alpha = .5) +
  geom_smooth(method = "lm") +
  facet_wrap(is_BoS ~ cond)
```

TODO think more about!!
outcome ~ log(words_first_10)*game_type*words_now ?

```{r, eval=F}
early_words <- for_mod |>
  filter(repNum < 10) |>
  group_by(gameId) |>
  summarize(early_words = sum(words))

for_early_word_mod <- for_mod |>
  filter(repNum > 9) |>
  left_join(early_words) |>
  filter(early_words > 0) |>
  mutate(log_early_words = log(early_words))

early_word_mod <- brm(
  payoff ~ log_early_words * is_BoS * is_chat + repNum +
    (is_BoS | gameId) +
    (is_BoS | playerId),
  data = for_early_word_mod,
  prior = perf_priors,
  control = list(adapt_delta = .95)
)
```
-->

### 2025-06-14 manuscript/prep.R

@@ -1,0 +1,141 @@
+library(tidyverse)
+library(jsonlite)
+library(here)
+library(rlang)
+
+ParseJSONColumn <- function(x) {
+  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
+    fromJSON(flatten = T)
+}
+
+## Data import constants
+
+
+study_1_loc <- "data/study1_220107"
+study_1_date_start <- lubridate::ymd("2022-01-07")
+
+study_2_loc <- "data/study2_220413"
+study_2_date_start <- lubridate::ymd("2022-04-13")
+
+study_3_loc <- "data/study3_220627"
+study_3_date_start <- lubridate::ymd("2022-06-24")
+
+study_4_loc <- "data/study4_240520"
+study_4_date_start <- lubridate::ymd("2024-05-20")
+
+study_5_loc <- "data/study5_241028"
+study_5_date_start <- lubridate::ymd("2024-10-26")
+
+# expt 1
+one_chat <- read_csv(here(study_1_loc, "raw_chat.csv")) %>%
+  mutate(gametype = game_cond) |>
+  mutate(cond = "normal") |>
+  mutate(expt = str_c("1", "_", gametype)) |>
+  select(expt, gameId, text, type, repNum, playerId, chat_cond, gametype, cond, targets)
+
+one_rounds <- read_csv(here(study_1_loc, "raw_results.csv")) %>%
+  mutate(gametype = game_cond) |>
+  mutate(cond = "normal") |>
+  mutate(expt = str_c("1", "_", gametype)) |>
+  select(expt, gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, cond, role)
+
+# expt 2
+two_chat <- read_csv(here(study_2_loc, "raw_chat.csv")) %>%
+  mutate(cond = "normal") %>%
+  mutate(expt = "2") |>
+  select(expt, gameId, text, type, repNum, playerId, chat_cond, gametype, cond, targets)
+
+two_rounds <- read_csv(here(study_2_loc, "raw_results.csv")) %>%
+  mutate(cond = "normal") %>%
+  mutate(expt = "2") |>
+  select(expt, gameId, repNum, playerId, payoff, response, time, box_A, box_B, AA, AB, BA, BB, chat_cond, gametype, cond, role)
+
+# expt 3
[…]

### 2025-06-14 manuscript/run_brm_mod.R

@@ -1,0 +1,256 @@
+library(tidyverse)
+library(here)
+library(brms)
+library(rstan)
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+source(here("manuscript/prep.R"))
+
+
+all_rounds_for_mod <- all_rounds |> 
+  mutate(is_chat_cond =ifelse(chat_cond=="chat", 1,0),
+         is_bos = ifelse(gametype=="BoS", 1,0),
+         expt=ifelse(expt %in% c("1_BoS", "1_PD"), "1", expt),
+         combo_type=str_c(gametype, "_", cond) |> factor(levels=c("PD_easy", "PD_normal", "PD_hard", "BoS_normal", "BoS_spike"))
+  )
+
+# TODO need to think about this more
+# trial order?
+
+perf_priors <- c(
+  set_prior("normal(5,2)", class = "Intercept"),
+  set_prior("normal(0, 2)", class = "b"),
+  set_prior("normal(0, 2)", class = "sd"))
+ # set_prior("lkj(1)", class = "cor"))
+
+
+payout_mega <- brm(payoff ~ is_chat_cond * is_bos + repNum+
+                     (1|expt) +
+                     (1|gameId),
+                   prior=perf_priors,
+                   control = list(adapt_delta = .95),
+                   file=here("manuscript/model_files/raw", "payout_mega"),
+                   data=all_rounds_for_mod)
+
+payout_mega_combo <- brm(payoff ~ is_chat_cond * combo_type+ repNum+
+                           (1|expt) +
+                           (1|gameId),
+                         prior=perf_priors,
+                         control = list(adapt_delta = .95),
+                         file=here("manuscript/model_files/raw", "payout_mega_combo"),
+                         data=all_rounds_for_mod)
+
+payout_3 <-  brm(payoff ~ is_chat_cond * combo_type +repNum+
+                   (1|gameId)+
+                   (1|playerId),
+                 prior=perf_priors,
+                 control = list(adapt_delta = .95),
+                 file=here("manuscript/model_files/raw", "payout_combo_3"),
+                 data=all_rounds_for_mod |> filter(expt=="3"))
+
+payout_4 <-  brm(payoff ~ is_chat_cond * combo_type + repNum+
+                   (1|gameId)+
+                   (1|playerId),
+                 prior=perf_priors,
+                 control = list(adapt_delta = .95),
+                 file=here("manuscript/model_files/raw", "payout_combo_4"),
[…]

### 2025-06-17 manuscript/manuscript.Rmd
many graph much fuss

@@
     extra_dependencies: ["float"]
+header-includes:
+ - \usepackage{float}
+ - \usepackage{setspace}\singlespacing
+ - \renewcommand{\textfraction}{0.00}
+ - \renewcommand{\topfraction}{1}
+ - \renewcommand{\bottomfraction}{1}
+ - \renewcommand{\floatpagefraction}{1}
+ - \setcounter{topnumber}{5}
+ - \setcounter{bottomnumber}{5}
+ - \setcounter{totalnumber}{6}  
+ - \usepackage{caption, subcaption}
+ - \usepackage{placeins}
 
@@
 
-As shown in Figure TODO, groups vary substantially in how many trials they use the chat on, but many groups rarely use the chat.
+As shown in Figure  \ref{chat-pct}, groups vary substantially in how many trials they use the chat on, but many groups rarely use the chat.
 The horizontal lines correspond to 10% and 50% of the time, so games lower than the 10% line are using chat on fewer than 4 (out of 40) trials.
@@
 
-```{r chat-pct, fig.width=10, fig.height=3, }
+```{r chat-pct, fig.width=10, fig.height=3, fig.cap="In chat conditions, how many trials each game choose to use the chat at all.
+Within each experiment, games are arranged in frequency of chat use.
+Horizontal lines denote 10% and 50% use of chat.
+\\label{chat-pct}"}
 
@@
   facet_grid(.~expt) + 
-  coord_cartesian(ylim = c(-.05, 1.05),xlim=c(-.05, 1.05), expand=F) +
   geom_hline(yintercept = .1) +
   geom_hline(yintercept = .5) +
-  labs(x="Fraction games", y="Fraction trials with chat messages")
+  labs(x="Fraction games", y="Fraction trials with chat messages")+
+    scale_x_continuous(breaks=c(0, .25, .5, .75, 1), labels=c("0", ".25", ".5", ".75", "1"))+
+      scale_y_continuous(breaks=c(0, .25, .5, .75, 1), labels=c("0", ".25", ".5", ".75", "1"))
 ```
@@
 ## Main results
+
+```{r}
+msum_loc="manuscript/model_files/summary"
+
+payout_mega <- read_rds(here(msum_loc, "payout_mega.rds"))
+
+payout_combo <- read_rds(here(msum_loc,"payout_mega_combo.rds"))
+payout_3 <- read_rds(here(msum_loc, "payout_combo_3.rds"))
+payout_4 <- read_rds(here(msum_loc, "payout_combo_4.rds"))
+payout_5 <- read_rds(here(msum_loc, "payout_combo_5.rds"))
+
+# 
+bos_aligned <- read_rds(here(msum_loc, "BoS_aligned_mega.rds"))
+bos_3 <- read_rds(here(msum_loc, "BoS_aligned_3.rds"))
+bos_4 <- read_rds(here(msum_loc, "BoS_aligned_4.rds"))
+bos_5 <- read_rds(here(msum_loc, "BoS_aligned_5.rds"))
+
+PD_cooperate <- read_rds(here(msum_loc, "PD_cooperate_mega.rds"))
+PD_unneven <- read_rds(here(msum_loc, "PD_unneven_mega.rds"))
+
+```
 
@@
 
-The first way we assess whether access to chat makes a difference on game performance is to look at the mean number of points recieved per trial.
+The first way we assess whether access to chat makes a difference on game performance is to look at the mean number of points received per trial.
 Because BoS and PD have different point structures, we compare within game type.
 
-```{r points_pd, fig.cap="TODO fig cap"}
+```{r points, fig.cap="Rewards achieved per trial depending on experiment, game condition, and chat condition. \\label{points}"}
+
+pal = c("PD_easy"="#1B9E77FF","PD_normal"= "#D95F02FF", "PD_hard"="#7570B3FF", "BoS_normal"= "#E7298AFF", "BoS_spike"="#66A61EFF")
 bonuses <- all_rounds %>%
   group_by(playerId, gameId, chat_cond, cond, gametype, expt) %>%
-  summarize(payoff = mean(payoff))
+  summarize(payoff = mean(payoff)) |> 
+  mutate( combo_type=str_c(gametype, "_", cond) |> factor(levels=c("PD_easy", "PD_normal", "PD_hard", "BoS_normal", "BoS_spike")))
 
-ggplot(bonuses, aes(x = chat_cond, y = payoff, color = str_c(cond, "_", gametype), group = cond)) +
+ggplot(bonuses, aes(x = chat_cond, y = payoff, color = combo_type, group = cond)) +
   geom_point(position = position_jitterdodge(jitter.width = .1, dodge.width = .7), alpha = .3) +
   facet_grid(gametype~expt) +
-  scale_color_brewer(palette = "Dark2") +
+  scale_color_manual(values=pal) +
   labs(y = "mean_reward", x = "condition") +
-  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .7), size = .7, color = "black") +
+  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .7), size = .3, color = "black") +
     stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .7), size = .7, color = "black", geom="line") +
-
+guides(color = guide_legend(override.aes = list(alpha = 1, size=3)))+
   theme(legend.position = "bottom") +
@@
     legend.title=element_blank(),
-    axis.title.x=element_blank(),
-    axis.text = element_text(size = 12),
-    axis.title = element_text(size = 14),
-    strip.text = element_text(size = 14)
+    axis.title.x=element_blank()
   )
@@
 
-TODO set per-condition color scheme!
+In a model of payouts by coarse conditions (BoS v PD), there is an interaction between condition and chat condition where chat condition leads to higher payout for BoS (`r stats_text(payout_mega, 4)`).
+Using finer grained measures of condition, and treating "easy PD" as baseline, there are interactions between chat and both normal BoS (`r stats_text(payout_combo, 7)`) and spike BoS (`r stats_text(payout_combo, 8)`).
+The same directionality occurs in individual experiments, but with wide credible intervals.
+In Experiment 3, the chat interaction with normal BoS is `r stats(payout_3, 6)` and with spike BoS is `r stats(payout_3, 7)`.
+In Experiment 4, the interaction between chat and normal BoS is `r stats(payout_4, 4)`.
+In Experiment 5, the chat interaction with normal BoS is `r stats(payout_5, 6)` and with spike BoS is `r stats(payout_5, 7)`.
+Full model results are in the Appendix.
 
@@
 Looking at points introduces some noise as the points were varying even within a type of game. To understand more what is being selected, we can look at what quadrant games are ending up in.
-
-TODO include topline model results
 
@@
   mutate(outcome = str_c(p1, p2)) %>%
-  left_join(game_chat)
+  left_join(game_chat) |> 
+   mutate( combo_type=str_c(gametype, "_", cond) |> factor(levels=c("PD_easy", "PD_normal", "PD_hard", "BoS_normal", "BoS_spike")))
+
 ```
@@
 For BoS, the desirable outcome is to both choose the same box.
-As can be seen in Figure TODO, this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.
+As can be seen in Figure \@ref(fig-bosquad), this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.
+Across experiments, the chat condition has a higher rate of good (aligned) outcomes (`r stats_text(bos_aligned, 3)`).
+The estimates are less certain for individual experiments.
+In experiment 3 the effect was `r stats_text(bos_3, 3)`; in experiment 4, `r stats_text(bos_4,2)`; and in experiment 5, `r stats_text(bos_5,3)`.
 
-```{r, fig.cap="In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both. "}
+```{r bosquad, fig.cap="For BoS trials, how frequently pairs made coordinated selections that recieved rewards. \\label{bosquad} ", fig.height=3, fig.width=8}
 outcome %>%
@@
   mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
-  ggplot(aes(x = str_c(chat_cond), y = outcome_parity, color=cond, group=cond)) +
+  ggplot(aes(x = str_c(chat_cond), y = outcome_parity, color=combo_type, group=combo_type)) +
   facet_grid(~ expt) +
+  scale_color_manual(values=pal)+
   stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
   geom_hline(aes(yintercept = .5)) +
-  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")+
+  labs(y = "Good outcome?", x="Chat condition", color="")+
   theme(legend.position="bottom")
@@
 
-```{r}
+```{r , fig.cap="For PD trials, how frequently pairs made selections that results in each type of outcome.
+AA is the cooperate outcome, AB/BA is mixed -- one cooperate, one defect, BB is both defect.
+\\label{pdquad} ", fig.height=5, fig.width=8}
 outcome %>%
@@
   pivot_longer(c("AA", "AB_BA", "BB"), names_to="outcome_type", values_to="selected") |> 
-  ggplot(aes(x = chat_cond, y = selected, color=cond)) +
+  ggplot(aes(x = chat_cond, y = selected, color=combo_type)) +
   facet_grid(outcome_type ~ expt) +
+    scale_color_manual(values=pal)+
+
   stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
   geom_hline(aes(yintercept = .5)) +
+  labs(y="fraction selected", x="Chat condition", color="")+
   theme(legend.position="bottom")
@@
 
-There is a slight tendency for more AA selections in easy than hard PD, especially in the chat conditions, but slight.
+Overall, for PD, the preference was to align on the fair-but-high square, regardless of whether chat was available.
+In the hard PD, AA selections were less common (`r stats_text(PD_cooperate, 2)`) and AA selections were slightly lowered still by a hard PD x chat interaction (`r stats_text(PD_cooperate, 5)`).
+The interaction effect had a credible interval overlapping 0 when experiments 3 or 5 were analyzed individually.
+In parallel, there was a hard PD x chat interaction increasing the rate of AB/BA selections (`r stats_text(PD_unneven, 5)`), consistent with the idea that this is the best outcome for hard PD, but requires more complex coordination to achieve.
 
-# Analyses based on use of chat
+Overall, the results based on condition assignment are consistent with the prediction that language would be more beneficial for BoS than for PD.
+However, conditional differences are small and often noisy (presumably due to low update of using language), and so the experiments do not provide compelling evidence.
 
-## Per game does talking help?
+## Uptake results
 
-One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.
-
-So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?
+To confirm that language use is, indeed useful, we also provide visualizations looking at both the option and choice to use language on a trial by trial basis.
+Note that a pairs' choice to use language is not assigned and thus this exploratory analysis could be confounded.
 
 ```{r}
-group_talk <- game_chat |>
-  group_by(gameId) |>
-  summarize(total_words = sum(words), pct_chat = mean(is_chat))
-
-overall_talk <- game_chat |> left_join(group_talk)
-ggplot(overall_talk |> filter(gametype == "BoS", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
-  # stat_summary(fun.data="mean_cl_boot")+
-  # geom_point()+
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-
-ggplot(overall_talk |> filter(gametype == "PD", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
-  # stat_summary(fun.data="mean_cl_boot")+
-  # geom_point()+
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
+outcome_chatted <- outcome |> 
+  mutate(chatted = case_when(
+    chat_cond =="nochat" ~ "nochat",
+    is_chat == 1 ~ "chatted",
+    is_chat == 0 ~ "quiet"
+  ) |> factor(levels=c("chatted", "quiet", "nochat")))
 ```
-
-So, in mixed games, talking on more trials helps regardess of whether you talked on this particular trial.
-(Although overall talk might also be confounding with conscientiousness..., to fix you'd have to experimentally manipulate when chat is or isn't available)
-
-Talking more might help with hard PD, maybe in mixed?
-
-```{r}
-ggplot(overall_talk |> filter(gametype == "BoS", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-
-ggplot(overall_talk |> filter(gametype == "PD", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-```
-
-Here total words is going to mix how much you talk each time with how often you talk, so maybe this is just totally redundant with previous.
-More talk helps on BoS, even as transfer from other trials.
-PD seems messier?
-
-## Quadrant analysis per trial
-
- - chat_0 is had access to chat and didn't use on that trial
- - chat_1 is used chat on that trial
- - nochat_0 did not have access to chat
-
-```{r}
-outcome <- all_rounds %>%
-  select(gameId, repNum, response, chat_cond, gametype, cond, role) |>
-  pivot_wider(names_from = role, values_from = response) %>%
-  mutate(outcome = str_c(p1, p2)) %>%
-  left_join(game_chat)
-```
-
-## BoS
 
@@
 
-```{r}
-outcome %>%
+```{r, fig.cap="For BoS trials, how frequently pairs made coordinated selections that recieved rewards as a function of whether chat was available or used that trial.
+\\label{bosquad-uptake} ", fig.height=3, fig.width=8}
+outcome_chatted %>%
   filter(gametype == "BoS") %>%
   mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
-  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
-  facet_grid(. ~ cond) +
-  stat_summary(fun.data = "mean_cl_boot") +
+  ggplot(aes(x = chatted, y = outcome_parity, color=combo_type, group=combo_type)) +
+  facet_grid(~ expt) +
+  scale_color_manual(values=pal)+
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
   geom_hline(aes(yintercept = .5)) +
-  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
+  labs(y = "Good outcome?", x="Chat usage", color="")+
+  theme(legend.position="bottom")
 ```
@@
 Near chance if you can't talk, above chance if you could but don't, far above chance if you do.
+This pattern suggests that some of the benefits to chatting spill over onto other trials, perhaps via setting up conventions for determining what strategy to use to break the symmetry.
+
+```{r, fig.cap="For PD trials, how frequently pairs made selections that results in each type of outcome as a function of chat availability and usage on each trial.
+AA is the cooperate outcome, AB/BA is mixed -- one cooperate, one defect, BB is both defect.
+\\label{pdquad} ", fig.height=5, fig.width=8}
+outcome_chatted %>%
+  filter(gametype == "PD") %>%
+  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0)) %>%
+  mutate(AB_BA= ifelse(outcome %in% c("AB", "BA"), 1, 0)) %>%
+  mutate(BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
+  pivot_longer(c("AA", "AB_BA", "BB"), names_to="outcome_type", values_to="selected") |> 
+  ggplot(aes(x = chatted, y = selected, color=combo_type)) +
+  facet_grid(outcome_type ~ expt) +
+    scale_color_manual(values=pal)+
+
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
+  geom_hline(aes(yintercept = .5)) +
+  theme(legend.position="bottom")
+```
+Noteably, the selection of AB/BA options in hard PD is primarly being driven by trials where people did communicate with each other.
+
+Another question is whether the amount of communication on a trial matters. Are more words more helpful?
 
 ```{r}
-outcome %>%
-  filter(gametype == "BoS") %>%
-  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
-  ggplot(aes(x = repNum, y = outcome_parity, color = cond)) +
-  facet_grid(~ str_c(chat_cond, "_", is_chat)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_hline(aes(yintercept = .5)) +
-  labs(title = "BoS:fraction choosing good outcomes", y = "Good outcome?")
-```
+game_chat |> filter(words > 0) |> 
+   mutate( combo_type=str_c(gametype, "_", cond) |> factor(levels=c("PD_easy", "PD_normal", "PD_hard", "BoS_normal", "BoS_spike"))) |> 
+  ggplot(aes(x = words, y = mean_payout, color=combo_type)) +
+  scale_x_log10(breaks=c(1,5,25))+
+  scale_color_manual(values=pal)+
+  geom_point(alpha=.1) +
+  facet_grid(gametype ~ expt) +
+  geom_smooth(method = "lm", show.legend=F)+
+  guides(color = guide_legend(override.aes = list(alpha = 1, size=3)))+
+  theme(legend.position="bottom")+
+  labs(color="", x="Words in a trial", y="Payout")
 
-## PD
-
-Expts 1 and 2 are suspect, but on most trials will pattern with easy PD.
-
-In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.
-
-In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.
-
-```{r}
-outcome %>%
-  filter(gametype == "PD") %>%
-  mutate(outcome_parity = ifelse(outcome %in% c("AA"), 1, 0)) %>%
-  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
-  facet_grid(. ~ cond) +
-  stat_summary(fun.data = "mean_cl_boot") +
-  geom_hline(aes(yintercept = .5)) +
-  labs(title = "PD:fraction choosing AA outcome")
-
-outcome %>%
-  filter(gametype == "PD") %>%
-  mutate(outcome_parity = ifelse(outcome %in% c("AB", "BA"), 1, 0)) %>%
-  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
-  facet_grid(. ~ cond) +
-  stat_summary(fun.data = "mean_cl_boot") +
-  geom_hline(aes(yintercept = .5)) +
-  labs(title = "PD:fraction choosing AB or BA outcome")
-
-outcome %>%
-  filter(gametype == "PD") %>%
-  mutate(outcome_parity = ifelse(outcome %in% c("BB"), 1, 0)) %>%
-  ggplot(aes(x = str_c(chat_cond, "_", is_chat), y = outcome_parity)) +
-  facet_grid(. ~ cond) +
-  stat_summary(fun.data = "mean_cl_boot") +
-  geom_hline(aes(yintercept = .5)) +
-  labs(title = "PD:fraction choosing BB outcome")
-```
-
-Can get a reasonable option no matter what. If you talk, you can get the uneven, but welfare-maximizing in hard PD.
-
-Looks like using language tends to help, on the trials it's used on, and slightly on the trials it's not used on for BoS. Helps in some cases for PD.
-
-Is there a dose-response relationship, or is one word enough?
-
-```{r}
-ggplot(game_chat |> filter(gametype == "BoS", words > 0), aes(x = log(words), y = mean_payout)) +
-  geom_point() +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-
-ggplot(game_chat |> filter(gametype == "PD", words > 0), aes(x = log(words), y = mean_payout)) +
-  geom_point() +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
+#TODO label in not log space 
 ```
@@
 
-## Per game does talking help?
+A qualitative perusal of the data suggests various strategies.
+On a trial by trial reference basis, cells can be identified by color or color combination, sometimes even reducing to single letter (such as r for red).
+The numeric values can also be used to identify.
 
-One idea is that talking on some trials may set up coordination strategies that can then effectively be used on later trials without talking on those trials.
-
-So we want to look at overall volume of talking (in words or in # of trial talked) as a predictor for performance, controlling for talk on that round?
-
-```{r}
-group_talk <- game_chat |>
-  group_by(gameId) |>
-  summarize(total_words = sum(words), pct_chat = mean(is_chat))
-
-overall_talk <- game_chat |> left_join(group_talk)
-ggplot(overall_talk |> filter(gametype == "BoS", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
-  # stat_summary(fun.data="mean_cl_boot")+
-  # geom_point()+
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-
-ggplot(overall_talk |> filter(gametype == "PD", chat_cond == "chat"), aes(x = pct_chat, y = mean_payout, color = as.character(is_chat))) +
-  # stat_summary(fun.data="mean_cl_boot")+
-  # geom_point()+
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-```
-
-So, in mixed games, talking on more trials helps regardess of whether you talked on this particular trial.
-(Although overall talk might also be confounding with conscientiousness..., to fix you'd have to experimentally manipulate when chat is or isn't available)
-
-Talking more might help with hard PD, maybe in mixed?
-
-```{r}
-ggplot(overall_talk |> filter(gametype == "BoS", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-
-ggplot(overall_talk |> filter(gametype == "PD", total_words > 0), aes(x = log(total_words), y = mean_payout, color = as.character(is_chat))) +
-  facet_grid(gametype ~ cond) +
-  geom_smooth(method = "lm")
-```
-
-Here total words is going to mix how much you talk each time with how often you talk, so maybe this is just totally redundant with previous.
-More talk helps on BoS, even as transfer from other trials.
-PD seems messier?
-
-# Language results: what do people say
-and what does this say about coordination strategies
-
-## Language
-
-How much language?
-
-Filter only for games that talked at least a little.
-
-Second graph filters for *trials* that talked.
-
-```{r}
-game_chat |>
-  group_by(gameId) |>
-  summarize(total_words = sum(words)) |>
-  filter(total_words != 0) |>
-  left_join(game_chat) |>
-  mutate(expt = str_sub(cond, -1)) |>
-  ggplot(aes(x = repNum, y = words)) +
-  geom_point() +
-  facet_wrap(~expt) +
-  geom_smooth()
-
-game_chat |>
-  group_by(gameId) |>
-  summarize(total_words = sum(words)) |>
-  filter(total_words != 0) |>
-  left_join(game_chat) |>
-  filter(words > 0) |>
-  mutate(expt = str_sub(cond, -1)) |>
-  ggplot(aes(x = repNum, y = words)) +
-  geom_point() +
-  facet_wrap(~expt) +
-  geom_smooth()
-```
-
-## how often do *both* people talk
-
-```{r}
-pair_chat <- all_chat %>%
-  filter(!is.na(targets)) %>%
-  select(-type) %>%
-  full_join(all_rounds) %>%
-  mutate(words = str_count(text, "\\W+") %|% int(0)) |>
-  group_by(gameId, playerId, gametype, cond, chat_cond, repNum) %>%
-  summarize(
-    words = sum(words),
-    is_chat = ifelse(words > 0, 1, 0)
-  ) |>
-  group_by(gameId, gametype, cond, chat_cond, repNum) |>
-  summarize(how_many_chat = sum(is_chat)) |>
-  group_by(gameId, gametype, cond, chat_cond) |>
-  summarize(pct_both = mean(how_many_chat == 2)) |>
-  filter(chat_cond == "chat")
-
-ggplot(pair_chat, aes(x = cond, y = pct_both)) +
-  geom_jitter(width = .2) +
-  facet_wrap(~gametype)
-```
-Even in games that talk, there aren't *that* many trials where both people talk?
-
-Probably going to need to manually look at the language for negotiation v "choose red" "yup" etc.
+Between the chat logs and the post game strategy notes, a number of strategies were identified.
+For example, some participants coordinated using the running score totals choosing the option that favored the person with a lower current score.
+Other identified strategies included fairness, being generous to the other person, or generally being fair but occassionally trying to take advantage and get a higher payout.
 
@@
 
-3.
+3. The repeated nature of the game and extra information such as running score allowed for coordinating in ways not available in the base single-shot games.
+
+One of the hopes was to see if using language on some perhaps early trials would set up conventions or patterns that could then be successfully executed on later with no or little additional language.
 
@@
 
+## Payout models
+
+These models all predict bonus score as an outcome in terms of either the coarse grained conditions (PD v BoS) or the finer grained conditions, all with interactions with assigned chat condition.
+
+Priors were normal(5,2) for intercept and normal(0,2) for beta and sd.
+Model were run both over all the experiments, and individually on each of experiments 3-5.
+Due to errors, we do not run separate models on 1 and 2.
+
 ```{r}
-do_table("payout_mega.rds", cap="TODO")
-do_table("payout_mega_combo.rds", cap="TODO")
-do_table("payout_combo_3.rds", cap="TODO")
-do_table("payout_combo_4.rds", cap="TODO")
-do_table("payout_combo_5.rds", cap="TODO")
-
-do_table("BoS_aligned_mega.rds", cap="TODO")
-do_table("BoS_aligned_3.rds", cap="TODO")
-do_table("BoS_aligned_4.rds", cap="TODO")
-do_table("BoS_aligned_5.rds", cap="TODO")
-
-do_table("PD_cooperate_mega.rds", cap="TODO")
-do_table("PD_cooperate_3.rds", cap="TODO")
-do_table("PD_cooperate_4.rds", cap="TODO")
-do_table("PD_cooperate_5.rds", cap="TODO")
-
-do_table("PD_unneven_mega.rds", cap="TODO")
-do_table("PD_unneven_3.rds", cap="TODO")
-do_table("PD_unneven_4.rds", cap="TODO")
-do_table("PD_unneven_5.rds", cap="TODO")
-
-do_table("PD_nash_mega.rds", cap="TODO")
-do_table("PD_nash_3.rds", cap="TODO")
-do_table("PD_nash_4.rds", cap="TODO")
-do_table("PD_nash_5.rds", cap="TODO")
+do_table("payout_mega.rds", cap="Mega-analytic model predicting bonus based on coarse game type and chat condition.")
+do_table("payout_mega_combo.rds", cap="Mega-analytic model predicting bonus based on fine-grained game type and chat condition.")
+do_table("payout_combo_3.rds", cap="Model predicting bonus for experiment 3. ")
+do_table("payout_combo_4.rds", cap="Model predicting bonus for experiment 4. ")
+do_table("payout_combo_5.rds", cap="Model predicting bonus for experiment 5. ")
 
@@
 
- [1] "BoS_aligned_3.rds"     "BoS_aligned_4.rds"     "BoS_aligned_5.rds"    
- [4] "BoS_aligned_mega.rds"  "payout_mega_3.rds"     "payout_mega_4.rds"    
- [7] "payout_mega_combo.rds" "payout_mega.rds"       "PD_cooperate_3.rds"   
-[10] "PD_cooperate_4.rds"    "PD_cooperate_5.rds"    "PD_cooperate_mega.rds"
-[13] "PD_nash_3.rds"         "PD_nash_4.rds"         "PD_nash_5.rds"        
-[16] "PD_nash_mega.rds"      "PD_unneven_3.rds"      "PD_unneven_4.rds"     
-[19] "PD_unneven_5.rds"      "PD_unneven_mega.rds"
+\FloatBarrier
 
-## payout models
+## BoS models
 
-First up, payout models.
+Abstracting over the randomly selected payout values, we also examined each class of game by where the selections fell in the 2x2 grid.
+For BoS, due to symmetry, outcomes are either aligned (positive outcome) or misaligned (0 value outcome).
+We can look at the effect of chat on getting aligned outcomes.
 
-## BoS alignment models
+These were logistic models with a prior of normal(0,1) for beta and sd.
 
-## PD alignment models
+```{r}
+do_table("BoS_aligned_mega.rds", cap="Mega-analytic model predicting occurrance of aligned outcome in BoS. ")
+do_table("BoS_aligned_3.rds", cap="Model predicting aligned BoS outcome for experiment 3. ")
+do_table("BoS_aligned_4.rds", cap="Model predicting aligned BoS outcome for experiment 4. ")
+do_table("BoS_aligned_5.rds", cap="Model predicting aligned BoS outcome for experiment 5. ")
 
-In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.
-
-In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.
-
-TODO run Bayesian
-TODO save Bayesian
-TODO tables
-
-# Random crap
-
-## Condition assignment on performance
-
-```{r, eval=F}
-perf_priors <- c(
-  set_prior("normal(5,2)", class = "Intercept"),
-  set_prior("normal(0, 2)", class = "b"),
-  set_prior("normal(0, 2)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
-
-perf_mod <- brm(
-  payoff ~ has_chat * is_BoS + repNum +
-    (is_BoS | gameId) +
-    (is_BoS | playerId),
-  data = for_mod,
-  prior = perf_priors,
-  control = list(adapt_delta = .95)
-)
-
-perf_mod_tot <- brm(
-  payoff ~ (use_chat | has_chat) * is_BoS +
-    repNum +
-    (is_BoS | gameId) +
-    (is_BoS | playerId),
-  data = for_mod,
-  prior = perf_priors,
-  control = list(adapt_delta = .95)
-)
-
-freq_tot <- lm(payoff ~ (use_chat | has_chat) * is_BoS +
-  repNum, data = for_mod)
 ```
 
-## Words over time
+## PD models
 
-log(words) ~ game_type * trial_number
+For PD, there are 3 different outcomes.
+The AA cooperative outcome, the AB/BA outcome, and the Nash equilibrium BB outcome.
+We ran separate logistic models for the occurrence of each outcome, although note that model results are inherently interdependent because each trial has exactly 1 of these outcomes.
 
-```{r, eval=F}
-for_log_words <- for_mod |>
-  filter(words > 0) |>
-  mutate(log_words = log(words))
+These were logistic models with a prior of normal(0,1) for beta and sd.
 
-lang_priors <- c(
-  set_prior("normal(5,5)", class = "Intercept"),
-  set_prior("normal(0, 5)", class = "b"),
-  set_prior("normal(0, 5)", class = "sd"),
-  set_prior("lkj(1)", class = "cor")
-)
+```{r}
+do_table("PD_cooperate_mega.rds", cap="Mega-analytic model predicting AA outcomes in PD trials")
+do_table("PD_cooperate_3.rds", cap="Model predicting AA outcomes in PD trials for experiment 3.")
+do_table("PD_cooperate_4.rds", cap="Model predicting AA outcomes in PD trials for experiment 4.")
+do_table("PD_cooperate_5.rds", cap="Model predicting AA outcomes in PD trials for experiment 5.")
 
-log_words_mod <- brm(
-  log_words ~ is_BoS * repNum +
-    (is_BoS * repNum | gameId),
-  data = for_log_words,
-  prior = lang_priors,
-  control = list(adapt_delta = .95)
-)
+do_table("PD_unneven_mega.rds", cap="Mega-analytic model predicting AB/BA outcomes in PD trials. ")
+do_table("PD_unneven_3.rds", cap="Model predicting AB/BA outcomes in PD trials for experiment 3.")
+do_table("PD_unneven_4.rds", cap="Model predicting AB/BA outcomes in PD trials for experiment 4.")
+do_table("PD_unneven_5.rds", cap="Model predicting AB/BA outcomes in PD trials for experiment 5.")
+
+do_table("PD_nash_mega.rds", cap="Mega-analytic model predicting BB (Nash equilibrium) outcomes in PD trials.")
+do_table("PD_nash_3.rds", cap="Model predicting BB outcomes in PD trials for experiment 3.")
+do_table("PD_nash_4.rds", cap="Model predicting BB outcomes in PD trials for experiment 4.")
+do_table("PD_nash_5.rds", cap="Model predicting BB outcomes in PD trials for experiment 5.")
+
 ```
-
-## Words on outcome
-
-outcome ~ game_type * log(words)
-
-```{r, eval=F}
-perf_word_mod <- brm(
-  payoff ~ log_words * is_BoS + repNum +
-    (is_BoS | gameId) +
-    (is_BoS | playerId),
-  data = for_log_words,
-  prior = perf_priors,
-  control = list(adapt_delta = .95)
-)
-```
-
-```{r, eval=F}
-ggplot(for_log_words, aes(x = log_words, y = payoff, color = repNum)) +
-  geom_jitter(alpha = .5) +
-  geom_smooth(method = "lm") +
-  facet_wrap(is_BoS ~ cond)
-```
-
-TODO think more about!!
-outcome ~ log(words_first_10)*game_type*words_now ?
-
-```{r, eval=F}
-early_words <- for_mod |>
-  filter(repNum < 10) |>
-  group_by(gameId) |>
-  summarize(early_words = sum(words))
-
-for_early_word_mod <- for_mod |>
-  filter(repNum > 9) |>
-  left_join(early_words) |>
-  filter(early_words > 0) |>
-  mutate(log_early_words = log(early_words))
-
-early_word_mod <- brm(
-  payoff ~ log_early_words * is_BoS * is_chat + repNum +
-    (is_BoS | gameId) +
-    (is_BoS | playerId),
-  data = for_early_word_mod,
-  prior = perf_priors,
-  control = list(adapt_delta = .95)
-)
-```
--->

### 2025-06-17 manuscript/manuscript.Rmd
work on figs and tables and methods

@@
   rticles::arxiv_article:
-    extra_dependencies: ["float"]
+    extra_dependencies: ["float", "subcaption", "multirow", "array"]
 header-includes:
  - \usepackage{float}
+ - \usepackage{subcaption}
+ - \usepackage{multirow}
+ - \usepackage{array}
  - \usepackage{setspace}\singlespacing
@@
 
+```{=latex}
+	\begin{figure}
+		\caption{Reward structure for each game. In each cell, the reward for X is shown before the reward for Y. }
+		\label{payoff-matrix}
+		\setlength{\extrarowheight}{2pt}
+		\begin{subfigure}[b]{0.45\textwidth}
+			\caption{PD generalized -- Each player prefers B to A, but players prefer AA to BB. }
+			\label{payoff-PD}
+		\begin{tabular}{cc|c|c|}
+			& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+			& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+			\multirow{2}*{Player $X$}  & $A$ & $(2,2)$ & $(0,3)$ \\\cline{3-4}
+			& $B$ & $(3,0)$ & $(1,1)$ \\\cline{3-4}
+		\end{tabular}
+		\end{subfigure}
+		~~~~
+			\begin{subfigure}[b]{0.45\textwidth}
+			\caption{BoS generalized-- X prefers AA and Y prefers BB, but both prefer either to AB or BA.}
+			\label{payoff-BoS}
+			\begin{tabular}{cc|c|c|}
+				& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+				& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+				\multirow{2}*{Player $X$}  & $A$ & $(2,1)$ & $(0,0)$ \\\cline{3-4}
+				& $B$ & $(0,0)$ & $(1,2)$ \\\cline{3-4}
+			\end{tabular}
+		\end{subfigure}
+		\bigskip
+
+		\bigskip
+								\begin{subfigure}[b]{0.45\textwidth}
+			\caption{Example PD easy: Here the total reward from AA is greater that the total reward from AB or BA.}
+			\label{payoff-PDeasy}
+			\begin{tabular}{cc|c|c|}
+				& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+				& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+				\multirow{2}*{Player $X$}  & $A$ & $(5,5)$ & $(0,7)$ \\\cline{3-4}
+				& $B$ & $(7,0)$ & $(2,2)$ \\\cline{3-4}
+			\end{tabular}
+		\end{subfigure}
+		~~~~
+		\begin{subfigure}[b]{0.45\textwidth}
+			\caption{Example BoS normal: The scale of the two rewards is relatively similar. }
+			\label{payoff-BoSnormal}
+			\begin{tabular}{cc|c|c|}
+				& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+				& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+				\multirow{2}*{Player $X$}  & $A$ & $(6, 2)$ & $(0,0)$ \\\cline{3-4}
+				& $B$ & $(0,0)$ & $(2,6)$ \\\cline{3-4}
+			\end{tabular}
+		\end{subfigure}
+
+		\bigskip
+		\bigskip
+			\begin{subfigure}[b]{0.45\textwidth}
+			\caption{Example PD hard: Here the total reward from AB or BA is greater than AA because the highest reward is more than double the second highest.}
+			\label{payoff-PDhard}
+			\begin{tabular}{cc|c|c|}
+				& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+				& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+				\multirow{2}*{Player $X$}  & $A$ & $(4,4)$ & $(0,11)$ \\\cline{3-4}
+				& $B$ & $(11,0)$ & $(2,2)$ \\\cline{3-4}
+			\end{tabular}
+		\end{subfigure}
+		~~~~
+				\begin{subfigure}[b]{0.45\textwidth}
+			\caption{Example BoS spike: One of the rewards is much higher than the other, raising the stakes of which of AA or BB to aim for. }
+			\label{payoff-BoSspike}
+			\begin{tabular}{cc|c|c|}
+				& \multicolumn{1}{c}{} & \multicolumn{2}{c}{Player $Y$}\\
+				& \multicolumn{1}{c}{} & \multicolumn{1}{c}{$A$}  & \multicolumn{1}{c}{$B$} \\\cline{3-4}
+				\multirow{2}*{Player $X$}  & $A$ & $(28,5)$ & $(0,0)$ \\\cline{3-4}
+				& $B$ & $(0,0)$ & $(5,28)$ \\\cline{3-4}
+			\end{tabular}
+		\end{subfigure}
+
+		\end{figure}
+
+```
+
 ## Overall methods
@@
 The experiment consisted first of consent and instructions.
-Then participants (in all conditions) (TODO confirm which experiments this was true for) were given 3 minutes to freely chat with their partner using the chat box.
-This was intended to convince participants their partner really was human and not a bot (after high "is a bot" ratings in pilot studies).
+Then participants (in all conditions) were given 3 minutes to freely chat with their partner using the chat box.
+Instructions for experiments 1-3 were: "You may have seen similar games where your partner is really a bot.
+This isn't one of those!
+You have been paired with another human participant who was also recruited from Prolific at the same time." In experiments 4 and 5, conducted after the widespread awareness of LLM chatbots like Chat-GPT, we became more explicit: "We are aware that there are similar games where your partner is a bot (or ChatGPT).
+*THIS IS NOT ONE OF THOSE.* We are cognitive scientists at a university, and our ethics approval (IRB) doesn't allow us to deceive participants.
+You have been paired with another participant who was also recruited from Prolific via the same task you were." This was intended to convince participants their partner really was human and not a bot (after high "is a bot" ratings in pre-experiment-1 pilot studies).
 A few potential prompts for conversation were given.
-The exact instructions were changed in later experiments after LLM chatbots became common to explicitly tell participants TODO what we actually said.
 
@@
 Each pair completed 40 trials, with the payoffs dependent on the experiment condition (see below).
-In some experiments, we showed a running payout score for each participant.
-(TODO which expts).
+In experiments 1,2,3 and 5, participants saw a running bonus score for themselves and their partner, as their bonus payment was proportional to the sum of the rewards they owned.
+On experiment 4, participants were instead paid proportional to the results of 4 randomly selected trials, and were not shown a cumulative payoff score.
 
@@
   pivot_wider(names_from = chat_cond, values_from = n) |>
-  kable(caption = "Number of participants in each experiment and each condition.")
+  kable(caption = "Number of participants in each experiment and each condition. \\label{participants}")
 ```
@@
 
-All experiments did both chat and no-chat versions as a between groups manipulation. 
+All experiments did both chat and no-chat versions as a between groups manipulation. The number of participants in each experiment and condition is in Table \ref{participants}. 
 TODO diagram
-Expt 1 (https://osf.io/8fnze): Between - group manipulation of either all PD trials or all BoS trials.
-PD trials were created by sampling 3 values from 1-9 (without replacement) for the payoffs, with 0 always as the sucker payoff, with the preference order corresponding to PD structure.
-For BoS, the off-diagonal payoff was always 1, the other payoffs were sampled from 2-9 (without replacement).
 
-TODO confirm payoffs!
+Trial types:
 
-Expt 2 (https://osf.io/5au2r): A within-group version of experiment 1. Each game had roughly half and half PD and BoS trials, with trials constructed in the same way as experiment 1.
+* normal PD trials were created by sampling 3 values from 1-9 (without replacement) for the payoffs, with 0 always as the sucker payoff, with the preference order corresponding to PD structure.
+
+* easy PD trials were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the second highest payoff (AA) being at least half of the highest AB/BA payoff.
+This ensures that the total payoff for AA is greater than or equal to the defect payoff.
+
+* hard PD trials  were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the highest payoff (AB/BA) being greater than double the cooperate AA payoff.
+This ensures that the defect payoff for AB/BA is greater than the total payoff for both cooperating.
+
+* normal BoS trials were created by sampling from 2-9 (without replacement) for the AA and BB payoffs. THe off diagonal payoffs were always 0.
+
+* spike BoS trials were created from sampling one reward from 3-7 and one reward from 25-30. Off diagonal payoffs were always 0.
+
+Experiments:
+
+*Expt 1* (https://osf.io/8fnze): Between - group manipulation of either all normal PD trials or all normal BoS trials.
+
+*Expt 2* (https://osf.io/5au2r): A within-group version of experiment 1. Each game had roughly half and half normal PD and normal BoS trials.
 
@@
 
-Expt 3 (https://osf.io/c274z): Each game had a mix of two different tyes of PD and two different types of BoS, so that overall roughly half of trials were BoS and half were PD.
-Of the 40 trials, exactly 4, at pre-set locations in the trial order were "spiked" BoS where one of the rewards was high (sampled from 25-30) and the other was normal (sampled from 3-7).
+*Expt 3* (https://osf.io/c274z): Each game had a mix of easy PD, hard PD, normal BoS, and spike BoS.
+Of the 40 trials, exactly 4, at pre-set locations in the trial order were spike BoS where one of the rewards was substantially higher than the other.
 Due to the higher stakes, we thought pairs might coordinate more on these trials.
-We also included roughly 16 "normal" BoS trials with payouts sampled as in Experiments 1 and 2.
+The other trials were sampled  in a 16:10:10 ratio from normal BoS, easy PD and hard PD for an overall roughly half BoS and half PD mixture.
 
@@
 Sometimes, however the payout for defecting is greater than twice the both-cooperate value and so the welfare maximizing option is to agree to defect-cooperate.
-Thus, in this situation, one might be able to cooperate for an overall better outcome than could otherwise be achieved, so we aimed to see whether people would recognize this type.
+Thus, in this situation, one might be able to cooperate for an overall better outcome than could otherwise be achieved, so we aimed to see whether people would recognize this type and whether language use would help.
 
-Of the roughly 20 trials in experiment 3 that were PD type, half were "easy" PD trials where where cooperating is welfare maximizing ( 2 * coop payoff greater than or equal to defect payoff).
-All rewards are chosen from [1-12] conditional on this distribution.
-The "sucker" payoff was 0.
-The other half of PD trials (roughly 1/4 of the overall game) were "hard" PD trials where having one person defect and the other take the sucker payoff is welfare maximizing (the defect payoff is > 2 * coop payoff).
-All rewards are chosen from [1-12] conditional on this constraint.
+*Expt 4* (https://osf.io/fymjc): In experiment 4, we tried to get a set-up that was closer to the predictions from game theory.
+We switched from providing a bonus proportional to the overall points earned to an incentive compatible bonus that sampled 4 trials at random and paid out those points.
+We also removed the points counter, given that participants previously reported using the visible money earned so far as a coordination locus.
+Games were a mix of roughly half easy PD and half normal BoS.
 
-Expt 4 (): In experiment 4, we tried to get a set-up that was closer to the predictions from game theory.
-We switched from providing a bonus proportional to the overall points earned to an incentive compatible bonus that sampled TODO N trials at random and paid out those points.
-We also removed the points counter, given that participants previously reported using the visible money earned so far as a coordination locus.
-TODO
-
-Expt 5 (): The condition distribution of experiment 5 was identical to experiment 3.  TODO
+*Expt 5* (https://osf.io/ksjmv): After a less successful experiment 4, we went back to the design of experiment 3.
+The condition distribution and incentive structure of experiment 5 was identical to experiment 3.
+Experiment 5 retained the modified pre-chat from experiment 4 that emphasized the absense of chatbots.
 
@@
   pivot_wider(names_from = chat_cond, values_from = pct) |>
-  kable(caption = "Percent of participants in each condition who thought their partner was human as reported in an exit survey.")
+  kable(caption = "Percent of participants in each condition who thought their partner was human as reported in an exit survey. \\label{human}")
 ```
 
-For the most part, we succeeded in convincing players their partner was indeed human. Rates of believing partner was human were slightly higher in the chat conditions than no-chat conditions.
+For the most part, we succeeded in convincing players their partner was indeed human (Table \ref{human}).
+Rates of believing partner was human were slightly higher in the chat conditions than no-chat conditions.
 
@@
 ```
-
+Points earned per trial in different games are shown in Figure \ref{points}, which suggests slight differences in reward for BoS games depending on chat versus nochat condition, and no differences for PD.
 In a model of payouts by coarse conditions (BoS v PD), there is an interaction between condition and chat condition where chat condition leads to higher payout for BoS (`r stats_text(payout_mega, 4)`).
@@
 The same directionality occurs in individual experiments, but with wide credible intervals.
-In Experiment 3, the chat interaction with normal BoS is `r stats(payout_3, 6)` and with spike BoS is `r stats(payout_3, 7)`.
-In Experiment 4, the interaction between chat and normal BoS is `r stats(payout_4, 4)`.
-In Experiment 5, the chat interaction with normal BoS is `r stats(payout_5, 6)` and with spike BoS is `r stats(payout_5, 7)`.
+In Experiment 3, the chat interaction with normal BoS is `r stats_text(payout_3, 6)` and with spike BoS is `r stats_text(payout_3, 7)`.
+In Experiment 4, the interaction between chat and normal BoS is `r stats_text(payout_4, 4)`.
+In Experiment 5, the chat interaction with normal BoS is `r stats_text(payout_5, 6)` and with spike BoS is `r stats_text(payout_5, 7)`.
 Full model results are in the Appendix.
@@
 For BoS, the desirable outcome is to both choose the same box.
-As can be seen in Figure \@ref(fig-bosquad), this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.
+As can be seen in Figure \ref{bosquad}, this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.
 Across experiments, the chat condition has a higher rate of good (aligned) outcomes (`r stats_text(bos_aligned, 3)`).
@@
 ```
-From Figure TODO, we see that the preferred outcome is AA which is welfare-maximizing in easy PD.
+From Figure \ref{pdquad}, we see that the preferred outcome is AA which is welfare-maximizing in easy PD.
 This is not what game theory predicts for rational actors, instead, participants are being nice and cooperative with each other.
@@
   geom_hline(aes(yintercept = .5)) +
-  labs(y = "Good outcome?", x="Chat usage", color="")+
+  labs(y = "Good outcome?", x="Chat usage on trial", color="")+
   theme(legend.position="bottom")
 ```
-
-Near chance if you can't talk, above chance if you could but don't, far above chance if you do.
+For BoS results, shown in Figure \ref{bosquad-uptake}, the outcomes are best on trials that chatted.
+Depending on experiment, trials that could talk are either intermediate, or pattern with the nochat condition.
 This pattern suggests that some of the benefits to chatting spill over onto other trials, perhaps via setting up conventions for determining what strategy to use to break the symmetry.
+However, most of the overall slight benefits from the chat condition are being driven by the small number of games and trials that used the chat.
 
@@
   geom_hline(aes(yintercept = .5)) +
+  labs(color="", y="Selected", x="Chat usage on trial" )+
   theme(legend.position="bottom")
 ```
+
+We similarly break down the PD results in terms of language use per trial in Figure \ref{pdquad}. 
 Noteably, the selection of AB/BA options in hard PD is primarly being driven by trials where people did communicate with each other.
 
-Another question is whether the amount of communication on a trial matters. Are more words more helpful?
+## Use of language
 
-```{r}
+From the uptake analysis, it appears that using chat has some benefits, especially on BoS trials.
+Does the amount of language used matter, or only its existance?
+In Figure \ref{wordpayout}, we compare rewards on a trial as a function of the number of words used, for trials where any language was used.
+Overall, there does not seem to be a benefit to using more words on a trial.
+
+```{r, fig.cap="On trials that used the chat, the relationship between number of words and the reward, for each experiment and condition. \\label{wordpayout} ", fig.height=4, fig.width=8}
 game_chat |> filter(words > 0) |> 
@@
   theme(legend.position="bottom")+
-  labs(color="", x="Words in a trial", y="Payout")
+  labs(color="", x="Words in a trial", y="Reward")
 
-#TODO label in not log space 
 ```
 
-Especially where we have more data, looks like one word is enough. Indicative of coordination rather than negotiation, probably?
+The utility of language, but without a clear dose-response relationship, suggests that the language may primarily be used for coordination rather than negotiation.
 

### 2025-06-19 manuscript/manuscript.Rmd
clean up text methods and start results

@@
 
-# Methods
 ```{r set-up, include=F}
@@
 
-## Overall methods
+# Methods
+We conducted a series of 5 real-time online experiments where pairs of participants were matched to play a sequence of 40 2x2 games together.
+Across the experiments, we varied the exact structure of trials, but all experiments involved trials with reward structures modeled on BoS and PD.
 
-All experiments were implemented in Empirica (v1).
+## Participants
+The number of participants included in each condition (after incomplete games were excluded) is shown in Table \ref{participants}.
 We recruited participants via Prolific and directed them to our website where they were paired up to play a game with another participant in real time.
+## Procedure
 
-The experiment consisted first of consent and instructions.
-Then participants (in all conditions) were given 3 minutes to freely chat with their partner using the chat box.
+All experiments were implemented in Empirica (v1). The experiment consisted first of consent and instruction.
+
+Then participants were matched with their partner and given 3 minutes to freely chat with their partner using the chat box.
+This initial free-chat was intended to convince participants their partner really was human and not a bot (after high levels of pilot participants reported they thought they had been playing with a bot).
+A few potential prompts for conversation were given.
 Instructions for experiments 1-3 were: "You may have seen similar games where your partner is really a bot.
@@
 *THIS IS NOT ONE OF THOSE.* We are cognitive scientists at a university, and our ethics approval (IRB) doesn't allow us to deceive participants.
-You have been paired with another participant who was also recruited from Prolific via the same task you were." This was intended to convince participants their partner really was human and not a bot (after high "is a bot" ratings in pre-experiment-1 pilot studies).
-A few potential prompts for conversation were given.
+You have been paired with another participant who was also recruited from Prolific via the same task you were."
 
-Then participants moved onto the main experiment, where on each trial, they saw a payout grid for differently colored treasure chests and were asked to click which chest they wanted to open.
+```{r,fig.cap="The user interface. Participants saw a reward table and selected what box to open. \\label{ui}"}
+knitr::include_graphics("overview.png")
+```
+
+After the three minutes, participants moved onto the main experiment.
+On each trial, they saw a payout grid for differently colored treasure chests and were asked to click which chest they wanted to open (Figure \ref{ui}).
 Each pair completed 40 trials, with the payoffs dependent on the experiment condition (see below).
-In experiments 1,2,3 and 5, participants saw a running bonus score for themselves and their partner, as their bonus payment was proportional to the sum of the rewards they owned.
+Participants in the "chat" condition could message their partner freely during the trials using a chat interface; participants in the "no chat" condition did not have access to the chat during the main experiment.
+
+In experiments 1,2,3 and 5, participants saw a running bonus score for themselves and their partner, as their bonus payment was proportional to the sum of the rewards they earned.
 On experiment 4, participants were instead paid proportional to the results of 4 randomly selected trials, and were not shown a cumulative payoff score.
 
-After the experiment, participants completed a survey about their experience, including their strategy and whether they thought they were playing with a real person, before seeing a debrief and being redirected back to prolific to receive payment.
+After the experiment, participants completed a survey about their experience, including their strategy and whether they thought they were playing with a real person, before seeing a debrief and being redirected back to Prolific to receive payment.
 
@@
 
-## Conditions
-For all experiments, we set parameters for the ratios between trial types and the ranges of payout values for each trial type.
-We used a sampling and randomization approach (see code for details), so different games saw different orders of trials and different values of trials, from the same distributions.
+## Trial types 
+We sampled different types of trials across different experiments.
+For each trial type, we constructed trials by randomly sampling reward values subject to some constraints.
+Examples of each trial type are shown in Figure \ref{payoff-matrix}
 
-All experiments did both chat and no-chat versions as a between groups manipulation. The number of participants in each experiment and condition is in Table \ref{participants}. 
-TODO diagram
-
-Trial types:
-
-* normal PD trials were created by sampling 3 values from 1-9 (without replacement) for the payoffs, with 0 always as the sucker payoff, with the preference order corresponding to PD structure.
-
-* easy PD trials were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the second highest payoff (AA) being at least half of the highest AB/BA payoff.
-This ensures that the total payoff for AA is greater than or equal to the defect payoff.
-
-* hard PD trials  were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the highest payoff (AB/BA) being greater than double the cooperate AA payoff.
-This ensures that the defect payoff for AB/BA is greater than the total payoff for both cooperating.
-
-* normal BoS trials were created by sampling from 2-9 (without replacement) for the AA and BB payoffs. THe off diagonal payoffs were always 0.
-
-* spike BoS trials were created from sampling one reward from 3-7 and one reward from 25-30. Off diagonal payoffs were always 0.
-
-Experiments:
-
-*Expt 1* (https://osf.io/8fnze): Between - group manipulation of either all normal PD trials or all normal BoS trials.
-
-*Expt 2* (https://osf.io/5au2r): A within-group version of experiment 1. Each game had roughly half and half normal PD and normal BoS trials.
-
-During data analysis for experiment 2, we realized that we had been misdisplaying the off diagonal rewards for PD for one player in each game. This was fixed for subsequent experiments.
-
-*Expt 3* (https://osf.io/c274z): Each game had a mix of easy PD, hard PD, normal BoS, and spike BoS.
-Of the 40 trials, exactly 4, at pre-set locations in the trial order were spike BoS where one of the rewards was substantially higher than the other.
-Due to the higher stakes, we thought pairs might coordinate more on these trials.
-The other trials were sampled  in a 16:10:10 ratio from normal BoS, easy PD and hard PD for an overall roughly half BoS and half PD mixture.
+**Normal PD** trials were created by sampling 3 values from 1-9 (without replacement) for the payoffs, with 0 always as the sucker payoff, with the preference order corresponding to PD structure.
 
@@
 Sometimes, however the payout for defecting is greater than twice the both-cooperate value and so the welfare maximizing option is to agree to defect-cooperate.
-Thus, in this situation, one might be able to cooperate for an overall better outcome than could otherwise be achieved, so we aimed to see whether people would recognize this type and whether language use would help.
+Thus, in this situation, one might be able to cooperate for an overall better outcome than could otherwise be achieved.
+We anecdotally observed this occur rarely in normal PD games in early experiments, so in later experiments, we explicitly sampled PD games based on which of both-cooperate or defect-cooperate was welfare-maximizing.
 
-*Expt 4* (https://osf.io/fymjc): In experiment 4, we tried to get a set-up that was closer to the predictions from game theory.
+**Easy PD** trials (Figure \ref{payoff-PDeasy}) were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the second highest payoff (AA) being at least half of the highest AB/BA payoff.
+This ensured that the total payoff for AA was greater than or equal to the defect payoff.
+
+**Hard PD** trials (Figure \ref{payoff-PDhard})  were created by sampling 3 values from 1-12 (without replacement) for the payoffs, with 0 always as the sucker payoff, conditioned on the highest payoff (AB/BA) being greater than double the cooperate AA payoff.
+This ensured that the defect payoff for AB/BA was greater than the total payoff for both cooperating.
+We predicted that cooperation and communication might lead to better outcomes in hard PD if participants coordinated on the welfare-maximizing but unequal AB or BA outcomes.
+
+**Normal BoS** trials (Figure \ref{payoff-BoSnormal}) were created by sampling from 2-9 (without replacement) for the AA and BB payoffs.
+The off diagonal payoffs were always 0.
+We predicted that communication would aid in coordinating which of AA or BB to choose.
+
+**Spike BoS** trials (Figure \ref{payoff-BoSspike}) were created from sampling one reward from 3-7 and one reward from 25-30.
+Off diagonal payoffs were always 0.
+We thought the higher rewards might lead to higher stakes and thus incentivize coordinating or negotiating about the selection more than in normal BoS trials.
+
+## Experiment conditions
+For all experiments, we set parameters for the ratios between trial types and the ranges of payout values for each trial type.
+We used a sampling and randomization approach, so different games saw different orders of trials and different values of trials, from the same distributions.
+
+All experiments had presence of chat as a between-groups manipulation. The number of participants in each experiment and condition is in Table \ref{participants}.
+
+Here we briefly describe the set of trial types used in each experiment and link to the individual pre-registrations.
+
+**Expt 1** (https://osf.io/8fnze): A between-group manipulation of either all normal PD trials or all normal BoS trials.
+
+**Expt 2** (https://osf.io/5au2r): A within-group version of experiment 1. Trials were sampled from a 50/50 distribution of normal BoS and normal PD trials.
+
+During data analysis for experiment 2, we realized that we had mis-displayed the off-diagonal rewards for PD for one player in each game in experiments 1 and 2.
+This was fixed for subsequent experiments.
+
+**Expt 3** (https://osf.io/c274z): Each game had a mix of easy PD, hard PD, normal BoS, and spike BoS trials.
+Of the 40 trials, exactly 4, at pre-set locations in the trial order were spike BoS trials.
+Due to the higher stakes, we thought pairs might coordinate more on these trials.
+The other trials were sampled in a 16:10:10 ratio from normal BoS, easy PD, and hard PD for an overall roughly half BoS and half PD mixture.
+
+**Expt 4** (https://osf.io/fymjc): In experiment 4, we aimed for a set-up that was closer to the predictions from game theory.We removed the points counter, given that participants in prior experiments sometimes used the visible money earned so far as a locus for coordination.
 We switched from providing a bonus proportional to the overall points earned to an incentive compatible bonus that sampled 4 trials at random and paid out those points.
-We also removed the points counter, given that participants previously reported using the visible money earned so far as a coordination locus.
-Games were a mix of roughly half easy PD and half normal BoS.
+Because the games that mattered were unknown and randomly sampled, we reasoned that the rational strategy might be closer to the formal predictions of one-shot games.
+Trials were sampled from a 50/50 distribution of easy PD and normal BoS.
 
-*Expt 5* (https://osf.io/ksjmv): After a less successful experiment 4, we went back to the design of experiment 3.
+**Expt 5** (https://osf.io/ksjmv): After a disappointing experiment 4, we returned to the design of experiment 3 and collected a larger sample.
 The condition distribution and incentive structure of experiment 5 was identical to experiment 3.
-Experiment 5 retained the modified pre-chat from experiment 4 that emphasized the absense of chatbots.
 
 ## Data processing
-All data processing was done in R, the code is available. TODO
+We excluded participant pairs that did not finish. Data processing and analysis was done in R. We ran regression models in BRMS and report the point estimates and 95% credible intervals.
+
+Experimental code, data, and analysis scripts are all available at https://github.com/vboyce/game-theory-language.
 
@@
 
-the goal was to look at how people might use langauge to coordinate and thus achieve better outcomes than they could without language.
-We were also interested in what sort of coordination language might be used, and whether procedural or meta-conventions might occur.
-
-In terms of outcomes, we can look at points earned (which includes the random noise from what the points spreads were). We can also look at what outcome quadrant participants are selecting.
-
-One issue that arose was that we had some difficulty convincing participants their partners were indeed real humans, and substantial difficulty creating a situation where people in the chat condition would choose to use the chat.
-
-We also do analysis breaking out our chat condition into whether or not people chose to use the chat.
-However, this breaks the randomization and introduces as confounds any exogenous factor (other than access to chat) that determines whether people chat (i.e.
-contentiousness, motivation, etc).
+Our goal was to test hypotheses around when using language would allow people to achieve better outcomes in a strategic game than they could without using language to coordinate.
+We also hoped to explore what sort of language use would be used for coordination, whether any negotiation would occur, and whether procedural and meta-conventions might occur.
 
@@
 
-The interpretability of results depends on two factors: a) whether participants generally think they are playing with another human and b) whether chat use actually differs substantially by condition.
+The interpretability of results depended on two factors: a) whether participants generally thought they were playing with another human and b) whether chat use actually differed substantially by chat condition.
 
@@
 
+In the exit survey, we asked participants if they thought their partner was human.
 For the most part, we succeeded in convincing players their partner was indeed human (Table \ref{human}).
@@
 
-Our other manipulation check was how well participants took up the manipulation of being able to use a chat to communicate with their partner.
-
-As shown in Figure  \ref{chat-pct}, groups vary substantially in how many trials they use the chat on, but many groups rarely use the chat.
-The horizontal lines correspond to 10% and 50% of the time, so games lower than the 10% line are using chat on fewer than 4 (out of 40) trials.
-There's a lot of not chatting in the access-to-chat condition.
+Our primary manipulation of interest was whether or not participants were assigned to the chat condition, where they could use a chat to communicate with their partner, or the no-chat condition where they could not.
+For groups in the chat condition, the rate of chat usage varied widely, with many participants who had access to chat not using it.
+As shown in Figure \ref{chat-pct}, many groups rarely use the chat.
+The low uptake of the chat manipulation considerably reduces our ability to determine how chat usage impacts performance.
 
@@
 
+Despite the low rates of chat usage, we focus our main analyses on the impact of the (assigned) chat condition.
+
 ```{r}
@@
 
-The prediction is that having access to the communication modality will be helpful for aligning coordination in BoS where both players want to choose the same thing.
-
-Coordination should be less helpful in PD.
-According to theory, for single-shot, rational actors should choose the nash equilibrium.
-Participants are nicer than that and instead coordinate on the unstable mutually beneficial equilibrium.
-
-After observing that occasionally there were PD trials were the welfare-maximizing option was for one person to get a large payout and the other to get nothing (called "hard" or "sacrifice" PD trials in expt 3 onwards), we explicitly sampled both this "hard" structure where the best overall outcome is to coordinate on one person getting the big reward and balance out later by switching off.
-
-In contrast, we refer to the trials in which only one option is welfare maximizing as "easy PD" where we do not expect language to help.
-
-The first way we assess whether access to chat makes a difference on game performance is to look at the mean number of points received per trial.
-Because BoS and PD have different point structures, we compare within game type.
+The predictions are that a communication channel should improve performance in BoS trials where communication is credible and can be used to coordinate but that communication channels should not improve performance in PD trials where communication is not credible.
+According to formal game theory, for single-shot games, rational actors should take actions that lead to Nash equilibria. Both AA and BB are equilibria in BoS, and only BB is a Nash equilibrium in PD.
 
@@
 ```
-Points earned per trial in different games are shown in Figure \ref{points}, which suggests slight differences in reward for BoS games depending on chat versus nochat condition, and no differences for PD.
-In a model of payouts by coarse conditions (BoS v PD), there is an interaction between condition and chat condition where chat condition leads to higher payout for BoS (`r stats_text(payout_mega, 4)`).
-Using finer grained measures of condition, and treating "easy PD" as baseline, there are interactions between chat and both normal BoS (`r stats_text(payout_combo, 7)`) and spike BoS (`r stats_text(payout_combo, 8)`).
-The same directionality occurs in individual experiments, but with wide credible intervals.
+
+We first analyse participant's per-trial rewards in terms of trial type and chat condition.
+Because each trial type had a different sampling range, points are not directly comparable across conditions.
+Points earned per trial in different games are shown in Figure \ref{points}.
+Visually, there are slightly higher average rewards for chat BoS games than no-chat BoS games, and no visible effect of chat condition on PD outcomes.
+
+In a mega-analytic model of rewards predicted by coarse trial type (BoS v PD) across all expeirments, trial type and chat condition interacted in the expected direction, where chat condition led to higher rewards for BoS (`r stats_text(payout_mega, 4)`).
+We also ran a mega-analytic model using fine grained trial types and treating easy PD as the baseline condition, again pooling across all experiments.
+Here, there was an interaction effect between chat and normal BoS (`r stats_text(payout_combo, 7)`) and between chat and spike BoS (`r stats_text(payout_combo, 8)`).
+
+When we analyze results on a per-experiment basis, as originally planned, the interaction results are directionally consistent, but with wide credible intervals that sometimes overlap 0.
 In Experiment 3, the chat interaction with normal BoS is `r stats_text(payout_3, 6)` and with spike BoS is `r stats_text(payout_3, 7)`.
@@
 In Experiment 5, the chat interaction with normal BoS is `r stats_text(payout_5, 6)` and with spike BoS is `r stats_text(payout_5, 7)`.
-Full model results are in the Appendix.
+Estimates for all coefficients in these models are in the Appendix.
 
-In BoS, there is a generally a slightly higher score in the chat condition than the nochat condition, whereas there is not a difference between condtions in PD.
+To summarize, the only effect of chat that we find is a very weak interaction effect where access to chat marginally improves outcomes in BoS games.
+However, the effect is hard to quantify due to the low rates of chat usage, even in the chat condition.
 
-Looking at points introduces some noise as the points were varying even within a type of game. To understand more what is being selected, we can look at what quadrant games are ending up in.
-
-### Quadrant analysis per trial
+Rewards were randomly sampled, which introduces noise. To look at participant selections more discretely, we also examine how frequently trials end up in each outcome quadrant.
 
@@
 
-For BoS, the desirable outcome is to both choose the same box.
-As can be seen in Figure \ref{bosquad}, this good outcome is achieved at above change levels when people can chat and usually at chance levels when people cannot.
+For BoS, the desirable outcome is to both choose the same box, with player X preferring the AA outcome and player Y preferring the BB outcome.
+As can be seen in Figure \ref{bosquad}, a good outcome is usually achieved at above chance levels in the chat condition and at chance levels in the no-chat condition.
 Across experiments, the chat condition has a higher rate of good (aligned) outcomes (`r stats_text(bos_aligned, 3)`).
@@
 ```
-We again note that the off-diagonal rewards were misdisplayed for one of the players in experiments 1 and 2 for PD. On most trials "normal_PD" will pattern with easy PD.
-
-In easyPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. AA is welfare maximizing.
-
-In hardPD: P1 prefers BA > AA > BB > AB and P2 prefers AB > AA > BB > BA. BA and AB are welfare maximizing.
+For PD trials, there are three distinct outcomes.
+If both players cooperate, the outcome is AA and each gets a moderately high payoff.
+If both players defect, the outcome is BB, the Nash equilibrium, and each player get a low payoff.
+If one player cooperates and the other defects, the outcome is AB (or BA), and one player get a high payoff and the other gets 0.
+From a perspective of wefare-maximization, the selection with best total reward is AA in easy PD and AB/BA in hard PD.
+We expect normal PD to generally be the of the type easy PD; and again note that the errors in displaying off-diagonal rewards would primarily affect hard PD type trials, where off diagonal options are more attractive.
+The rates of selecting each of the three outcome options by experiment and conditions are shown in Figure \ref{quad}.
 
@@
 ```
-From Figure \ref{pdquad}, we see that the preferred outcome is AA which is welfare-maximizing in easy PD.
-This is not what game theory predicts for rational actors, instead, participants are being nice and cooperative with each other.
+A first observation is that the generally preferred option, across games and conditions, is AA, the higher-value equal outcome spot.
+Notably, this is not the Nash equilibrium that is predicted by game theory for rational actors in a finite time horizon situation.
+However, human participants seem to be nicer and more cooperative with each other, regardless of whether chat is available.
 
-Overall, for PD, the preference was to align on the fair-but-high square, regardless of whether chat was available.
-In the hard PD, AA selections were less common (`r stats_text(PD_cooperate, 2)`) and AA selections were slightly lowered still by a hard PD x chat interaction (`r stats_text(PD_cooperate, 5)`).
+In hard PD trials, AA selections were less common than in easy PD trials (`r stats_text(PD_cooperate, 2)`) and AA selections were slightly lowered still by a hard PD x chat interaction (`r stats_text(PD_cooperate, 5)`), when all experiments were analyzed together.
 The interaction effect had a credible interval overlapping 0 when experiments 3 or 5 were analyzed individually.
-In parallel, there was a hard PD x chat interaction increasing the rate of AB/BA selections (`r stats_text(PD_unneven, 5)`), consistent with the idea that this is the best outcome for hard PD, but requires more complex coordination to achieve.
+
+In parallel, the rate of AB/BA selections was increased by a hard PD x chat interaction (`r stats_text(PD_unneven, 5)`), consistent with the idea that from a total points perspective, this is the best outcome for hard PD, but requires more complex coordination to achieve.
 
 Overall, the results based on condition assignment are consistent with the prediction that language would be more beneficial for BoS than for PD.
-However, conditional differences are small and often noisy (presumably due to low update of using language), and so the experiments do not provide compelling evidence.
+However, condition differences were small and often noisy (presumably due to low update of using language), and so the experiments do not provide compelling evidence.
+Additionally, the analytic import of using well-studied game theory scenarios with predictions about the utility of language is weakened by the inconsistency between predicted behaviors in the PD condition and actual selections.
+
+TODO you are here!
 

### 2025-06-19 manuscript/manuscript.Rmd
citations

@@
 ---
-title: "Game theory and language: a cautionary tale"
+title: "Use of communication channels in 2x2 strategic games"
 authors:
@@
 abstract: |
-  An attempt was made TODO
+  How do we use language to coordinate with each other?
+This question is of interest both from the cognitive science of language perspective as well as the game theory perspective.
+Here we attempt to bridge the gap between the empirical tradition of studying language use for coordination in reference games with the theoretical and analytic framework of game theory.
+We conducted a series of 5 real-time online experiments where pairs of participants were matched to play a sequence of 40 2x2 games together (total N=616).
+Across the experiments, we varied the exact structure of trials, but all experiments involved trials with reward structures modeled on Prisoner's Dilemma (PD) and Bach or Stravinsky (BoS).
+We manipulated the presence or absence of a chat interface as a between-pairs manipulation.
+Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and nochat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games.
+In contrast to the predictions of rational actors, participants in both chat and no-chat conditions generally aligned on both cooperating rather than both defecting in PD.
+Overall, we were limited in our ability to discern the role of communication channels in game strategies due to low uptake in using the communication channels when available.
+
 bibliography: references.bib
@@
 
-How do we use language to coordinate with each other?
+How do we use language to coordinate with each other? This question is of interest both from the cognitive science of language perspective as well as the game theory perspective.
 
-The use of language and the formation of linguistic conventions in pure reference games is studied frequently, but the role of language is studied less in other situations with more complicated goals that may not be fully aligned between participants.
+In studies of conversation, there's work on how people align with each other in their word use when doing joint task and how repeatedly referring to the same targets in a matching game leads to the formation of linguistic conventions.
+As reference is a key part of other linguistic speech acts such as negotiation, coordinating actions, or persuasion, one might wonder whether referential coordination can also be seen during interactions in pursuit of these more complex goals.
+In a similar vein, more complex situations may allow for linguistic alignment and convention formation not just of terms of reference but of terms for entire strategic approaches.
+One could also imagine alignment in approach, creating a conventional structure for how to coordinate or negotiate.
+(todo possible citations)
 
-In prior work, we tried to look at the role of language in a 3-player negotiation game where incentives were either fully-aligned or competitive between players.
+Despite the relevance of these questions, from both a linguistic perspective and an applied psychology perspective, the large body of work on conversational alignment and referential communication has rarely branched out to situations where incentives are not fully aligned between agents.
+
+In prior work, we tried to look at the role of language in a 3-player negotiation game where incentives were either fully-aligned or competitive between players [@mankewitz2021; @boyce2023b].
 We found some evidence that language use improved performance, and that the referential language used to pick out targets showed some signs of conventionalization over time.
@@
 
-Here, we attempted to bridge that gap, by looking at the role of language in two classic game theory games.
+Game theory provides a formalism for understanding rational action when different actors have different incentive structures over outcomes that are jointly determined by the choices of all the actors.
+There is some theoretical work analyzing how communication availability may shift the strategies in games and thus when communicative acts are credible signals versus cheap talk [@cooper1992; @foerster2021; @baliga2002].
+Game theory economic games have been used to empirically study questions such as social prediction [@vanbaar2021], emotional signals [@demelo2020], and coordination [@duguid2014].
+Some studies test predictions about the role of communication in economic games [@crandall2018; @cross; @crawford2019a].
+(todo: check/reread citations, possible find more citations?)
+
+In our current work, we attempted to bridge that gap, by looking at the role of language in two classic game theory games.
 By using set-ups that are well studied in the game theory literature, we had a rich body of work understanding how rational agents would play, and when linguistic coordination could be useful.
 
-We focus on two games from within the 2x2 game space: Prisoner's Dilemma (PD) and Bach or Stravinsky (BoS).
+We focused on two games from within the 2x2 game space: Prisoner's Dilemma (PD) and Bach or Stravinsky (BoS).
 
-TODO payout table
+In Prisoner's Dilemma (Figure \ref{payoff-PD}), each player individually prefers the rewards they get from playing B to the rewards they get from playing A (holding the other player's action constant).
+However, the outcome from both playing A is better for each player than the outcome from both playing B.
+Thus, the Nash equilibrium, where no player can benefit from making a unilateral change in action is BB, and so game theorists expect rational agents playing a one-shot game of PD to each choose B.
+The prediction from game theory is that availability of communication should not change behavior in a PD games because it would be "cheap talk" -- there is no incentive to be truthful about one's planned action.
 
-From game theory analysis, communication should not help in PD as it would be "cheap talk" -- that is, regardless of what the opponent player says, the equilibrium does not change, because there is not an incentive to be truthful.
-Thus, the prediction from game theory is that among rational agents playing a one-shot game of PD, they will end up in the Nash equilibrium.
-
-For BoS, there are two pure Nash equilibrium, and thus there is a coordination challenge to agree on which one to go to.
-In this case, a message from one player saying what they will select will be credible -- having said that, they are incentivized to act on it, and their partner can do better by listening to it.
-There may be competition over which of the two equilibrium options to agree on, but both players want to go for the same cell.
-Thus, the prediction is that language will be beneficial in BoS.
-
-To foreshadow the results, we found that *using* language showed some effect of helping improve results when people used it (including the possibility of cross-round transfer).
-However, the rate of uptake of the chat interface was low (despite our best efforts), and so we did not find robust effects of condition (chat or no chat).
-We also cannot know that differences between *those who could use and did use the chat* and other groups were due to the chat use as opposed to selection confounds on who chooses to use the chat.
-A qualitative look at the language and post-experiment reports reveals various ways overall strategies for achieving fairness (in both conditions) and various strategies for identifying the target each round (when language was used).
-
-Thus, we still suspect that there probably is something interesting about when language is strategically useful and what sort of procedural and referential language is used.
-That said, we are unable to provide evidence because over 5 experiments (+ pilots) we continually ran into issues getting sufficient uptake of the condition manipulation, from which we eventually concluded that it was not worth our running additional experiments.
-
-All data and code is available.
+In Bach or Stravinsky (Figure \ref{payoff-BoS}), both players only receive rewards if they choose the same option.
+However, player X prefers the AA option and B prefers the BB option.
+Thus, there are two Nash equilibria, either AA or BB.
+In a one-shot game of BoS, there is no way for participants to break the symmetry.
+Communication is predicted to be potentially useful, if it allows participants to break this symmetry and agree on which cell to align on.
 
@@
 
-All experiments were implemented in Empirica (v1). The experiment consisted first of consent and instruction.
+All experiments were implemented in Empirica v1 [@almaatouq2021]. The experiment consisted first of consent and instruction.
 
@@
 
-```{r,fig.cap="The user interface. Participants saw a reward table and selected what box to open. \\label{ui}"}
+```{r,out.width=".8\\textwidth", fig.cap="The user interface. Participants saw a reward table and selected what box to open. \\label{ui}"}
 knitr::include_graphics("overview.png")
@@
   kable(caption = "Number of participants in each experiment and each condition. \\label{participants}")
+
+total_n <- all_rounds |> select(gameId) |> unique() |> tally() |> mutate(n=2*n)
 ```
@@
 
-```{r chat-pct, fig.width=10, fig.height=3, fig.cap="In chat conditions, how many trials each game choose to use the chat at all.
+```{r chat-pct, fig.width=10, fig.height=2.5, fig.cap="In chat conditions, how many trials each game choose to use the chat at all.
 Within each experiment, games are arranged in frequency of chat use.
@@
 
-```{r points, fig.cap="Rewards achieved per trial depending on experiment, game condition, and chat condition. \\label{points}"}
+```{r points, fig.cap="Rewards achieved per trial depending on experiment, game condition, and chat condition. \\label{points}", fig.height=4, fig.width=8}
 
@@
   geom_hline(aes(yintercept = .5)) +
+  coord_cartesian(ylim=c(0,1))+
   labs(y = "Good outcome?", x="Chat condition", color="")+
@@
 We expect normal PD to generally be the of the type easy PD; and again note that the errors in displaying off-diagonal rewards would primarily affect hard PD type trials, where off diagonal options are more attractive.
-The rates of selecting each of the three outcome options by experiment and conditions are shown in Figure \ref{quad}.
+The rates of selecting each of the three outcome options by experiment and conditions are shown in Figure \ref{pdquad}.
 
@@
     scale_color_manual(values=pal)+
-
+  coord_cartesian(ylim=c(0,1))+
   stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.3)) +
@@
 
-TODO you are here!
-
 ## Uptake results
 
-To confirm that language use is, indeed useful, we also provide visualizations looking at both the option and choice to use language on a trial by trial basis.
-Note that a pairs' choice to use language is not assigned and thus this exploratory analysis could be confounded.
+Given the low rates of chat usage even in the chat condition, we also provide visualizations of outcomes split out by both the option and choice to use language on a trial by trial basis.
+We note that a pairs' choice to use language is not assigned and thus that the observed effect of using chat encompasses both the direct effect of using chat and any selection effects related to which people, when presented with the option to chat, choose to do so.
 
@@
     chat_cond =="nochat" ~ "nochat",
-    is_chat == 1 ~ "chatted",
+    is_chat == 1 ~ "chatty",
     is_chat == 0 ~ "quiet"
-  ) |> factor(levels=c("chatted", "quiet", "nochat")))
+  ) |> factor(levels=c("chatty", "quiet", "nochat")))
 ```
 
-In BoS: P1 prefers AA to BB, P2 prefers BB to AA. AB and BA are 0 for both.
+We repeat the quadrant analyses, now with a three-way split between "chatty" games who used the chat on the given trial, "quiet" games who were in the chat condition, but produced no chat messages on the given trial, and "nochat" games were assigned to the no chat condition.
 
@@
   facet_grid(~ expt) +
+    coord_cartesian(ylim=c(0,1))+
   scale_color_manual(values=pal)+
@@
 ```
-For BoS results, shown in Figure \ref{bosquad-uptake}, the outcomes are best on trials that chatted.
-Depending on experiment, trials that could talk are either intermediate, or pattern with the nochat condition.
-This pattern suggests that some of the benefits to chatting spill over onto other trials, perhaps via setting up conventions for determining what strategy to use to break the symmetry.
-However, most of the overall slight benefits from the chat condition are being driven by the small number of games and trials that used the chat.
+For BoS results, shown in Figure \ref{bosquad-uptake}, chatty trials had the best outcomes, with high levels of aligning.
+Depending on experiment, quiet trials generally patterned with the nochat condition, expect in the first experiment that was entirely BoS.
+Thus, it appears that there may be slight benefits to chatting that spill over onto other trials, perhaps via setting up conventions for determining what strategy to use to break the symmetry.
+Strategy set-up may be easiest when the trials are most homogeneous.
+However, most of the overall slight benefits associated with the chat condition are being driven by the small number of games and trials that used the chat.
 
@@
 We similarly break down the PD results in terms of language use per trial in Figure \ref{pdquad}.
-Noteably, the selection of AB/BA options in hard PD is primarly being driven by trials where people did communicate with each other.
+Trials that used chat were slightly more consistent at selecting the cooperate option on easy and normal PD (although this in particular could reflect selection effects for careful or attentive participants).
+The larger difference in selections between trials that used chat and those that did not was on the hard PD trials where selection of the AB/BA option appears to be primarily driven by chat-usage.
+
+Overall, the slight differences observed in the chat condition and nochat conditions appear to primarily come from the relatively small number of trials that actually used the chat while quiet trials generally patterned with behavior from the nochat condition.
 
@@
 From the uptake analysis, it appears that using chat has some benefits, especially on BoS trials.
-Does the amount of language used matter, or only its existance?
+Did the amount of language used affect performance, or only its existence?
 In Figure \ref{wordpayout}, we compare rewards on a trial as a function of the number of words used, for trials where any language was used.
-Overall, there does not seem to be a benefit to using more words on a trial.
+Overall, we do not observe a benefit to using more words on a trial.
 
@@
 The utility of language, but without a clear dose-response relationship, suggests that the language may primarily be used for coordination rather than negotiation.
+A qualitative perusal of the data reveals various strategies that pairs employ.
+One approach is for one player to identify the cell they will both choose, either by the color or color combination (sometimes even reduced to single letters, such as "r" for red) or by the numeric values.
 
-A qualitative perusal of the data suggests various strategies.
-On a trial by trial reference basis, cells can be identified by color or color combination, sometimes even reducing to single letter (such as r for red).
-The numeric values can also be used to identify.
-
-Between the chat logs and the post game strategy notes, a number of strategies were identified.
+Between the chat logs and the post game strategy notes, some longer-run strategies were identified.
 For example, some participants coordinated using the running score totals choosing the option that favored the person with a lower current score.
-Other identified strategies included fairness, being generous to the other person, or generally being fair but occassionally trying to take advantage and get a higher payout.
+Other identified strategies included fairness, being generous to the other person, or generally being fair but occasionally trying to take advantage and get a higher payout.
 
 # Discussion
-Some caveats: Despite our desire to use game-theory aligned paradigms for the analytic theory and predictions, the experiments had several factors that cause potential theoretical misalignment.
 
-1. For practical reasons, we ran repeated trials with the same partners. This iterated approach is also studied, but potentially changes what exactly the predictions are.
+Our intention with this line of research was to bridge between the literature on linguistic coordination and referential language use in reference games with the literature on strategic games and the value of communication from game theory.
+We tested how people would use a linguistic communication channel while playing a series of 2x2 strategic games with a partner, and how the outcomes of these 2x2 games would vary between people with the communication channel and those without.
+
+Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and nochat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games and slightly more likely to choose welfare-maximizing unequal outcomes in hard PD games.
+In contrast to the predictions of rational actors, participants in both chat and no-chat conditions generally aligned on both cooperating rather than both defecting in PD.
+The tendency to cooperate has previously been observed in experiments with repeated PD-style games [@rand2013].
+Overall, we were limited in our ability to discern the role of communication channels due to low uptake in using the communication channels when available.
+
+Despite our desire to use game-theory aligned paradigms for the analytic theory and predictions, the experimental paradigm contained several design factors that misalign with the assumptions of the analytic theory.
+
+1.
+We ran repeated trials with the same pairs to amortize costs related to instructing and pairing participants and letting participants get used to the games and interface.
+This iterated approach is also studied in game theory, but potentially changes what exactly the predictions for optimal play are.
 
 2.
-The incentives of participants are not the same as the incentives of the "agents".
-People have values above money such as their reputation and desire to be nice to other people.
-Additionally, the monetary value participants get from the task is not what is shown in the boxes.
-They were paid a base rate for participation + a bonus aligned to task performance.
-Strategies of random clicking could be cost efficient if it means the task takes sufficiently shorter that base rate / short time is a better wage than (base rate + higher bonus)/ longer time.
+The incentives of participants were not the same as the incentives of the "agents".
+People have values above money such as their reputation and desire to be nice to other people, even when other people are strangers on the internet.
+Additionally, the monetary value participants got from the task was not equivalent to the rewards from the trials.
+Participants were paid a base rate for participation plus the bonus aligned to task performance.
+Thus, their wage from this task depended both on how long the game took them and how well they did on trials.
+Strategies of random clicking or using easy to implement heuristics may have been higher utility than slower strategies using communication depending on how participants valued higher bonuses versus shorter completion times.
 
-3. The repeated nature of the game and extra information such as running score allowed for coordinating in ways not available in the base single-shot games.
+3.
+The repeated nature of the game, mix of trial types, and extra information such as running score allowed for coordination around alternating or evening out scores in the long run in ways not available in the base single-shot games.
 
-One of the hopes was to see if using language on some perhaps early trials would set up conventions or patterns that could then be successfully executed on later with no or little additional language.
+In conclusion, we still think there are interesting questions about how communication, and linguistic communication in particular, might affect what strategies around coordination and competition can be used in strategic games.
+Coming from the perspective of convention formation in reference games, we are still interested in when and how people can form procedural or "meta" conventions around how to negotiate with each other rather than just around referents.
+However, over the course of these experiments, we have learned that there are methodological and implementational barriers.
+We were able to mostly overcome initial issue with participants thinking they were playing with bots by adding an initial short free chat between each partnership.
+However, we were not able to overcome other problems.
+We had difficulties finding tasks that have tractable theoretical predictions but are also sufficiently engaging that humans will take them seriously, and compromised on a design that was not ideal from either perspective.
+Additionally, while we previously had no problems getting participants to use linguistic communication in online chat-box mediated tasks where the goal was reference [@boyce2024], we had difficulty getting participants to use chat in strategic contexts [both here and in @mankewitz2021].
 
-# Appendix: Model Results
+It is possible that other research teams could design experiments, recruitment contexts, interaction contexts, or instructions that produce different baseline preferences around the use of communication, and could lead to more successful uptake of communication channel manipulations.
+We hope this is true, as we think the research questions are interesting.
+
+# Appendix: Model Results {-}
 
 Here we present the statistical results we ran.
-Note that these are not all the statistics we ever ran and may not be the pre-registered ones.
-These are what I thought best summarized the data descriptively and addressed the hypotheses at the time this summary was written.
+Note that these are not all the statistics we ever ran and may not necessarily line up with the analyses we pre-registered.
+Instead, these represent what we thought best summarized the data descriptively and addressed the hypotheses at the time this summary was written.
 
@@
 
-## Payout models
+## Payout models {-}
 
@@
 Model were run both over all the experiments, and individually on each of experiments 3-5.
-Due to errors, we do not run separate models on 1 and 2.
+Due to errors in experimental display, we do not run separate models on 1 and 2.
 
@@
 
-## BoS models
+## BoS models {-}
 
@@
 
-## PD models
+\FloatBarrier
+
+## PD models {-}
 
@@
 ```
+
+# Bibliography {-}

### 2025-07-16 manuscript/manuscript.Rmd
fix typos

@@
 We manipulated the presence or absence of a chat interface as a between-pairs manipulation.
-Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and nochat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games.
+Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and no-chat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games.
 In contrast to the predictions of rational actors, participants in both chat and no-chat conditions generally aligned on both cooperating rather than both defecting in PD.
@@
 One could also imagine alignment in approach, creating a conventional structure for how to coordinate or negotiate.
-(todo possible citations)
 
@@
 
-In prior work, we tried to look at the role of language in a 3-player negotiation game where incentives were either fully-aligned or competitive between players [@mankewitz2021; @boyce2023b].
+In prior work, we tried to look at the role of language in a 3-player negotiation game where incentives were either fully-aligned or competitive between players [@mankewitz2021; @boyce2023flowers].
 We found some evidence that language use improved performance, and that the referential language used to pick out targets showed some signs of conventionalization over time.
@@
 Game theory economic games have been used to empirically study questions such as social prediction [@vanbaar2021], emotional signals [@demelo2020], and coordination [@duguid2014].
-Some studies test predictions about the role of communication in economic games [@crandall2018; @cross; @crawford2019a].
-(todo: check/reread citations, possible find more citations?)
+Some studies test predictions about the role of communication in economic games [@crandall2018; @cross2020; @crawford2019a].
 
@@
 In Bach or Stravinsky (Figure \ref{payoff-BoS}), both players only receive rewards if they choose the same option.
-However, player X prefers the AA option and B prefers the BB option.
+However, player X prefers the AA option and player Y prefers the BB option.
 Thus, there are two Nash equilibria, either AA or BB.
@@
 We recruited participants via Prolific and directed them to our website where they were paired up to play a game with another participant in real time.
+
 ## Procedure
-
 All experiments were implemented in Empirica v1 [@almaatouq2021]. The experiment consisted first of consent and instruction.
@@
 Each pair completed 40 trials, with the payoffs dependent on the experiment condition (see below).
-Participants in the "chat" condition could message their partner freely during the trials using a chat interface; participants in the "no chat" condition did not have access to the chat during the main experiment.
+Participants in the "chat" condition could message their partner freely during the trials using a chat interface; participants in the "no-chat" condition did not have access to the chat during the main experiment.
 
-In experiments 1,2,3 and 5, participants saw a running bonus score for themselves and their partner, as their bonus payment was proportional to the sum of the rewards they earned.
+In experiments 1, 2, 3, and 5, participants saw a running bonus score for themselves and their partner, as their bonus payment was proportional to the sum of the rewards they earned.
 On experiment 4, participants were instead paid proportional to the results of 4 randomly selected trials, and were not shown a cumulative payoff score.
@@
 For each trial type, we constructed trials by randomly sampling reward values subject to some constraints.
-Examples of each trial type are shown in Figure \ref{payoff-matrix}
+Examples of each trial type are shown in Figure \ref{payoff-matrix}.
 
@@
 
-**Expt 4** (https://osf.io/fymjc): In experiment 4, we aimed for a set-up that was closer to the predictions from game theory.We removed the points counter, given that participants in prior experiments sometimes used the visible money earned so far as a locus for coordination.
+**Expt 4** (https://osf.io/fymjc): In experiment 4, we aimed for a set-up that was closer to the predictions from game theory.
+We removed the points counter, given that participants in prior experiments sometimes used the visible money earned so far as a locus for coordination.
 We switched from providing a bonus proportional to the overall points earned to an incentive compatible bonus that sampled 4 trials at random and paid out those points.
@@
   select(expt, chat_cond, pct) |>
-  mutate(chat_cond = ifelse(chat_cond == "chat", "chat", "nochat"))
+  mutate(chat_cond = ifelse(chat_cond == "chat", "chat", "no-chat"))
 
@@
 If both players cooperate, the outcome is AA and each gets a moderately high payoff.
-If both players defect, the outcome is BB, the Nash equilibrium, and each player get a low payoff.
-If one player cooperates and the other defects, the outcome is AB (or BA), and one player get a high payoff and the other gets 0.
+If both players defect, the outcome is BB, the Nash equilibrium, and each player gets a low payoff.
+If one player cooperates and the other defects, the outcome is AB (or BA), and one player gets a high payoff and the other gets 0.
 From a perspective of wefare-maximization, the selection with best total reward is AA in easy PD and AB/BA in hard PD.
@@
 Overall, the results based on condition assignment are consistent with the prediction that language would be more beneficial for BoS than for PD.
-However, condition differences were small and often noisy (presumably due to low update of using language), and so the experiments do not provide compelling evidence.
+However, condition differences were small and often noisy (presumably due to low uptake of using language), and so the experiments do not provide compelling evidence.
 Additionally, the analytic import of using well-studied game theory scenarios with predictions about the utility of language is weakened by the inconsistency between predicted behaviors in the PD condition and actual selections.
@@
   mutate(chatted = case_when(
-    chat_cond =="nochat" ~ "nochat",
+    chat_cond =="nochat" ~ "no-chat",
     is_chat == 1 ~ "chatty",
     is_chat == 0 ~ "quiet"
-  ) |> factor(levels=c("chatty", "quiet", "nochat")))
+  ) |> factor(levels=c("chatty", "quiet", "no-chat")))
 ```
 
-We repeat the quadrant analyses, now with a three-way split between "chatty" games who used the chat on the given trial, "quiet" games who were in the chat condition, but produced no chat messages on the given trial, and "nochat" games were assigned to the no chat condition.
+We repeat the quadrant analyses, now with a three-way split between "chatty" games who used the chat on the given trial, "quiet" games who were in the chat condition, but produced no chat messages on the given trial, and "no-chat" games were assigned to the no-chat condition.
 
@@
 For BoS results, shown in Figure \ref{bosquad-uptake}, chatty trials had the best outcomes, with high levels of aligning.
-Depending on experiment, quiet trials generally patterned with the nochat condition, expect in the first experiment that was entirely BoS.
+Depending on experiment, quiet trials generally patterned with the no-chat condition, except in the first experiment that was entirely BoS.
 Thus, it appears that there may be slight benefits to chatting that spill over onto other trials, perhaps via setting up conventions for determining what strategy to use to break the symmetry.
@@
 Trials that used chat were slightly more consistent at selecting the cooperate option on easy and normal PD (although this in particular could reflect selection effects for careful or attentive participants).
-The larger difference in selections between trials that used chat and those that did not was on the hard PD trials where selection of the AB/BA option appears to be primarily driven by chat-usage.
+The larger difference in selections between trials that used chat and those that did not was on the hard PD trials where selection of the AB/BA option appears to be primarily driven by chat usage.
 
-Overall, the slight differences observed in the chat condition and nochat conditions appear to primarily come from the relatively small number of trials that actually used the chat while quiet trials generally patterned with behavior from the nochat condition.
+Overall, the slight differences observed in the chat condition and no-chat conditions appear to primarily come from the relatively small number of trials that actually used the chat while quiet trials generally patterned with behavior from the no-chat condition.
 
@@
 
-Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and nochat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games and slightly more likely to choose welfare-maximizing unequal outcomes in hard PD games.
+Despite low levels of use of the communication channel, we found slight differences in outcome between the chat and no-chat groups, where pairs who could communicate were slightly more likely to reach coordinated beneficial outcomes in BoS games and slightly more likely to choose welfare-maximizing unequal outcomes in hard PD games.
 In contrast to the predictions of rational actors, participants in both chat and no-chat conditions generally aligned on both cooperating rather than both defecting in PD.
@@
 However, over the course of these experiments, we have learned that there are methodological and implementational barriers.
-We were able to mostly overcome initial issue with participants thinking they were playing with bots by adding an initial short free chat between each partnership.
+We were able to mostly overcome the initial issue with participants thinking they were playing with bots by adding an initial short free chat between each partnership.
 However, we were not able to overcome other problems.
